### 2017-04-04 SpringRateProject/MSXII Supsension Parameters.tex
Opperation reduce complexity
Decided to approximate a 50% 50% wheel base with equivalent spring and
damping rates on the front and rear to reduce the problem’s complexity.

\documentclass[12pt]{article}

% Packages
\usepackage{lipsum}
\usepackage{setspace}
\usepackage{indentfirst}
\usepackage{geometry}
\usepackage[nonumberlist, toc]{glossaries}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage[numbered]{./latex/mcode}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage [autostyle, english = american]{csquotes}
\usepackage{fancyhdr}
\usepackage{amsmath}

% Formatting
\geometry{letterpaper, portrait, margin=.85in}

\pagestyle{fancy}
\lhead{MSXII Suspension Parameters}
\rhead{Midnight Sun Solar Car Team}

\MakeOuterQuote{"}

\begin{document}

% Title Ppage
\begin{titlepage}
	\vspace*{3cm}
	\centering
	\includegraphics[width=.2\textwidth]{./LaTex/midnightSunLogoCircle.png}\par
	\vspace{1.5cm}
	{\scshape\LARGE Midnight Sun Solar Car Team \par}
	{\scshape\large University of Waterloo\par}
	\vspace{1.5cm}
	{\huge\bfseries MSXII Suspension Parameters\par}
	\vspace{0.2cm}
	\large Prepared for Project 3 of ME321 Winter 2017
	\vfill
	Prepared by:\par
	Devon Copeland\par
	\vspace{1cm}
	\today\par
\end{titlepage}

% Main Matter
\section{Background}
\subsection{Vehicle Design}
Midnight Sun Twelve (MSXII) is a cruiser class solar electric vehicle being designed with the goal of competing in the 2018 American Solar Challenge (ASC 2018) and the 2019 World Solar Challenge (WSC 2019). Cruiser class vehicles are must be multi-occupant and are designed with the intent of being more practical than a typical solar car. Because of the unique requirements of this class, the spring rates and target damping coefficients on MSXII's suspension must be selected to optimize for efficient while still keeping driver comfort in mind. This report proposes and analytical technique for determining the response of the suspension system to a variety of road conditions.

\subsection{Important Vehicle Parameters}
Table \ref{tab:params} lists predicted vehicle parameters that are of importance to the design and analysis of the suspension. 
\begin{table}[]
	\centering
	\caption{Important Vehicle Parameters}
	\label{tab:params}
	\begin{tabular}{lll}
	Mass                             & 500  & $kg$                    \\
	Distance from Front Wheel to CoG & 1.32 & $m$                     \\
	Moment of Inertia                & 550  & $kgm^2$ \\
	Wheelbase                        & 1.6  & $m$                    
	\end{tabular}
\end{table}
\subsection{Front Suspension} 
MSXII's front suspension comprises of a double wishbone linkage with an outboard coilover. Important dimensions required for subsequent analysis are shown in Figure \ref{fig:}:

\subsection{Rear Suspension} 
MSXII's rear suspension comprises of independent trailing arms allowing for zero scrub and thus less rolling resistance. Dimensions required for subsequent analysis are shown in Figure \ref{fig:}:

\subsection{Selecting Spring Rates}

\section{Suspension Model}

\subsection{Overview}
MSXII's suspension is modelled using a point mass connected to two vertical, linear spring mass damper systems at the two wheels. This half car model assumes infinite tire stiffness and that the vehicle pitches about it's center of gravity. Figure \ref{fig:systemModel} describes the proposed model where: 
\begin{align*}
	\theta & \text{ is the pitch angle about positive x axis}\\        
	x_0 & \text{ is the displacement of the center of gravity in the vertical axis} \\
	x_1 \text{ \& } x_2 & \text{ are the displacements of the mounting points for the front and rear wheels respectively} \\
	y_1 \text{ \& } y_2 & \text{ is the change in road elevation below the front  and rear wheels respectively} \\
	M & \text{ is the mass of the vehicle} \\
	I & \text{ is the moment of inertia normal to the half car plane at the center of gravity} \\
	k1,k2,c1 \text{ \& } c_2 & \text{ are the spring and damping coefficients} \\
	a \text{ \& } b & \text{ are the distances from center of gravity to the front and rear wheels respectively}
\end{align*}
\begin{figure}[h!]
	\centering
	\label{fig:systemModel}
	\includegraphics[width=.8\textwidth]{./LaTex/systemModel.jpg}
	\caption{Half car model with infinite tire stiffness}
\end{figure}

\subsection{Equations of Motion}
The spring mass damper system shown in Figure \ref{fig:systemModel} can be described by the folowing two differential equations: 
\begin{equation}
	M \ddot x + c_1(\dot x_1 - \dot y_1) + c_2(\dot x_2 - \dot y_2) + k_1(x_1 - y_1) + k_2(x_2 - y_2) = 0
\end{equation}
\begin{equation}
	I \ddot\theta - ac_1(\dot x_1 - \dot y_1) + bc_2(\dot x_2 - \dot y_2) - ak_1(x_1 - y_1) + bk_2(x_2 - y_2) = 0
\end{equation}
Rearranging to move the inputs $y_1$ and $y_2$ to right hand side: 
\begin{equation}
	M \ddot x + c_1\dot x_1 + c_2\dot x_2 + k_1x_1 + k_2x_2 = c_1\dot y_1 + c_2\dot y_2 + k_1y_1 + k_2y_2
\end{equation}
\begin{equation}
	I \ddot\theta - ac_1\dot x_1 + bc_2\dot x_2 - ak_1x_1 + bk_2x_2 = -ac_1\dot y_1 + bc_2\dot y_2 - ak_1y_1 + bk_2y_2
\end{equation}
By assuming small pitch angles, $\theta$, the angle between the spring mass damper system becomes approximately perpendicular and the displacements $x_1$ and $x_2$ can be approximated as follows:
\begin{equation}
	x_1 = x_0 - a\theta
\end{equation}
\begin{equation}
	x_2 = x_0 + b\theta
\end{equation}
It follows that the time derivatives of the above two equations are: 
\begin{equation}
	\dot x_1 = x_0 - a\dot\theta
\end{equation}
\begin{equation}
	\dot x_2 = x_0 + b\dot\theta
\end{equation}
Substituting into the original differential equations: 
\begin{equation}
	M \ddot x + \dot x_0(c_2+c_1) + \dot\theta(bc_2-ac_1) + x_0(k_2+k_1) + \theta(bk_2-ak_1)= c_1\dot y_1 + c_2\dot y_2 + k_1y_1 + k_2y_2
\end{equation}
\begin{equation}
	I \ddot\theta + \dot x_0(bc_2-ac_1) + \dot\theta(b^2c_2-a^2c_1) + x_0(bk_2-ak_1) + \theta(b^2k_2+a^2k_1) = -ac_1\dot y_1 + bc_2\dot y_2 - ak_1y_1 + bk_2y_2
\end{equation}
By taking the unilateral laplace transform of the above two equations with all initial conditions set to zero, the following $s$ domain equations are obtained:
\begin{equation}
	\left[Ms^2+(c_2+c_1)s+(k_2+k_1)\right]X_0 + \left[(bc_2-ac_1)s+(bk_2+ak_1)\right]\Theta = c_1sY_1 + c_2sY_2 + k_1Y_1 + k_2Y_2
\end{equation}
\begin{equation}
	\left[(bc_2-ac_1)s+(bk_2-ak_1)\right]X_0 + \left[Is^2+ (b^2c_2+a^2c_1)s+(b^2k_2+a^2k_1)\right]\Theta = c_1sY_1 + c_2sY_2 + k_1Y_1 + k_2Y_2
\end{equation}
The above two equations can be expressed in matrix form as follows: 
\begin{equation}
	\label{eq:matrix1}
	\begin{bmatrix} 
	Q_{Ax}(s) & Q_{A\Theta}(s) \\
	Q_{Bx}(s) & Q_{B\Theta}(s)
	\end{bmatrix}
	\begin{bmatrix} 
	X_0 \\
	\Theta 
	\end{bmatrix}
	= 
	\begin{bmatrix} 
	P_{A1}(s) \\
	P_{B1}(s)
	\end{bmatrix}Y_1 +
	\begin{bmatrix} 
	P_{A2}(s) \\
	P_{B2}(s)
	\end{bmatrix}
	Y_2
\end{equation}
If it is assumed that the rear tire of the car experiences the exact same road conditions as the front tire, $Y1$ and $Y2$ can be expressed as functions of the same input, $Y$ where $Y2$ is simply a phase shifted version of $Y1$:
\begin{equation}
	\begin{split}
		y_1(t) &= y(t)\\
		y_2(t) &= y(t-t_0)
	\end{split} \implies
	\begin{split}
		Y_1(s) &= Y(s)\\
		Y_2(s) &= Y(s)e^{-st_0}
	\end{split}
\end{equation}
Equation \ref{eq:matrix1} can then be rewritten as follows:
\begin{equation}
	\label{eq:matrix1}
	Q
	\begin{bmatrix} 
	X_0 \\
	\Theta 
	\end{bmatrix}
	= 
	\left(\left[P_1\right]+\left[P_2\right]e^{-st_0}\right)Y
\end{equation}
Since the system is linear linear, the above matrix equation can be solved in two parts to obtain the transfer functions for $X_0$ and $\Theta$:
\begin{equation}
	\frac{\begin{bmatrix} 
		X_0 \\
		\Theta 
	\end{bmatrix}}{Y} =
	\left[H\right] = 
	\left[H_1\right] + \left[H_2\right] = 
	[Q^{-1}][P_1] + [Q^{-1}][P_1]e^{-st_0} 
\end{equation}

\section{Analysis}
\subsection{Natural Frequency}
Using the MATLAB script show in Appendix \ref{app:code}, the transfer function $H(s)$ is solved for symbolically and the parameters for $I$, $M$, $k_1$, $k_2$, $a$ and $b$ are substituted into the equation while the damping coefficients $c_1$ and $c_2$ are set to zero. The inverse laplace transform is then taken to obtain the impulse response as shown in Figure \ref{fig:impResp}.
\begin{figure}[h!]
	\centering
	\label{fig:impResp}
	\includegraphics[width=.8\textwidth]{./matlab/impResp.png}
	\caption{Undamped Impulse Response}
\end{figure}

By performing an Fast Fourier Transform on the undamped system the harmonics of the system can be approximated as shown in Figure \ref{fig:fftImpResp}. A sampling frequency of 200Hz and and a window of 30s was used for the FFT.
From figure \ref{fig:fftImpResp}, the largest harmonics of the system occur at Hz for translational displacement and Hz for angular displacement. 
\begin{figure}[h!]
	\centering
	\label{fig:impResp}
	\includegraphics[width=.8\textwidth]{./matlab/fftImpResp.png}
	\caption{FFT of the Undamped Impulse Response}
\end{figure}

\subsection{Critical Damping}

\subsection{Force Response}
\subsection{Worst Case Total Response}

% Appendix
\pagebreak
\appendix
\section{MATLAB Source Code}
\label{app:code}
\lstinputlisting[breaklines=true]{./matlab/createTransferFunctions.m}

\end{document}

### 2017-04-04 SpringRateProject/matlab/createTransferFunctions.m

@@ -1,0 +1,275 @@
+% ---------------------------------------------------------------------- %
+%
+% Suspension Spring Rates Project - createTransferFuncitons.m
+% Function setup transfer functions for a half car model to determine the
+% reponse of MSXII to disturbances in the road conditions
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2017
+%
+%
+% Notes:
+%   
+%   - 
+% ---------------------------------------------------------------------- %
+
+load_system('halfCarModel');
+close all;
+
+% ----------------------------- Constants ------------------------------ %
+
+c1_ = 0;     % Front damping coefficient (Ns/m)
+c2_ = 0;     % Rear damping coefficient (Ns/m) 
+k1_ = 50000;    % Front sring rate (N/m) 
+k2_ = 60000;    % Rear sring rate (N/m)
+m_ = 500;       % Mass of the vehicle (kg)
+I_ = 550;       % Moment of interial about global x axis (kg*m^2)
+a_ = 1.2;       % Distance from COG to front tire (m)
+b_ = 1.4;       % Distance from COG to rear tire (m)
+
+% ---------------------- Forcing Function Parameters ------------------- %
+
+amplitude = 0.015;              % Amplitude of bumps on road surface
+bumpSparation = 2;              % Distance between peaks on the road surface (m)
+speed = 22.2;                   % Cruising speed (m/s)
+wb = 2*pi*speed/bumpSparation;  % Frequency of base excitation (rad/sec)
+timeDelay = (a_+b_)/speed;        % delay between input striking front and
+                                % rear wheels (s)
+                                
+% Update simulink parameters
+% set_param('halfCarModel/Road','Frequency',mat2str(wb),'Amplitude',mat2str(amplitude));
+set_param('halfCarModel/DelayA','DelayTime',mat2str(timeDelay));
+set_param('halfCarModel/DelayB','DelayTime',mat2str(timeDelay));
+                              
[…]

### 2017-04-04 SpringRateProject/matlab/timeDomainAnalysis.m

@@ -1,0 +1,25 @@
+syms t c1 c2 k1 k2 m I a b t0 A wb xddot xdot x thetaddot thetadot theta c k
+
+Ax2 = m;% X terms of ODE (A) 
+Ax1 = (c1+c2) ; 
+Ax0 = (k1+k2);
+Atheta2 = 0;
+Atheta1 = (b*c2-a*c1);
+Atheta0 = (b*k2-a*k1);
+Bx2 = 0;
+Bx1 = (b*c2-a*c1); 
+Bx0 = (b*k2-a*k1); 
+Btheta2 = I;
+Btheta1 = (b^2*c2+a^2*c1);
+Btheta0 = (b^2*k2+a^2*k1);
+
+PA = c1*A*wb*cos(wb*t) + c2*A*wb*cos(wb*(t-t0)) + k1*A*sin(wb*t) + k2*A*sin(wb*(t-t0)); 
+PB = -c1*a*A*wb*cos(wb*t) + c2*b*A*wb*cos(wb*(t-t0)) + k1*a*A*sin(wb*t) + k2*b*A*sin(wb*(t-t0)); 
+
+% Create linear systems in to solve for transfer functions
+
+A = [Ax2, Ax1, Ax0, Atheta2, Atheta1, Atheta0, PA;
+     Bx2, Bx1, Bx0, Btheta2, Btheta1, Btheta0, PB];
+ 
+A = subs(A, {b, c1, c2, k1, k2}, {a, c, c, k, k});
+A

### 2017-04-04 SpringRateProject/MSXII Supsension Parameters.tex
Finished first draft of spring rate project

@@
 \usepackage{amsmath}
+\usepackage{cite}
 
@@
 	\centering
-	\includegraphics[width=.2\textwidth]{./LaTex/midnightSunLogoCircle.png}\par
+	\includegraphics[width=.25\textwidth]{./LaTex/midnightSunLogoCircle.png}\par
 	\vspace{1.5cm}
@@
 	{\scshape\large University of Waterloo\par}
-	\vspace{1.5cm}
+	\vspace{3.5cm}
 	{\huge\bfseries MSXII Suspension Parameters\par}
@@
 \subsection{Important Vehicle Parameters}
-Table \ref{tab:params} lists predicted vehicle parameters that are of importance to the design and analysis of the suspension. 
-\begin{table}[]
+Table \ref{tab:params} lists predicted vehicle parameters that are of importance to the design and analysis of the suspension. The mass and moment of inertia was computed by CAD software. 
+\begin{table}[htbp]
 	\centering
@@
 	\begin{tabular}{lll}
-	Mass                             & 500  & $kg$                    \\
-	Distance from Front Wheel to CoG & 1.32 & $m$                     \\
-	Moment of Inertia                & 550  & $kgm^2$ \\
-	Wheelbase                        & 1.6  & $m$                    
+	Mass ($M$)                            	& 550  & $kg$       \\
+	Wheelbase                       	 	& 2.60  & $m$  		\\
+	Track                       	 		& 1.60  & $m$  		\\
+	Distance from Front Wheel to CoG ($a$)	& 1.32 & $m$  		\\
+	Distance from Road to CoG ($h$)		& 0.52 & $m$  		\\
+	Moment of Inertia Normal to Symmetry Plane at CoG ($I_{xx}$) & 550  & $kgm^2$ \\
+	Full Wheel Travel in Compression 	& 45   & $mm$ 		\\
+	Front Coilover Angle from Vertical	& 40   & $\circ$ 	\\
+	Rear Coilover Angle from Vertical	& 30   & $\circ$
 	\end{tabular}
 \end{table}
-\subsection{Front Suspension} 
-MSXII's front suspension comprises of a double wishbone linkage with an outboard coilover. Important dimensions required for subsequent analysis are shown in Figure \ref{fig:}:
+\subsection{Suspension Architecture}
+MSXII's front suspension comprises of a double wishbone linkage with an outboard coilover while the rear suspension comprises of independent trailing arms allowing for zero scrub and thus less rolling resistance.
+Figure \ref{fig:solidworksScreenshots} shows the screenshots of the front and rears suspension.
+\begin{figure}[htbp]
+	\centering
+    \begin{subfigure}[b]{.32\textwidth}
+		\caption{Front Isometric View}
+		\centering
+        \includegraphics[height=4cm]{./LaTex/frontIso.PNG}
+    \end{subfigure}
+    \begin{subfigure}[b]{.32\textwidth}
+		\caption{Rear Isometric View}
+		\centering
+        \includegraphics[height=4cm]{./LaTex/rearIso.PNG}
+        \label{fig:c8}
+    \end{subfigure}
+    \begin{subfigure}[b]{.32\textwidth}
+		\caption{Rear Sie View}
+		\centering
+        \includegraphics[height=4cm]{./LaTex/rearRight.PNG}
+        \label{fig:c9}
+    \end{subfigure}
+    \caption{MSXII Suspension Design (Screenshots from Solidworks)}
+	\label{fig:solidworksScreenshots}
+\end{figure}
 
-\subsection{Rear Suspension} 
-MSXII's rear suspension comprises of independent trailing arms allowing for zero scrub and thus less rolling resistance. Dimensions required for subsequent analysis are shown in Figure \ref{fig:}:
+\subsection{Selecting Suspension Parameters}
+\label{sec:paramSelection}
+The spring rates are to be calculated based on the travel of the wheel and the maximum loading conditions.
+Assuming a smooth driving surface, the largest normal force acting on the tire from the road will occur while braking in a turn.
+To approximate this loading condition the a superposition of two static half car models is used; one for cornering and one for hard braking.
+\subsubsection{Hard Braking}\begin{figure}[h!]
+	\centering
+	\includegraphics[width=.5\textwidth]{./LaTex/brakingFBD.jpg}
+	\caption{Free Body Diagram of a Car Braking}
+	\label{fig:barkingFBD}
+\end{figure}
+Figure \ref{fig:barkingFBD} shows a free body diagram of a half car model undergoing braking.
+MSXII will only have brake callipers on the front wheels however regenerative braking from the hub motors will also provide a braking force at the rear contact patch.
+Assuming a generous coefficient of static friction, $\mu _s$, of 1.0 and the extreme case where the both the font and rear tires are about to begin sliding the combined braking force can be expressed as:
+\begin{equation}
+	F_{braking} = F_{Nf} + F_{Nr} = \mu _s F_{N_{net}} =  \mu _s Mg = (1.0)(550kg)\left(9.81\frac{m}{s^2}\right) = 5.40kN
+\end{equation}
+Summing the moments about the centre of gravity and the vertical forces results in the following equations: 
+\begin{equation}
+	aF_{Nf} = hF_{braking} + bF_{Nr}
+\end{equation}
+\begin{equation}
+	F_{Nf} + F_{Nr} = Mg
+\end{equation}
+Solving for the front normal force: 
+\begin{equation}
+\begin{split}
+	aF_{Nf} &= hF_{braking} + b(Mg - F_{Nr})\\
+	F_{Nf} &= \frac{hF_{braking} + bMg}{a+b} = \frac{(0.52m)(5.40kN)+(1.28m)(550kg)\left(9.81\frac{m}{s^2}\right)}{(1.32m)+(1.28m)} = 2.66kN
+\end{split}
+\end{equation}
+Note that the 2.66kN above is for both front wheels. Therefore there is approximately \textbf{1.33kN} of normal force on each front tire during hard braking.
 
-\subsection{Selecting Spring Rates}
+\subsubsection{Cornering}
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=.5\textwidth]{./LaTex/dynamicScrutineering.PNG}
+	\caption{Figure of Eight Course from ASC 2018 Regulations 
+	\cite{ASC2018regs}}
+	\label{fig:dynamicScrutineering}
+\end{figure}
+The ASC 2018 regulations stipulate that a vehicle must be able to navigate a figure of eight as shown in Figure \ref{fig:dynamicScrutineering} in less than 18 seconds \cite{ASC2018regs}.
+Since the average arc length of the figure of eight is $34\pi$, the net cornering force required to navigate the figure of eight can be found as follows:
+\begin{equation}
+	F_{si} + F_{so} = F_{steer} = \frac{Mv^2}{r} = \frac{(550kg)\left(\frac{34\pi m}{18s}\right)^2}{8.5m} = 2.28kN
+\end{equation}
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=.5\textwidth]{./LaTex/steerFBD.jpg}
+	\caption{Free Body Diagram of a Car Cornering}
+	\label{fig:steerFBD}
+\end{figure}
+For the half car model shown if Figure \ref{fig:steerFBD} umming the moments about the centre of gravity and the vertical forces results in the following equations: 
+\begin{equation}
+	\frac{track}{2}F_{Nf} = hF_{steer} + \frac{track}{2}F_{Nr}
+\end{equation}
+\begin{equation}
+	F_{No} + F_{Ni} = Mg
+\end{equation}
+Solving for the front normal force: 
+\begin{equation}
+\begin{split}
+	\frac{track}{2}F_{No} &= hF_{steer} + \frac{track}{2}(Mg - F_{Ni})\\
+	F_{No} &= \frac{hF_{steer} + \frac{track}{2}Mg}{track} = \frac{(0.52m)(2.28kN)+(0.80m)(550kg)\left(9.81\frac{m}{s^2}\right)}{1.60m} = 2.70kN
+\end{split}
+\end{equation}
+Note that the 2.70kN above is for both front and rear outside wheels.
+Assuming the force is evenly split between the front and rear, there is approximately \textbf{1.35kN} of normal force on each outside tire during maximum cornering.
+\subsubsection{Spring Rate Selection}
+Adding the maximum normal forces from cornering and braking results in 2.68kN.
+The maximum travel of the suspension for MSXII in compression has been set to 45mm.
+It is desired that under the combined of braking and cornering, the suspension compresses to a maximum of 70\% of the travel.
+As such, the desired spring rate is found to be:
+\begin{equation}
+	k = \frac{force}{displacement} = \frac{2.68N}{(70\%)(0.045m)} = 85.0kN/m
+\end{equation}
+Note that the above rate assumes a vertical spring.
+The vertical spring rate can be mapped to an equivalent spring rate by dividing by the cosine of the angle that the coilover makes with the vertical direction:
+\begin{equation}
+	\begin{split}
+		k_{front} &= \frac{85.0}{cos(35)} = 110.6kN \\
+		k_{rear} &= {85.0}{cos(30)} = 98.1kN
+	\end{split}
+\end{equation}
+MSXII will be using Ohlins TTX25 MkII coilovers for both the front and rear.
+These coilovers can be configured to have spring rates from 150 lb/in to 650 lb/in in increments of 50 lb/in.
+1 lb/in is approximately 175 N/m therefore 550 lb/in coils should be selected for the front and 650 lb/in coils for the rear.
+
+\subsubsection{Damping Coefficient Selection}
+Ohlins has published curves for their TTX25 MkII's however real world damping rates are non constant.
+For this report, the damping coefficient will be idealized as constant and will be selected such that the damping ratio, $\zeta$, is 0.5.
+This is typical of performance consumer vehicles \cite{dampers}.
+Given an equivalent vertical spring rate or 85.0kN/m, a damping ratio of 0.5 maps to a vertical damping coefficient of \textbf(4.83kNs/m).
 
 \section{Suspension Model}
-
 \subsection{Overview}
@@
 	\centering
-	\label{fig:systemModel}
 	\includegraphics[width=.8\textwidth]{./LaTex/systemModel.jpg}
 	\caption{Half car model with infinite tire stiffness}
+	\label{fig:systemModel}
 \end{figure}
@@
 \subsection{Equations of Motion}
+\subsubsection{Dynamic Equilibrium}
 The spring mass damper system shown in Figure \ref{fig:systemModel} can be described by the folowing two differential equations: 
@@
 \begin{equation}
-	I \ddot\theta + \dot x_0(bc_2-ac_1) + \dot\theta(b^2c_2-a^2c_1) + x_0(bk_2-ak_1) + \theta(b^2k_2+a^2k_1) = -ac_1\dot y_1 + bc_2\dot y_2 - ak_1y_1 + bk_2y_2
+	I \ddot\theta + \dot x_0(bc_2-ac_1) + \dot\theta(b^2c_2+a^2c_1) + x_0(bk_2-ak_1) + \theta(b^2k_2+a^2k_1) = -ac_1\dot y_1 + bc_2\dot y_2 - ak_1y_1 + bk_2y_2
 \end{equation}
-By taking the unilateral laplace transform of the above two equations with all initial conditions set to zero, the following $s$ domain equations are obtained:
+The above two equations are considerably complex and are not particularly conducive of solving analytically by hand.
+However, the fact that the vehicle has the center of gravity approximately centered along the wheel offers an avenue for simplification.
+Additionally if the front and rear spring and damping coefficients are assumed to be the same ($k$ and $c$ respectively), the problem's complexity is drastically reduced as shown in the following two equations:
+\begin{equation}
+	M \ddot x + 2c\dot x_0 + 2kx_0 = c\dot y_1 + c\dot y_2 + ky_1 + ky_2
+\end{equation}
+\begin{equation}
+	I \ddot\theta + 2a^2c\dot\theta + 2a^2k\theta = -ac\dot y_1 + bc_2\dot y_2 - ak_1y_1 + bk_2y_2
+\end{equation}
+
+\subsubsection{Natural Frequency}
+The natural frequency of the translational and rotational components of the system are found as follows: 
+\begin{equation}
+	\begin{split}
+	\omega _{nx} &= \sqrt{\frac{k_{eq}}{m_{eq}}} = \sqrt{\frac{2k}{M}}\\
+	\omega _{n\Theta} &= \sqrt{\frac{k_{eq}}{m_{eq}}}= \sqrt{\frac{2a^2k}{I}}
+	\end{split}  
+\end{equation}
+
+\subsubsection{Critical Damping}
+The critical camping coefficient for the translational and rotational components of the system are found as follows:
+\begin{equation}
+	\begin{split}
+	c_{eqx} = 2c, c_{rx} = 2m_{eq}\omega _n = \sqrt{8kM}, \zeta _x = \frac{c_{eqx}}{c_r}\\
+	c_{eq\Theta} = 2a^2c, c_{r\Theta} = 2m_{eq}\omega _n = \sqrt{8a^2kI}, \zeta _\Theta = \frac{c_{eq\Theta}}{c_r}
+	\end{split}  
+\end{equation}
+
+\subsubsection{Final Equations fo Motion}
+The road displacement can be approximated by a sinusoid with and amplitude $A$ and a base excitation frequency $\omega _b$.
+Since rear tires experience the parts to the road as the front $y_2$ can be expressed as a phase shifted version of $y_1$.
+Combining this with the natural frequency and damping ratio results in the following equations of motion:
+\begin{equation}
+	\label{eq:eomx}
+	\ddot x + (2\zeta _x M \omega _{nx}) \dot x_0 + (\omega _{nx}^2)x_0 = \frac{cA\omega _b}{M}cos(\omega _b t) + \frac{cA\omega _b}{M}cos(\omega _b (t-t_0))  + \frac{kA}{M}sin(\omega _b t) + \frac{kA}{M}sin(\omega _b (t-t_0))
+\end{equation}
+\begin{equation}
+	\label{eq:eomtheta}
+	\begin{split}
+		\ddot\theta + &(2\zeta _\Theta  I \omega _{n\Theta}) \dot\theta + (\omega _{n\Theta}^2)\theta \\&= -\frac{acA\omega _b}{I}cos(\omega _b t) + \frac{acA\omega _b}{I}cos(\omega _b (t-t_0)) - \frac{kcA\omega _b}{I}sin(\omega _b t) + \frac{akA\omega _b}{I}sin(\omega _b (t-t_0))
+	\end{split}
+\end{equation}
+
+\subsection{Total Response}
+\subsubsection{Base Excitation Parameters}
+To model real road conditions conditions a sinusoid with a amplitude of 0.015 and a frequency of 69.7 rad/s is used.
+This corresponds to 30mm peak to peak bumps that are separated by a distance of 2m while driving at 80km/hr (calculated in Appendix \ref{app:simple}).
+
+\subsubsection{Initial conditions}
+For the total response, the initial conditions of the system are all set to zero. This represents the vehicle traveling on a perfectly smooth road before encountering disturbances at time $t = 0$.
+
+\subsubsection{Solution}
+Based on the idealize parameters proposed in Section \ref{sec:paramSelection} , the natural frequency and damping ratio are found as follows: 
+\begin{equation}
+	\begin{split}
+	\omega _{nx} &= \sqrt{\frac{2k}{M}} = \sqrt{\frac{(2)(85000N/m)}{550kg}} = 17.58rad/s = 2.80Hz \\
+	\omega _{n\Theta} &= \sqrt{\frac{2a^2k}{I}} = \sqrt{\frac{(2)(1.3^2m^2)(85000N)}{550kgm^2}} = 22.86rad/s = 3.64Hz 
+	\end{split}  
+\end{equation}
+\begin{equation}
+	\begin{split}
+		c_{eqx} &= 2c = (2)(4830Ns/m) = 9660Ns/m \\ 
+		c_{rx} &= 2M\omega _n = (2)(550kg)(17.58rad/s) = 19340Ns/m \\ 
+		\zeta _x& = \frac{c_{eqx}}{c_r} = \frac{9660Ns/m}{19340Ns/m} = 0.50\\
+		c_{eq\Theta} &= 2a^2c =  2(1.3^2m^2)(4830Ns/m) = 16330Nms\\
+		c_{r\Theta} &= 2m_{eq}\omega _n = (2)(550kgm^2)(22.86rad/s) = 25150Nms\\ 
+		\zeta _\Theta &= \frac{c_{eq\Theta}}{c_r} = \frac{16330Nms}{ 25150Nms} = 0.65
+	\end{split}  
+\end{equation}
+Using the MATLAB script shown in Appendix \ref{app:simple}, values for the parameters $I$, $M$, $k$, $c$, and $a$ are substituted into Equations \ref{eq:eomx} and \ref{eq:eomtheta} and the Laplace transform is taken.
+Solving for $X(s)$ and $\Theta(s)$ and taking the inverse Laplace transform yeilds:
+\begin{equation}
+		\begin{split}
+		x(t) = &(3.12\cdot 10^{-5})\, \sin\!(69.7\, t) - 0.00201\, \cos\!(69.7\, t) + 1.0\, \mathop{\mathrm{u}}\nolimits\!(1.0\, t - 0.117)\, ((5.91\cdot 10^{-4})\, \cos\!(69.7\, t) \\
+		&- 0.00192\, \sin\!(69.7\, t)) + \mathrm{e}^{- 8.78\, t}\, (0.00201\, \cos\!(15.2\, t) + 0.00102\, \sin\!(15.2\, t) \\
+		&- \mathop{\mathrm{u}}\nolimits\!(1.0\, t - 0.117)\, (0.00396\, \cos\!(15.2\, t) - 0.00489\, \sin\!(15.2\, t)))
+		\end{split}
+\end{equation}
+
+\begin{equation}
+\begin{split}
+	\theta (t) =& 0.00242\, \cos\!(69.7\, t) + (3.31\cdot 10^{-4})\, \mathop{\mathrm{u}}\nolimits\!(1.0\, t - 0.117)\, \cos\!(69.7\, t) - 0.00242\, \mathrm{e}^{- 14.8\, t}\, \cos\!(9.42\, t) \\
+	 &- (5.64\cdot 10^{-4})\, \mathrm{e}^{- 14.8\, t}\, \sin\!(9.42\, t) - 1.0\, \sin\!(69.7\, t)\, (0.00244\, \mathop{\mathrm{u}}\nolimits\!(1.0\, t - 0.117) + 4.39\cdot 10^{-4}) \\
+	 &+ 0.00334\, \mathop{\mathrm{u}}\nolimits\!(1.0\, t - 0.117)\, \mathrm{e}^{- 14.8\, t}\, \cos\!(9.42\, t) + 0.0137\, \mathop{\mathrm{u}}\nolimits\!(1.0\, t - 0.117)\, \mathrm{e}^{- 14.8\, t}\, \sin\!(9.42\, t)
+\end{split}
+\end{equation}
+
+\subsubsection{Plotting Total Response}
+Figure \ref{fig:totalResp} shows the total response of the system subjected to the simulated road input. 
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width=.8\textwidth]{./matlab/totalResp.png}
+	\caption{Total Response of the System}
+	\label{fig:totalResp}
+\end{figure}
+
+\subsection{Alternate Formulation}
+If the center of gravity is not centered between the front and rear wheel axes, the problem increases significantly in complexity and an alternate approach to solving the problem is required.
+By taking the unilateral Laplace transform of the above two equations with all initial conditions set to zero, the following $s$ domain equations are obtained:
 \begin{equation}
@@
 
-\section{Analysis}
-\subsection{Natural Frequency}
-Using the MATLAB script show in Appendix \ref{app:code}, the transfer function $H(s)$ is solved for symbolically and the parameters for $I$, $M$, $k_1$, $k_2$, $a$ and $b$ are substituted into the equation while the damping coefficients $c_1$ and $c_2$ are set to zero.
+Using the MATLAB script shown in Appendix \ref{app:alternative}, the transfer function $H(s)$ is solved for symbolically as described above and the parameters for $I$, $M$, $k_1$, $k_2$, $c_1$, $c_2$, $a$ and $b$ are substituted into the equation.
 The inverse laplace transform is then taken to obtain the impulse response as shown in Figure \ref{fig:impResp}.
@@
 	\centering
-	\label{fig:impResp}
 	\includegraphics[width=.8\textwidth]{./matlab/impResp.png}
 	\caption{Undamped Impulse Response}
+	\label{fig:impResp}
 \end{figure}
@@
 A sampling frequency of 200Hz and and a window of 30s was used for the FFT.
-From figure \ref{fig:fftImpResp}, the largest harmonics of the system occur at Hz for translational displacement and Hz for angular displacement. 
 \begin{figure}[h!]
 	\centering
-	\label{fig:impResp}
 	\includegraphics[width=.8\textwidth]{./matlab/fftImpResp.png}
 	\caption{FFT of the Undamped Impulse Response}
+	\label{fig:fftImpResp}
 \end{figure}
+From Figure \ref{fig:fftImpResp}, the largest harmonics of the system occur at approximately 2Hz for translational displacement and 4Hz for angular displacement.
+Note that these frequencies align closely with the natural frequencies that were determined analytically in the simplified case.
 
-\subsection{Critical Damping}
-
-\subsection{Force Response}
-\subsection{Worst Case Total Response}
+% Source
+\pagebreak
+\bibliography{bib}{}
+\bibliographystyle{plain}
 
@@
 \appendix
-\section{MATLAB Source Code}
-\label{app:code}
-\lstinputlisting[breaklines=true]{./matlab/createTransferFunctions.m}
+\section{MATLAB Source Code for Simplified Model}
+\label{app:simple}
+\lstinputlisting[breaklines=true]{./matlab/alternativeFormulation.m}
+\pagebreak
+\section{MATLAB Source Code for Alternative Formulation}
+\label{app:alternative}
+\lstinputlisting[breaklines=true]{./matlab/simplifiedFormulation.m}
 

### 2017-04-04 SpringRateProject/matlab/alternativeFormulation.m

@@ -1,0 +1,152 @@
+% ---------------------------------------------------------------------- %
+%
+% Suspension Spring Rates Project - alternativeFormulation.m
+% Function to setup transfer functions for a half car model to determine the
+% reponse of MSXII to disturbances in the road conditions
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2017
+%
+%
+% Notes:
+%   
+%   - 
+% ---------------------------------------------------------------------- %
+
+load_system('halfCarModel');
+close all;
+clear all
+
+% ----------------------------- Constants ------------------------------ %
+
+c1_ = 4830;     % Front damping coefficient (Ns/m)
+c2_ = 4830;     % Rear damping coefficient (Ns/m) 
+k1_ = 650*175.1*cosd(40);    % Front sring rate (N/m) 
+k2_ = 550*175.1*cosd(30);    % Rear sring rate (N/m)
+m_ = 550;       % Mass of the vehicle (kg)
+I_ = 550;       % Moment of interial about global x axis (kg*m^2)
+a_ = 1.32;       % Distance from COG to front tire (m)
+b_ = 1.28;       % Distance from COG to rear tire (m)
+
+% ---------------------- Forcing Function Parameters ------------------- %
+
+amplitude = 0.015;              % Amplitude of bumps on road surface
+bumpSparation = 2;              % Distance between peaks on the road surface (m)
+speed = 22.2;                   % Cruising speed (m/s)
+wb = 2*pi*speed/bumpSparation;  % Frequency of base excitation (rad/sec)
+timeDelay = (a_+b_)/speed;      % delay between input striking front and
+                                % rear wheels (s)
+                                
+% Update simulink parameters
+set_param('halfCarModel/DelayA','DelayTime',mat2str(timeDelay));
+set_param('halfCarModel/DelayB','DelayTime',mat2str(timeDelay));
+                              
+% ----------------- Create Symbolic Transfer Functions ----------------- %
+
[…]

### 2017-04-04 SpringRateProject/matlab/createTransferFunctions.m

@@ -19,14 +19,14 @@ close all;
 
 % ----------------------------- Constants ------------------------------ %
 
-c1_ = 0;     % Front damping coefficient (Ns/m)
-c2_ = 0;     % Rear damping coefficient (Ns/m) 
-k1_ = 50000;    % Front sring rate (N/m) 
-k2_ = 60000;    % Rear sring rate (N/m)
-m_ = 500;       % Mass of the vehicle (kg)
+c1_ = 9670;     % Front damping coefficient (Ns/m)
+c2_ = 9670;     % Rear damping coefficient (Ns/m) 
+k1_ = 650*175.1;    % Front sring rate (N/m) 
+k2_ = 550*175.1;    % Rear sring rate (N/m)
+m_ = 550;       % Mass of the vehicle (kg)
 I_ = 550;       % Moment of interial about global x axis (kg*m^2)
-a_ = 1.2;       % Distance from COG to front tire (m)
-b_ = 1.4;       % Distance from COG to rear tire (m)
+a_ = 1.3;       % Distance from COG to front tire (m)
+b_ = 1.3;       % Distance from COG to rear tire (m)
 
 % ---------------------- Forcing Function Parameters ------------------- %
 
@@ -45,12 +45,6 @@ set_param('halfCarModel/DelayB','DelayTime',mat2str(timeDelay));
 % ----------------- Create Symbolic Transfer Functions ----------------- %
 
 syms t s c1 c2 k1 k2 m I a b t0 A w0 c k 
-
-c1 = c;
-c2 = c;
-k1 = k;
-k2 = k;
-b = a;
 
 QA1 = m*s^2 + (c1+c2)*s + (k1+k2); % X terms of ODE (A) 
 QA2 = (b*c2-a*c1)*s + (b*k2-a*k1); % Theta terms of ODE (A)

### 2017-04-04 SpringRateProject/matlab/simplifiedFormulation.m

@@ -1,0 +1,91 @@
+% ---------------------------------------------------------------------- %
+%
+% Suspension Spring Rates Project - simplifiedFormulation.m
+% Function to setup simplified transfer functions for a half car model to 
+% determine the reponse of MSXII to disturbances in the road conditions
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2017
+%
+%
+% Notes:
+%   
+%   - 
+% ---------------------------------------------------------------------- %
+
+close all;
+clear all;
+clc;
+
+% ----------------------------- Constants ------------------------------ %
+
+c_ = 4830;     % Front damping coefficient (Ns/m)
+k_ = 85000;    % Front sring rate (N/m) 
+M_ = 550;      % Mass of the vehicle (kg)
+I_ = 550;      % Moment of interial about global x axis (kg*m^2)
+a_ = 1.3;      % Distance from COG to front tire and rear tire (m)
+zx_ = 0.50;     % Damping ratio of translational  
+wnx_ = 17.58;   % Natural frequency of pitch
+zth_ = 0.65;    % Damping ratio of pitch
+wnth_ = 22.86;  % Natural frequency of pitch
+
+
+% ---------------------- Forcing Function Parameters ------------------- %
+
+A_ = 0.015;                     % Amplitude of bumps on road surface
+bumpSparation = 2;              % Distance between peaks on the road surface (m)
+speed = 22.2;                   % Cruising speed (m/s)
+wb_ = 2*pi*speed/bumpSparation; % Frequency of base excitation (rad/sec)
+t0_ = (2*a_)/speed;            % delay between input striking front and
+                                % rear wheels (s)
+
+% ----------------- Create Symbolic Transfer Functions ----------------- %                                
+                                
+syms t s M I a t0 A wb c k zx wnx zth wnth
+
+QA = s^2 + 2*c/M*s + wnx^2; % ODE A 
+QB = s^2 + 2*a^2*c/I*s + wnx^2; % ODE B
+
+phaseShift = exp(-s*t0);
+
[…]

### 2017-04-06 SpringRateProject/MSXII Supsension Parameters.tex
Revisions to wording in report

@@
 \subsection{Vehicle Design}
-Midnight Sun Twelve (MSXII) is a cruiser class solar electric vehicle being designed with the goal of competing in the 2018 American Solar Challenge (ASC 2018) and the 2019 World Solar Challenge (WSC 2019).
-Cruiser class vehicles are must be multi-occupant and are designed with the intent of being more practical than a typical solar car.
-Because of the unique requirements of this class, the spring rates and target damping coefficients on MSXII's suspension must be selected to optimize for efficient while still keeping driver comfort in mind.
-This report proposes and analytical technique for determining the response of the suspension system to a variety of road conditions.
+Midnight Sun Twelve (MSXII) is a cruiser class, solar electric vehicle being designed with the goal of competing in the 2018 American Solar Challenge (ASC 2018) and the 2019 World Solar Challenge (WSC 2019).
+By definition, cruiser class solar vehicles must be multi-occupant and are designed with the intent of being more practical than a typical, challenger class solar car.
+Because of the unique requirements of this class, the spring rates and target damping coefficients on MSXII's suspension must be selected to optimize for efficiency while still keeping driver comfort in mind.
+This report proposes an analytical technique for determining the response of a suspension system to any given road conditions.
 
 \subsection{Important Vehicle Parameters}
-Table \ref{tab:params} lists predicted vehicle parameters that are of importance to the design and analysis of the suspension. The mass and moment of inertia was computed by CAD software. 
+Table \ref{tab:params} lists vehicle parameters that are of importance to the design and analysis of MSXII's suspension. The mass and moment of inertia was computed by CAD software. 
 \begin{table}[htbp]
@@
 MSXII's front suspension comprises of a double wishbone linkage with an outboard coilover while the rear suspension comprises of independent trailing arms allowing for zero scrub and thus less rolling resistance.
-Figure \ref{fig:solidworksScreenshots} shows the screenshots of the front and rears suspension.
+Figure \ref{fig:solidworksScreenshots} shows the screenshots of the front and rear suspension.
 \begin{figure}[htbp]
@@
 		\centering
-        \includegraphics[height=4cm]{./LaTex/frontIso.PNG}
+        \includegraphics[height=3.5cm]{./LaTex/frontIso.PNG}
     \end{subfigure}
@@
 		\centering
-        \includegraphics[height=4cm]{./LaTex/rearIso.PNG}
+        \includegraphics[height=3.5cm]{./LaTex/rearIso.PNG}
         \label{fig:c8}
@@
 		\centering
-        \includegraphics[height=4cm]{./LaTex/rearRight.PNG}
+        \includegraphics[height=3.5cm]{./LaTex/rearRight.PNG}
         \label{fig:c9}
@@
 \label{sec:paramSelection}
-The spring rates are to be calculated based on the travel of the wheel and the maximum loading conditions.
-Assuming a smooth driving surface, the largest normal force acting on the tire from the road will occur while braking in a turn.
-To approximate this loading condition the a superposition of two static half car models is used; one for cornering and one for hard braking.
+The spring rates are calculated based on the travel of the wheel and the maximum vertical loading of the wheel due to normal force from the road.
+Assuming a smooth driving surface, the largest normal force acting on the tire will occur while braking in a turn.
+To approximate this loading condition, the a superposition of two static half car models is used; one for cornering and one for hard braking.
 \subsubsection{Hard Braking}\begin{figure}[h!]
@@
 Figure \ref{fig:barkingFBD} shows a free body diagram of a half car model undergoing braking.
-MSXII will only have brake callipers on the front wheels however regenerative braking from the hub motors will also provide a braking force at the rear contact patch.
-Assuming a generous coefficient of static friction, $\mu _s$, of 1.0 and the extreme case where the both the font and rear tires are about to begin sliding the combined braking force can be expressed as:
+MSXII will only have brake callipers on the front wheels however regenerative braking from the hub motors will also provide a braking force at the rear contact patches.
+Assuming a generous coefficient of static friction, $\mu _s$, of 1.0 and the extreme case where the both the font and rear tires are about to begin sliding, the combined braking force can be expressed as:
 \begin{equation}
@@
 \end{equation}
-Summing the moments about the centre of gravity and the vertical forces results in the following equations: 
+From Figure \ref{fig:barkingFBD}, summing the moments about the centre of gravity and the vertical forces results in the following equations: 
 \begin{equation}
@@
 \end{equation}
-Note that the 2.66kN above is for both front wheels. Therefore there is approximately \textbf{1.33kN} of normal force on each front tire during hard braking.
+Note that 2.66kN is for both front wheels. Therefore there is approximately \textbf{1.33kN} of normal force on each front tire during hard braking.
 
@@
 The ASC 2018 regulations stipulate that a vehicle must be able to navigate a figure of eight as shown in Figure \ref{fig:dynamicScrutineering} in less than 18 seconds \cite{ASC2018regs}.
-Since the average arc length of the figure of eight is $34\pi$, the net cornering force required to navigate the figure of eight can be found as follows:
+Since the average arc length of the figure of eight is $34\pi m$, the net cornering force required to navigate the figure of eight can be found as follows:
 \begin{equation}
@@
 \end{figure}
-For the half car model shown if Figure \ref{fig:steerFBD} umming the moments about the centre of gravity and the vertical forces results in the following equations: 
+
+\noindent For the half car model shown if Figure \ref{fig:steerFBD}, summing the moments about the centre of gravity and the vertical forces results in the following equations: 
 \begin{equation}
@@
 \end{equation}
-Solving for the front normal force: 
+Solving for the outside wheel normal force: 
 \begin{equation}
@@
 The maximum travel of the suspension for MSXII in compression has been set to 45mm.
-It is desired that under the combined of braking and cornering, the suspension compresses to a maximum of 70\% of the travel.
+It is desired that under the combined loading condition of braking and cornering, the suspension compresses to a maximum of 70\% of the travel.
 As such, the desired spring rate is found to be:
@@
 Note that the above rate assumes a vertical spring.
-The vertical spring rate can be mapped to an equivalent spring rate by dividing by the cosine of the angle that the coilover makes with the vertical direction:
+The vertical spring rate can be mapped to an equivalent angled spring rate by dividing by the cosine of the angle that the coilover makes with the vertical direction:
 \begin{equation}
@@
 \subsubsection{Damping Coefficient Selection}
-Ohlins has published curves for their TTX25 MkII's however real world damping rates are non constant.
+Ohlins has published damper curves for their TTX25 MkII's however real world damping rates are non constant.
 For this report, the damping coefficient will be idealized as constant and will be selected such that the damping ratio, $\zeta$, is 0.5.
-This is typical of performance consumer vehicles \cite{dampers}.
+This is typical of high performance consumer vehicles \cite{dampers}.
 Given an equivalent vertical spring rate or 85.0kN/m, a damping ratio of 0.5 maps to a vertical damping coefficient of \textbf(4.83kNs/m).
@@
 \begin{align*}
-	\theta & \text{ is the pitch angle about positive x axis}\\        
+	\theta & \text{ is the pitch angle about the axis normal to the vehicle's symmetry plane}\\        
 	x_0 & \text{ is the displacement of the center of gravity in the vertical axis} \\
@@
 	M & \text{ is the mass of the vehicle} \\
-	I & \text{ is the moment of inertia normal to the half car plane at the center of gravity} \\
+	I & \text{ is the moment of inertia normal to the vehicle’s symmetry plane at the center of gravity} \\
 	k1,k2,c1 \text{ \& } c_2 & \text{ are the spring and damping coefficients} \\
@@
 The above two equations are considerably complex and are not particularly conducive of solving analytically by hand.
-However, the fact that the vehicle has the center of gravity approximately centered along the wheel offers an avenue for simplification.
+However, the fact that the vehicle has the center of gravity approximately centered along the wheel such that $a$ and $b$ are equal, offers an avenue for simplification.
 Additionally if the front and rear spring and damping coefficients are assumed to be the same ($k$ and $c$ respectively), the problem's complexity is drastically reduced as shown in the following two equations:
@@
 
-\subsubsection{Final Equations fo Motion}
+\subsubsection{Final Equations of Motion}
 The road displacement can be approximated by a sinusoid with and amplitude $A$ and a base excitation frequency $\omega _b$.
-Since rear tires experience the parts to the road as the front $y_2$ can be expressed as a phase shifted version of $y_1$.
+Since the rear wheels experience the same displacements as the front $y_2$ can be expressed as a phase shifted version of $y_1$.
 Combining this with the natural frequency and damping ratio results in the following equations of motion:
@@
 \subsubsection{Base Excitation Parameters}
-To model real road conditions conditions a sinusoid with a amplitude of 0.015 and a frequency of 69.7 rad/s is used.
+To model real road conditions, a sinusoid with a amplitude of 0.015 and a frequency of 69.7 rad/s is used.
 This corresponds to 30mm peak to peak bumps that are separated by a distance of 2m while driving at 80km/hr (calculated in Appendix \ref{app:simple}).
@@
 \end{equation}
-Since the system is linear linear, the above matrix equation can be solved in two parts to obtain the transfer functions for $X_0$ and $\Theta$:
+Since the system is linear, the above matrix equation can be solved in two parts to obtain the transfer functions for $X_0$ and $\Theta$:
 \begin{equation}

### 2017-04-08 SpringRateProject/MSXII Supsension Parameters.tex
Fucked up the units
Fixed mistake made when finding applied loads on the vehicle (didn’t
convert kN to N in calculations). Updated report and matlab script
accordingly. Reorganized file structure

@@
 \begin{split}
-	aF_{Nf} &= hF_{braking} + b(Mg - F_{Nr})\\
-	F_{Nf} &= \frac{hF_{braking} + bMg}{a+b} = \frac{(0.52m)(5.40kN)+(1.28m)(550kg)\left(9.81\frac{m}{s^2}\right)}{(1.32m)+(1.28m)} = 2.66kN
+	aF_{Nf} &= hF_{braking} + b(Mg - F_{Nf})\\
+	F_{Nf} &= \frac{hF_{braking} + bMg}{a+b} = \frac{(0.52m)(5.40kN)+(1.28m)(550kg)\left(9.81\frac{m}{s^2}\right)}{(1.32m)+(1.28m)} = 3.74kN
 \end{split}
 \end{equation}
-Note that 2.66kN is for both front wheels. Therefore there is approximately \textbf{1.33kN} of normal force on each front tire during hard braking.
+Note that 3.74kN is for both front wheels. Therefore there is approximately \textbf{1.87kN} of normal force on each front tire during hard braking.
 
@@
 \begin{equation}
-	\frac{track}{2}F_{Nf} = hF_{steer} + \frac{track}{2}F_{Nr}
+	\frac{track}{2}F_{No} = hF_{steer} + \frac{track}{2}F_{Ni}
 \end{equation}
@@
 	\frac{track}{2}F_{No} &= hF_{steer} + \frac{track}{2}(Mg - F_{Ni})\\
-	F_{No} &= \frac{hF_{steer} + \frac{track}{2}Mg}{track} = \frac{(0.52m)(2.28kN)+(0.80m)(550kg)\left(9.81\frac{m}{s^2}\right)}{1.60m} = 2.70kN
+	F_{No} &= \frac{hF_{steer} + \frac{track}{2}Mg}{track} = \frac{(0.52m)(2.28kN)+(0.80m)(550kg)\left(9.81\frac{m}{s^2}\right)}{1.60m} = 3.44kN
 \end{split}
 \end{equation}
-Note that the 2.70kN above is for both front and rear outside wheels.
-Assuming the force is evenly split between the front and rear, there is approximately \textbf{1.35kN} of normal force on each outside tire during maximum cornering.
+Note that the 3.44kN above is for both front and rear outside wheels.
+Assuming the force is evenly split between the front and rear, there is approximately \textbf{1.72kN} of normal force on each outside tire during maximum cornering.
 \subsubsection{Spring Rate Selection}
-Adding the maximum normal forces from cornering and braking results in 2.68kN.
+Adding the maximum normal forces from cornering and braking results in 3.59kN.
 The maximum travel of the suspension for MSXII in compression has been set to 45mm.
-It is desired that under the combined loading condition of braking and cornering, the suspension compresses to a maximum of 70\% of the travel.
+It is desired that under the combined loading condition of braking and cornering, the suspension compresses to a maximum of 90\% of the travel.
 As such, the desired spring rate is found to be:
 \begin{equation}
-	k = \frac{force}{displacement} = \frac{2.68N}{(70\%)(0.045m)} = 85.0kN/m
+	k = \frac{force}{displacement} = \frac{3.59kN}{(70\%)(0.045m)} = 88.6kN/m
 \end{equation}
@@
 	\begin{split}
-		k_{front} &= \frac{85.0}{cos(35)} = 110.6kN \\
-		k_{rear} &= {85.0}{cos(30)} = 98.1kN
+		k_{front} &= \frac{88.6}{cos(35)} = 108.2kN \\
+		k_{rear} &= {88.6}{cos(30)} = 102.3kN
 	\end{split}
@@
 These coilovers can be configured to have spring rates from 150 lb/in to 650 lb/in in increments of 50 lb/in.
-1 lb/in is approximately 175 N/m therefore 550 lb/in coils should be selected for the front and 650 lb/in coils for the rear.
+1 lb/in is approximately 175 N/m therefore 600 lb/in coils should be selected for the front and 600 lb/in coils for the rear.
 
@@
 This is typical of high performance consumer vehicles \cite{dampers}.
-Given an equivalent vertical spring rate or 85.0kN/m, a damping ratio of 0.5 maps to a vertical damping coefficient of \textbf(4.83kNs/m).
+Given an equivalent vertical spring rate or 88.6/m, a damping ratio of 0.5 maps to a vertical damping coefficient of \textbf(4.83kNs/m).
 
@@
 	\begin{split}
-	\omega _{nx} &= \sqrt{\frac{2k}{M}} = \sqrt{\frac{(2)(85000N/m)}{550kg}} = 17.58rad/s = 2.80Hz \\
-	\omega _{n\Theta} &= \sqrt{\frac{2a^2k}{I}} = \sqrt{\frac{(2)(1.3^2m^2)(85000N)}{550kgm^2}} = 22.86rad/s = 3.64Hz 
+	\omega _{nx} &= \sqrt{\frac{2k}{M}} = \sqrt{\frac{(2)(88600N/m)}{550kg}} = 17.95rad/s = 2.86Hz \\
+	\omega _{n\Theta} &= \sqrt{\frac{2a^2k}{I}} = \sqrt{\frac{(2)(1.3^2m^2)(88600N)}{550kgm^2}} = 23.33rad/s = 3.71Hz 
 	\end{split}  
@@
 		c_{eqx} &= 2c = (2)(4830Ns/m) = 9660Ns/m \\ 
-		c_{rx} &= 2M\omega _n = (2)(550kg)(17.58rad/s) = 19340Ns/m \\ 
-		\zeta _x& = \frac{c_{eqx}}{c_r} = \frac{9660Ns/m}{19340Ns/m} = 0.50\\
+		c_{rx} &= 2M\omega _n = (2)(550kg)(17.75rad/s) = 19530Ns/m \\ 
+		\zeta _x& = \frac{c_{eqx}}{c_r} = \frac{9660Ns/m}{19530Ns/m} = 0.49\\
 		c_{eq\Theta} &= 2a^2c =  2(1.3^2m^2)(4830Ns/m) = 16330Nms\\
-		c_{r\Theta} &= 2m_{eq}\omega _n = (2)(550kgm^2)(22.86rad/s) = 25150Nms\\ 
-		\zeta _\Theta &= \frac{c_{eq\Theta}}{c_r} = \frac{16330Nms}{ 25150Nms} = 0.65
+		c_{r\Theta} &= 2m_{eq}\omega _n = (2)(550kgm^2)(23.33rad/s) = 25660Nms\\ 
+		\zeta _\Theta &= \frac{c_{eq\Theta}}{c_r} = \frac{16330Nms}{ 25660Nms} = 0.64
 	\end{split}  

### 2017-04-08 SpringRateProject/matlab/alternativeFormulation.m

@@ -22,8 +22,8 @@ clear all
 
 c1_ = 4830;     % Front damping coefficient (Ns/m)
 c2_ = 4830;     % Rear damping coefficient (Ns/m) 
-k1_ = 650*175.1*cosd(40);    % Front sring rate (N/m) 
-k2_ = 550*175.1*cosd(30);    % Rear sring rate (N/m)
+k1_ = 600*175.1*cosd(35);    % Front sring rate (N/m) 
+k2_ = 500*175.1*cosd(30);    % Rear sring rate (N/m)
 m_ = 550;       % Mass of the vehicle (kg)
 I_ = 550;       % Moment of interial about global x axis (kg*m^2)
 a_ = 1.32;       % Distance from COG to front tire (m)
@@ -122,6 +122,9 @@ xlabel('Frequency (Hz)');
 title('FFT of Impusle Response');
 xlim([0 10]);
 saveas(gcf,'fftImpResp','png');
+
+% --------------------------------- Forces ----------------------------- %
+
 
 % ------------------------------- Simulink ----------------------------- %
 

### 2017-04-08 SpringRateProject/matlab/simplifiedFormulation.m

@@ -21,14 +21,14 @@ clc;
 % ----------------------------- Constants ------------------------------ %
 
 c_ = 4830;     % Front damping coefficient (Ns/m)
-k_ = 85000;    % Front sring rate (N/m) 
+k_ = 88600;    % Front sring rate (N/m) 
 M_ = 550;      % Mass of the vehicle (kg)
 I_ = 550;      % Moment of interial about global x axis (kg*m^2)
 a_ = 1.3;      % Distance from COG to front tire and rear tire (m)
-zx_ = 0.50;     % Damping ratio of translational  
-wnx_ = 17.58;   % Natural frequency of pitch
-zth_ = 0.65;    % Damping ratio of pitch
-wnth_ = 22.86;  % Natural frequency of pitch
+zx_ = 0.49;     % Damping ratio of translational  
+wnx_ = 17.95;   % Natural frequency of pitch
+zth_ = 0.64;    % Damping ratio of pitch
+wnth_ = 23.33;  % Natural frequency of pitch
 
 
 % ---------------------- Forcing Function Parameters ------------------- %

### 2017-05-21 SpringRateProject/MSXII Supsension Parameters.tex
Updated loading condition
Corrected mistake resulting from counting gravity twice when doing the
superposition of corning and braking conditions

@@
 Assuming the force is evenly split between the front and rear, there is approximately \textbf{1.72kN} of normal force on each outside tire during maximum cornering.
+\subsubsection{Superposition}
+Both the braking and cornering conditions subject the front outside wheel to a normal force greater than the nominal normal force seen when the car is at rest.
+The change in normal force for each loading condition can be expressed as follows:
+\begin{equation}
+	\begin{split}
+		\Delta F_{N_{braking}} &= F_{Nf_{braking}} - \frac{Mg}{4} =  1.87kN - 1.35kN = 0.52kN\\
+		\Delta F_{N_{cornering}} &=  F_{No_{cornering}} - \frac{Mg}{4} =  1.72kN - 1.35kN = 0.37kN
+	\end{split}
+\end{equation}
+To approximate the worst case loading condition, a superposition of the changes in normal force and the nominal normal force is applied. 
+\begin{equation}
+	F_{N_{max}} = \Delta F_{N_{braking}} + \Delta F_{N_{cornering}} + F_{N_{nominal}} = 0.52kN + 0.37kN + 1.35kN = 2.24kN
+\end{equation}
+
 \subsubsection{Spring Rate Selection}
-Adding the maximum normal forces from cornering and braking results in 3.59kN.
 The maximum travel of the suspension for MSXII in compression has been set to 45mm.
-It is desired that under the combined loading condition of braking and cornering, the suspension compresses to a maximum of 90\% of the travel.
+It is desired that under the combined loading condition of braking and cornering, the suspension compresses to a maximum of 55\% of the travel.
 As such, the desired spring rate is found to be:
 \begin{equation}
-	k = \frac{force}{displacement} = \frac{3.59kN}{(70\%)(0.045m)} = 88.6kN/m
+	k = \frac{force}{displacement} = \frac{2.24kN}{(55\%)(0.045m)} = 88.6kN/m
 \end{equation}
@@
 	\begin{split}
-		k_{front} &= \frac{88.6}{cos(35)} = 108.2kN \\
-		k_{rear} &= {88.6}{cos(30)} = 102.3kN
+		k_{front} &= \frac{88.6}{cos(35)} = 108.2kN/m \\
+		k_{rear} &= \frac{88.6}{cos(30)} = 102.3kN/m
 	\end{split}

### 2017-09-27 FEABoundaryConditions/FEABoundaryConditions.tex
Midsun latex template, pvdr, fea boundary conditions

--- 2017-05-21 SpringRateProject/MSXII Supsension Parameters.tex
@@
 	\vspace{3.5cm}
-	{\huge\bfseries MSXII Suspension Parameters\par}
+	{\huge\bfseries FEA Boundary Conditions\par}
 	\vspace{0.2cm}
-	\large Prepared for Project 3 of ME321 Winter 2017
+	\large MSXII
 	\vfill
@@
 \end{figure}
+\subsection{Two Loading Conditions}
+Two cases of boundary conditions are considered for FEA.
+One case occurs when the suspension is statically loaded during the extreme case of braking while cornering.
+The other case is a dynamic loading scenario where the car suddenly hits a bump at speed.
 
-\subsection{Selecting Suspension Parameters}
+It is important to note that the former case (that of braking while cornering) is specific to the front suspension.
+The equivalent worst case loading scenario for the rear would occur while accelerating during a corner.
+However, since MSXII's motors produce only ***Nm of torque while the brakes require ***Nm to slow the car at the required rate, the worst loading for the front will far exceed that of the rear.
+For simplicity, this value is treated as the worst case for both the front and the rear.
+
+\section{Static Loading}
 \label{sec:paramSelection}
-The spring rates are calculated based on the travel of the wheel and the maximum vertical loading of the wheel due to normal force from the road.
 Assuming a smooth driving surface, the largest normal force acting on the tire will occur while braking in a turn.
@@
 
-\subsubsection{Spring Rate Selection}
-The maximum travel of the suspension for MSXII in compression has been set to 45mm.
-It is desired that under the combined loading condition of braking and cornering, the suspension compresses to a maximum of 55\% of the travel.
-As such, the desired spring rate is found to be:
-\begin{equation}
-	k = \frac{force}{displacement} = \frac{2.24kN}{(55\%)(0.045m)} = 88.6kN/m
-\end{equation}
-Note that the above rate assumes a vertical spring.
-The vertical spring rate can be mapped to an equivalent angled spring rate by dividing by the cosine of the angle that the coilover makes with the vertical direction:
-\begin{equation}
-	\begin{split}
-		k_{front} &= \frac{88.6}{cos(35)} = 108.2kN/m \\
-		k_{rear} &= \frac{88.6}{cos(30)} = 102.3kN/m
-	\end{split}
-\end{equation}
-MSXII will be using Ohlins TTX25 MkII coilovers for both the front and rear.
-These coilovers can be configured to have spring rates from 150 lb/in to 650 lb/in in increments of 50 lb/in.
-1 lb/in is approximately 175 N/m therefore 600 lb/in coils should be selected for the front and 600 lb/in coils for the rear.
-
-\subsubsection{Damping Coefficient Selection}
-Ohlins has published damper curves for their TTX25 MkII's however real world damping rates are non constant.
-For this report, the damping coefficient will be idealized as constant and will be selected such that the damping ratio, $\zeta$, is 0.5.
-This is typical of high performance consumer vehicles \cite{dampers}.
-Given an equivalent vertical spring rate or 88.6/m, a damping ratio of 0.5 maps to a vertical damping coefficient of \textbf(4.83kNs/m).
-
-\section{Suspension Model}
-\subsection{Overview}
-MSXII's suspension is modelled using a point mass connected to two vertical, linear spring mass damper systems at the two wheels.
-This half car model assumes infinite tire stiffness and that the vehicle pitches about it's center of gravity.
-Figure \ref{fig:systemModel} describes the proposed model where:
-\begin{align*}
-	\theta & \text{ is the pitch angle about the axis normal to the vehicle's symmetry plane}\\        
-	x_0 & \text{ is the displacement of the center of gravity in the vertical axis} \\
-	x_1 \text{ \& } x_2 & \text{ are the displacements of the mounting points for the front and rear wheels respectively} \\
-	y_1 \text{ \& } y_2 & \text{ is the change in road elevation below the front  and rear wheels respectively} \\
-	M & \text{ is the mass of the vehicle} \\
-	I & \text{ is the moment of inertia normal to the vehicle’s symmetry plane at the center of gravity} \\
-	k1,k2,c1 \text{ \& } c_2 & \text{ are the spring and damping coefficients} \\
-	a \text{ \& } b & \text{ are the distances from center of gravity to the front and rear wheels respectively}
-\end{align*}
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=.8\textwidth]{./LaTex/systemModel.jpg}
-	\caption{Half car model with infinite tire stiffness}
-	\label{fig:systemModel}
-\end{figure}
-
-\subsection{Equations of Motion}
-\subsubsection{Dynamic Equilibrium}
-The spring mass damper system shown in Figure \ref{fig:systemModel} can be described by the folowing two differential equations: 
-\begin{equation}
-	M \ddot x + c_1(\dot x_1 - \dot y_1) + c_2(\dot x_2 - \dot y_2) + k_1(x_1 - y_1) + k_2(x_2 - y_2) = 0
-\end{equation}
-\begin{equation}
-	I \ddot\theta - ac_1(\dot x_1 - \dot y_1) + bc_2(\dot x_2 - \dot y_2) - ak_1(x_1 - y_1) + bk_2(x_2 - y_2) = 0
-\end{equation}
-Rearranging to move the inputs $y_1$ and $y_2$ to right hand side: 
-\begin{equation}
-	M \ddot x + c_1\dot x_1 + c_2\dot x_2 + k_1x_1 + k_2x_2 = c_1\dot y_1 + c_2\dot y_2 + k_1y_1 + k_2y_2
-\end{equation}
-\begin{equation}
-	I \ddot\theta - ac_1\dot x_1 + bc_2\dot x_2 - ak_1x_1 + bk_2x_2 = -ac_1\dot y_1 + bc_2\dot y_2 - ak_1y_1 + bk_2y_2
-\end{equation}
-By assuming small pitch angles, $\theta$, the angle between the spring mass damper system becomes approximately perpendicular and the displacements $x_1$ and $x_2$ can be approximated as follows:
-\begin{equation}
-	x_1 = x_0 - a\theta
-\end{equation}
-\begin{equation}
-	x_2 = x_0 + b\theta
-\end{equation}
-It follows that the time derivatives of the above two equations are: 
-\begin{equation}
-	\dot x_1 = x_0 - a\dot\theta
-\end{equation}
-\begin{equation}
-	\dot x_2 = x_0 + b\dot\theta
-\end{equation}
-Substituting into the original differential equations: 
-\begin{equation}
-	M \ddot x + \dot x_0(c_2+c_1) + \dot\theta(bc_2-ac_1) + x_0(k_2+k_1) + \theta(bk_2-ak_1)= c_1\dot y_1 + c_2\dot y_2 + k_1y_1 + k_2y_2
-\end{equation}
-\begin{equation}
-	I \ddot\theta + \dot x_0(bc_2-ac_1) + \dot\theta(b^2c_2+a^2c_1) + x_0(bk_2-ak_1) + \theta(b^2k_2+a^2k_1) = -ac_1\dot y_1 + bc_2\dot y_2 - ak_1y_1 + bk_2y_2
-\end{equation}
-The above two equations are considerably complex and are not particularly conducive of solving analytically by hand.
-However, the fact that the vehicle has the center of gravity approximately centered along the wheel such that $a$ and $b$ are equal, offers an avenue for simplification.
-Additionally if the front and rear spring and damping coefficients are assumed to be the same ($k$ and $c$ respectively), the problem's complexity is drastically reduced as shown in the following two equations:
-\begin{equation}
-	M \ddot x + 2c\dot x_0 + 2kx_0 = c\dot y_1 + c\dot y_2 + ky_1 + ky_2
-\end{equation}
-\begin{equation}
-	I \ddot\theta + 2a^2c\dot\theta + 2a^2k\theta = -ac\dot y_1 + bc_2\dot y_2 - ak_1y_1 + bk_2y_2
-\end{equation}
-
-\subsubsection{Natural Frequency}
-The natural frequency of the translational and rotational components of the system are found as follows: 
-\begin{equation}
-	\begin{split}
-	\omega _{nx} &= \sqrt{\frac{k_{eq}}{m_{eq}}} = \sqrt{\frac{2k}{M}}\\
-	\omega _{n\Theta} &= \sqrt{\frac{k_{eq}}{m_{eq}}}= \sqrt{\frac{2a^2k}{I}}
-	\end{split}  
-\end{equation}
-
-\subsubsection{Critical Damping}
-The critical camping coefficient for the translational and rotational components of the system are found as follows:
-\begin{equation}
-	\begin{split}
-	c_{eqx} = 2c, c_{rx} = 2m_{eq}\omega _n = \sqrt{8kM}, \zeta _x = \frac{c_{eqx}}{c_r}\\
-	c_{eq\Theta} = 2a^2c, c_{r\Theta} = 2m_{eq}\omega _n = \sqrt{8a^2kI}, \zeta _\Theta = \frac{c_{eq\Theta}}{c_r}
-	\end{split}  
-\end{equation}
-
-\subsubsection{Final Equations of Motion}
-The road displacement can be approximated by a sinusoid with and amplitude $A$ and a base excitation frequency $\omega _b$.
-Since the rear wheels experience the same displacements as the front $y_2$ can be expressed as a phase shifted version of $y_1$.
-Combining this with the natural frequency and damping ratio results in the following equations of motion:
-\begin{equation}
-	\label{eq:eomx}
-	\ddot x + (2\zeta _x M \omega _{nx}) \dot x_0 + (\omega _{nx}^2)x_0 = \frac{cA\omega _b}{M}cos(\omega _b t) + \frac{cA\omega _b}{M}cos(\omega _b (t-t_0))  + \frac{kA}{M}sin(\omega _b t) + \frac{kA}{M}sin(\omega _b (t-t_0))
-\end{equation}
-\begin{equation}
-	\label{eq:eomtheta}
-	\begin{split}
-		\ddot\theta + &(2\zeta _\Theta  I \omega _{n\Theta}) \dot\theta + (\omega _{n\Theta}^2)\theta \\&= -\frac{acA\omega _b}{I}cos(\omega _b t) + \frac{acA\omega _b}{I}cos(\omega _b (t-t_0)) - \frac{kcA\omega _b}{I}sin(\omega _b t) + \frac{akA\omega _b}{I}sin(\omega _b (t-t_0))
-	\end{split}
-\end{equation}
-
-\subsection{Total Response}
-\subsubsection{Base Excitation Parameters}
-To model real road conditions, a sinusoid with a amplitude of 0.015 and a frequency of 69.7 rad/s is used.
-This corresponds to 30mm peak to peak bumps that are separated by a distance of 2m while driving at 80km/hr (calculated in Appendix \ref{app:simple}).
-
-\subsubsection{Initial conditions}
-For the total response, the initial conditions of the system are all set to zero. This represents the vehicle traveling on a perfectly smooth road before encountering disturbances at time $t = 0$.
-
-\subsubsection{Solution}
-Based on the idealize parameters proposed in Section \ref{sec:paramSelection} , the natural frequency and damping ratio are found as follows: 
-\begin{equation}
-	\begin{split}
-	\omega _{nx} &= \sqrt{\frac{2k}{M}} = \sqrt{\frac{(2)(88600N/m)}{550kg}} = 17.95rad/s = 2.86Hz \\
-	\omega _{n\Theta} &= \sqrt{\frac{2a^2k}{I}} = \sqrt{\frac{(2)(1.3^2m^2)(88600N)}{550kgm^2}} = 23.33rad/s = 3.71Hz 
-	\end{split}  
-\end{equation}
-\begin{equation}
-	\begin{split}
-		c_{eqx} &= 2c = (2)(4830Ns/m) = 9660Ns/m \\ 
-		c_{rx} &= 2M\omega _n = (2)(550kg)(17.75rad/s) = 19530Ns/m \\ 
-		\zeta _x& = \frac{c_{eqx}}{c_r} = \frac{9660Ns/m}{19530Ns/m} = 0.49\\
-		c_{eq\Theta} &= 2a^2c =  2(1.3^2m^2)(4830Ns/m) = 16330Nms\\
-		c_{r\Theta} &= 2m_{eq}\omega _n = (2)(550kgm^2)(23.33rad/s) = 25660Nms\\ 
-		\zeta _\Theta &= \frac{c_{eq\Theta}}{c_r} = \frac{16330Nms}{ 25660Nms} = 0.64
-	\end{split}  
-\end{equation}
-Using the MATLAB script shown in Appendix \ref{app:simple}, values for the parameters $I$, $M$, $k$, $c$, and $a$ are substituted into Equations \ref{eq:eomx} and \ref{eq:eomtheta} and the Laplace transform is taken.
-Solving for $X(s)$ and $\Theta(s)$ and taking the inverse Laplace transform yeilds:
-\begin{equation}
-		\begin{split}
-		x(t) = &(3.12\cdot 10^{-5})\, \sin\!(69.7\, t) - 0.00201\, \cos\!(69.7\, t) + 1.0\, \mathop{\mathrm{u}}\nolimits\!(1.0\, t - 0.117)\, ((5.91\cdot 10^{-4})\, \cos\!(69.7\, t) \\
-		&- 0.00192\, \sin\!(69.7\, t)) + \mathrm{e}^{- 8.78\, t}\, (0.00201\, \cos\!(15.2\, t) + 0.00102\, \sin\!(15.2\, t) \\
-		&- \mathop{\mathrm{u}}\nolimits\!(1.0\, t - 0.117)\, (0.00396\, \cos\!(15.2\, t) - 0.00489\, \sin\!(15.2\, t)))
-		\end{split}
-\end{equation}
-
-\begin{equation}
-\begin{split}
-	\theta (t) =& 0.00242\, \cos\!(69.7\, t) + (3.31\cdot 10^{-4})\, \mathop{\mathrm{u}}\nolimits\!(1.0\, t - 0.117)\, \cos\!(69.7\, t) - 0.00242\, \mathrm{e}^{- 14.8\, t}\, \cos\!(9.42\, t) \\
-	 &- (5.64\cdot 10^{-4})\, \mathrm{e}^{- 14.8\, t}\, \sin\!(9.42\, t) - 1.0\, \sin\!(69.7\, t)\, (0.00244\, \mathop{\mathrm{u}}\nolimits\!(1.0\, t - 0.117) + 4.39\cdot 10^{-4}) \\
-	 &+ 0.00334\, \mathop{\mathrm{u}}\nolimits\!(1.0\, t - 0.117)\, \mathrm{e}^{- 14.8\, t}\, \cos\!(9.42\, t) + 0.0137\, \mathop{\mathrm{u}}\nolimits\!(1.0\, t - 0.117)\, \mathrm{e}^{- 14.8\, t}\, \sin\!(9.42\, t)
-\end{split}
-\end{equation}
-
-\subsubsection{Plotting Total Response}
-Figure \ref{fig:totalResp} shows the total response of the system subjected to the simulated road input. 
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=.8\textwidth]{./matlab/totalResp.png}
-	\caption{Total Response of the System}
-	\label{fig:totalResp}
-\end{figure}
-
-\subsection{Alternate Formulation}
-If the center of gravity is not centered between the front and rear wheel axes, the problem increases significantly in complexity and an alternate approach to solving the problem is required.
-By taking the unilateral Laplace transform of the above two equations with all initial conditions set to zero, the following $s$ domain equations are obtained:
-\begin{equation}
-	\left[Ms^2+(c_2+c_1)s+(k_2+k_1)\right]X_0 + \left[(bc_2-ac_1)s+(bk_2+ak_1)\right]\Theta = c_1sY_1 + c_2sY_2 + k_1Y_1 + k_2Y_2
-\end{equation}
-\begin{equation}
-	\left[(bc_2-ac_1)s+(bk_2-ak_1)\right]X_0 + \left[Is^2+ (b^2c_2+a^2c_1)s+(b^2k_2+a^2k_1)\right]\Theta = c_1sY_1 + c_2sY_2 + k_1Y_1 + k_2Y_2
-\end{equation}
-The above two equations can be expressed in matrix form as follows: 
-\begin{equation}
-	\label{eq:matrix1}
-	\begin{bmatrix} 
-	Q_{Ax}(s) & Q_{A\Theta}(s) \\
-	Q_{Bx}(s) & Q_{B\Theta}(s)
-	\end{bmatrix}
-	\begin{bmatrix} 
-	X_0 \\
-	\Theta 
-	\end{bmatrix}
-	= 
-	\begin{bmatrix} 
-	P_{A1}(s) \\
-	P_{B1}(s)
-	\end{bmatrix}Y_1 +
-	\begin{bmatrix} 
-	P_{A2}(s) \\
-	P_{B2}(s)
-	\end{bmatrix}
-	Y_2
-\end{equation}
-If it is assumed that the rear tire of the car experiences the exact same road conditions as the front tire, $Y1$ and $Y2$ can be expressed as functions of the same input, $Y$ where $Y2$ is simply a phase shifted version of $Y1$:
-\begin{equation}
-	\begin{split}
-		y_1(t) &= y(t)\\
-		y_2(t) &= y(t-t_0)
-	\end{split} \implies
-	\begin{split}
-		Y_1(s) &= Y(s)\\
-		Y_2(s) &= Y(s)e^{-st_0}
-	\end{split}
-\end{equation}
-Equation \ref{eq:matrix1} can then be rewritten as follows:
-\begin{equation}
-	\label{eq:matrix1}
-	Q
-	\begin{bmatrix} 
-	X_0 \\
-	\Theta 
-	\end{bmatrix}
-	= 
-	\left(\left[P_1\right]+\left[P_2\right]e^{-st_0}\right)Y
-\end{equation}
-Since the system is linear, the above matrix equation can be solved in two parts to obtain the transfer functions for $X_0$ and $\Theta$:
-\begin{equation}
-	\frac{\begin{bmatrix} 
-		X_0 \\
-		\Theta 
-	\end{bmatrix}}{Y} =
-	\left[H\right] = 
-	\left[H_1\right] + \left[H_2\right] = 
-	[Q^{-1}][P_1] + [Q^{-1}][P_1]e^{-st_0} 
-\end{equation}
-
-Using the MATLAB script shown in Appendix \ref{app:alternative}, the transfer function $H(s)$ is solved for symbolically as described above and the parameters for $I$, $M$, $k_1$, $k_2$, $c_1$, $c_2$, $a$ and $b$ are substituted into the equation.
-The inverse laplace transform is then taken to obtain the impulse response as shown in Figure \ref{fig:impResp}.
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=.8\textwidth]{./matlab/impResp.png}
-	\caption{Undamped Impulse Response}
-	\label{fig:impResp}
-\end{figure}
-
-By performing an Fast Fourier Transform on the undamped system the harmonics of the system can be approximated as shown in Figure \ref{fig:fftImpResp}.
-A sampling frequency of 200Hz and and a window of 30s was used for the FFT.
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width=.8\textwidth]{./matlab/fftImpResp.png}
-	\caption{FFT of the Undamped Impulse Response}
-	\label{fig:fftImpResp}
-\end{figure}
-From Figure \ref{fig:fftImpResp}, the largest harmonics of the system occur at approximately 2Hz for translational displacement and 4Hz for angular displacement.
-Note that these frequencies align closely with the natural frequencies that were determined analytically in the simplified case.
+\section{Dynamic Loading}
+MSXII will be using Ohlins TTX25 dampers for both the front and rear suspension. From the Ohlins' published data, \cite{tbd}, the maximum damping coefficient achievable is approximately 3.5kNs/m.
 

### 2017-09-27 FEABoundaryConditions/matlab/alternativeFormulation.m

@@ -1,0 +1,155 @@
+% ---------------------------------------------------------------------- %
+%
+% Suspension Spring Rates Project - alternativeFormulation.m
+% Function to setup transfer functions for a half car model to determine the
+% reponse of MSXII to disturbances in the road conditions
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2017
+%
+%
+% Notes:
+%   
+%   - 
+% ---------------------------------------------------------------------- %
+
+load_system('halfCarModel');
+close all;
+clear all
+
+% ----------------------------- Constants ------------------------------ %
+
+c1_ = 4830;     % Front damping coefficient (Ns/m)
+c2_ = 4830;     % Rear damping coefficient (Ns/m) 
+k1_ = 600*175.1*cosd(35);    % Front sring rate (N/m) 
+k2_ = 500*175.1*cosd(30);    % Rear sring rate (N/m)
+m_ = 550;       % Mass of the vehicle (kg)
+I_ = 550;       % Moment of interial about global x axis (kg*m^2)
+a_ = 1.32;       % Distance from COG to front tire (m)
+b_ = 1.28;       % Distance from COG to rear tire (m)
+
+% ---------------------- Forcing Function Parameters ------------------- %
+
+amplitude = 0.015;              % Amplitude of bumps on road surface
+bumpSparation = 2;              % Distance between peaks on the road surface (m)
+speed = 22.2;                   % Cruising speed (m/s)
+wb = 2*pi*speed/bumpSparation;  % Frequency of base excitation (rad/sec)
+timeDelay = (a_+b_)/speed;      % delay between input striking front and
+                                % rear wheels (s)
+                                
+% Update simulink parameters
+set_param('halfCarModel/DelayA','DelayTime',mat2str(timeDelay));
+set_param('halfCarModel/DelayB','DelayTime',mat2str(timeDelay));
+                              
+% ----------------- Create Symbolic Transfer Functions ----------------- %
+
[…]

### 2017-09-27 FEABoundaryConditions/matlab/simplifiedFormulation.m

@@ -1,0 +1,91 @@
+% ---------------------------------------------------------------------- %
+%
+% Suspension Spring Rates Project - simplifiedFormulation.m
+% Function to setup simplified transfer functions for a half car model to 
+% determine the reponse of MSXII to disturbances in the road conditions
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2017
+%
+%
+% Notes:
+%   
+%   - 
+% ---------------------------------------------------------------------- %
+
+close all;
+clear all;
+clc;
+
+% ----------------------------- Constants ------------------------------ %
+
+c_ = 4830;     % Front damping coefficient (Ns/m)
+k_ = 88600;    % Front sring rate (N/m) 
+M_ = 550;      % Mass of the vehicle (kg)
+I_ = 550;      % Moment of interial about global x axis (kg*m^2)
+a_ = 1.3;      % Distance from COG to front tire and rear tire (m)
+zx_ = 0.49;     % Damping ratio of translational  
+wnx_ = 17.95;   % Natural frequency of pitch
+zth_ = 0.64;    % Damping ratio of pitch
+wnth_ = 23.33;  % Natural frequency of pitch
+
+
+% ---------------------- Forcing Function Parameters ------------------- %
+
+A_ = 0.015;                     % Amplitude of bumps on road surface
+bumpSparation = 2;              % Distance between peaks on the road surface (m)
+speed = 22.2;                   % Cruising speed (m/s)
+wb_ = 2*pi*speed/bumpSparation; % Frequency of base excitation (rad/sec)
+t0_ = (2*a_)/speed;            % delay between input striking front and
+                                % rear wheels (s)
+
+% ----------------- Create Symbolic Transfer Functions ----------------- %                                
+                                
+syms t s M I a t0 A wb c k zx wnx zth wnth
+
+QA = s^2 + 2*c/M*s + wnx^2; % ODE A 
+QB = s^2 + 2*a^2*c/I*s + wnx^2; % ODE B
+
+phaseShift = exp(-s*t0);
+
[…]

### 2017-09-28 FEABoundaryConditions/FEABoundaryConditions.tex
updated static boundary conditons

@@
 	Track                       	 		& 1.60  & $m$  		\\
+	Tire Diameter                       	 		& 0.53  & $m$  		\\
+	Wheel Assembly Moment of Inertia ($J_w$)   	& 0.25  & $kgm^2$  		\\
 	Distance from Front Wheel to CoG ($a$)	& 1.32 & $m$  		\\
 	Distance from Road to CoG ($h$)		& 0.52 & $m$  		\\
-	Moment of Inertia Normal to Symmetry Plane at CoG ($I_{xx}$) & 550  & $kgm^2$ \\
+	%Moment of Inertia Normal to Symmetry Plane at CoG ($I_{xx}$) & 550  & $kgm^2$ \\
 	Full Wheel Travel in Compression 	& 45   & $mm$ 		\\
@@
 \end{figure}
-\subsection{Two Loading Conditions}
-Two cases of boundary conditions are considered for FEA.
-One case occurs when the suspension is statically loaded during the extreme case of braking while cornering.
-The other case is a dynamic loading scenario where the car suddenly hits a bump at speed.
+\subsection{Approach to FEA}
+\textit{explain the different loading conditions to be tested (dynamic, static, fatigue etc.)}
 
-It is important to note that the former case (that of braking while cornering) is specific to the front suspension.
-The equivalent worst case loading scenario for the rear would occur while accelerating during a corner.
-However, since MSXII's motors produce only ***Nm of torque while the brakes require ***Nm to slow the car at the required rate, the worst loading for the front will far exceed that of the rear.
-For simplicity, this value is treated as the worst case for both the front and the rear.
-
+\pagebreak
 \section{Static Loading}
 \label{sec:paramSelection}
-Assuming a smooth driving surface, the largest normal force acting on the tire will occur while braking in a turn.
-To approximate this loading condition, the a superposition of two static half car models is used; one for cornering and one for hard braking.
-\subsubsection{Hard Braking}\begin{figure}[h!]
+Assuming a smooth driving surface, the largest normal force acting on the front tire will occur while braking in a turn.
+Similarly, the largest normal force acting on the rear tire will occur while accelerating in a turn.
+To approximate these loading condition, the a superposition of two static half car models is used; one for cornering and one for hard braking.
+
+\subsection{Hard Braking}
+\label{sec:hardBraking}
+\begin{figure}[h!]
 	\centering
@@
 
+\subsection{Maximum acceleration}
+MSXII will use SCM-150s which have a peak torque of 135Nm, however from the manufacturer provided torque speed curves, a more reasonable upper limit to for peak torque sustainable as the car accelerates is approximately 100Nm \cite{scm}.
+From this torque value and a number of other vehicle parameters the vehicle's maximum acceleration can be calculated as follows:
+
+\begin{equation}
+	a = \frac{F_w}{\frac{1}{2}M}
+\end{equation}
+\begin{equation}
+	\alpha_w = \frac{T_m - F_wr}{J_w}
+\end{equation}
+\begin{equation}
+	\alpha_w = ar
+\end{equation}
+where $a$ is the acceleration of the vehicle, $F_w$ is the friction force acting on a single tire while accelerating, $J_w$ is the moment of inertia of the wheel assembly, $r$ is the radius of the tire, $T_m$ is the torque from the motor and $\alpha_w$ is the angular acceleration of the wheel.
+
+Rearranging the above three equations, the acceleration of the vehicle can be expressed as a function of motor torque as follows: 
+\begin{equation}
+\begin{split}
+	a &= \frac{T_m r}{\frac{1}{2}Mr^2 + J_w} \\
+	a &= \frac{(100Nm)(0.27m)}{\frac{1}{2}(550kg)(0.27m^2)^2 + 0.25kgm^2} = 1.33m/s^2
+\end{split}
+\end{equation}
+Solving for friction force on a single rear tire: 
+\begin{equation}
+	F_w = Ma = (550kg)(1.33m/s^2) = 732N
+\end{equation}
+Using the same approach as demonstrated in section \ref{sec:hardBraking}, the resulting normal force on the rear wheels can be calculated as follows: 
+\begin{equation}
+\begin{split}
+	bF_{Nr} &= hF_{w} + a(Mg - F_{Nr})\\
+	F_{Nr} &= \frac{hF_{w} + aMg}{a+b} = \frac{(0.52m)(732N)+(1.32m)(550kg)\left(9.81\frac{m}{s^2}\right)}{(1.32m)+(1.28m)} = 2.89kN
+\end{split}
+\end{equation}
+Note that 2.88kN is for both rear wheels. Therefore there is approximately \textbf{1.44kN} of normal force on each rear tire during maximum acceleration.
+
 \subsubsection{Cornering}
@@
 \end{figure}
-The ASC 2018 regulations stipulate that a vehicle must be able to navigate a figure of eight as shown in Figure \ref{fig:dynamicScrutineering} in less than 18 seconds \cite{ASC2018regs}.
-Since the average arc length of the figure of eight is $34\pi m$, the net cornering force required to navigate the figure of eight can be found as follows:
+To determine the lateral force subjected to the vehicle, the worst case of two scenarios is selected.
+The first scenario originates from the ASC 2018 regulations while the second draws its roots from actual highway geometry and speed limits.
+
+For the first scenario, the ASC 2018 regulations stipulate that a vehicle must be able to navigate a figure of eight as shown in Figure \ref{fig:dynamicScrutineering} in less than 18 seconds \cite{ASC2018regs}.
+Since the average arc length of the figure of eight is $34\pi m$, the net lateral force required to navigate the figure of eight can be found as follows:
 \begin{equation}
@@
 \end{equation}
+
+\begin{figure}[htbp]
+	\centering
+    \begin{subfigure}[b]{.4\textwidth}
+		\caption{Conestoga Pkwy. \& Wellington St.}
+		\centering
+        \includegraphics[height=4.5cm]{./LaTex/hw1.PNG}
+    \end{subfigure}
+    \begin{subfigure}[b]{.4\textwidth}
+		\caption{Conestoga Pkwy. \& University Ave.}
+		\centering
+        \includegraphics[height=4.5cm]{./LaTex/hw2.PNG}
+    \end{subfigure}
+
+    \bigskip
+    \begin{subfigure}[b]{.4\textwidth}
+		\caption{Conestoga Pkwy. \& King St.}
+		\centering
+        \includegraphics[height=4.5cm]{./LaTex/hw3.PNG}
+    \end{subfigure}
+    \begin{subfigure}[b]{.4\textwidth}
+		\caption{Speed Limit on Conestoga Pkwy Exits}
+		\centering
+        \includegraphics[height=4.5cm]{./LaTex/speedLimit.PNG}
+    \end{subfigure}
+    \caption{Highway Exist Around the University of Waterloo}
+	\label{fig:highwayExits}
+\end{figure}
+For the second scenario, the diameter of several highway exit ramps around the University of Waterloo were measured as shown in Figure \ref{fig:highwayExits}.
+The recommended speed limit on these exits 50$km/hr$.
+Assuming a worst case scenario where the vehicle takes an exit with a 100$m$ diameter at 60$km/hr$ (16.67$m/s$), required lateral force required to navigate this corner can be found as follows:
+\begin{equation}
+	F_{steer} = \frac{Mv^2}{r} = \frac{(550kg)(16.67m/s)^2}{50m} = 3.06kN
+\end{equation}
+
+Since the lateral force required in the second scenario is highest, 3.06kN will be used as the worst case cornering force.
+
 \begin{figure}[h!]
@@
 
-\noindent For the half car model shown if Figure \ref{fig:steerFBD}, summing the moments about the centre of gravity and the vertical forces results in the following equations: 
+Using the half car model shown if Figure \ref{fig:steerFBD}, summing the moments about the centre of gravity and the vertical forces results in the following equations: 
 \begin{equation}
@@
 	\frac{track}{2}F_{No} &= hF_{steer} + \frac{track}{2}(Mg - F_{Ni})\\
-	F_{No} &= \frac{hF_{steer} + \frac{track}{2}Mg}{track} = \frac{(0.52m)(2.28kN)+(0.80m)(550kg)\left(9.81\frac{m}{s^2}\right)}{1.60m} = 3.44kN
+	F_{No} &= \frac{hF_{steer} + \frac{track}{2}Mg}{track} = \frac{(0.52m)(3.06kN)+(0.80m)(550kg)\left(9.81\frac{m}{s^2}\right)}{1.60m} = 3.69kN
 \end{split}
 \end{equation}
-Note that the 3.44kN above is for both front and rear outside wheels.
-Assuming the force is evenly split between the front and rear, there is approximately \textbf{1.72kN} of normal force on each outside tire during maximum cornering.
-\subsubsection{Superposition}
-Both the braking and cornering conditions subject the front outside wheel to a normal force greater than the nominal normal force seen when the car is at rest.
+Solving for the inside wheel normal force: 
+\begin{equation}
+	F_{Ni} = Mg - F_{No} = (550kg)\left(9.81\frac{m}{s^2}\right) - 3.69kN = 1.71kN
+\end{equation}
+Note that the forces above are for both front and rear outside wheels.
+Assuming the force is evenly split between the front and rear, there is approximately \textbf{1.85kN} of normal force on each outside tire and \textbf{0.85kN} for each inside tire during maximum cornering.
+
+\subsection{Superposition}
+The braking accelerating and cornering conditions subject the wheels to normal forces that differ from those seen when the car is at rest.
 The change in normal force for each loading condition can be expressed as follows:
@@
 	\begin{split}
-		\Delta F_{N_{braking}} &= F_{Nf_{braking}} - \frac{Mg}{4} =  1.87kN - 1.35kN = 0.52kN\\
-		\Delta F_{N_{cornering}} &=  F_{No_{cornering}} - \frac{Mg}{4} =  1.72kN - 1.35kN = 0.37kN
+		\Delta F_{Nf_{braking}} &= F_{Nf_{braking}} - \frac{Mg}{4} =  1.87kN - 1.35kN = 0.52kN\\
+		\Delta F_{Nr_{accelerating}} &= F_{Nf_{braking}} - \frac{Mg}{4} =  1.44kN - 1.35kN = 0.09kN\\
+		\Delta F_{No_{cornering}} &=  F_{No_{cornering}} - \frac{Mg}{4} =  1.85kN - 1.35kN = 0.50kN \\
+		\Delta F_{Ni_{cornering}} &=  F_{No_{cornering}} - \frac{Mg}{4} =  0.85kN - 1.35kN = -0.50kN
 	\end{split}
@@
 To approximate the worst case loading condition, a superposition of the changes in normal force and the nominal normal force is applied.
+Note that the nominal force used is the same for each wheel which assumes a centre of gravity approximately centred along the wheelbase.
 \begin{equation}
-	F_{N_{max}} = \Delta F_{N_{braking}} + \Delta F_{N_{cornering}} + F_{N_{nominal}} = 0.52kN + 0.37kN + 1.35kN = 2.24kN
+	\begin{split}	
+	F_{Nfi_{max}} = \Delta F_{Nf_{braking}} + \Delta F_{Ni_{cornering}} + F_{N_{nominal}} = 0.52kN - 0.50kN + 1.35kN = 1.24kN \\
+	F_{Nfo_{max}} = \Delta F_{Nf_{braking}} + \Delta F_{No_{cornering}} + F_{N_{nominal}} = 0.52kN + 0.50kN + 1.35kN = 2.37kN \\
+	F_{Nri_{max}} = \Delta F_{Nr_{braking}} + \Delta F_{Ni_{cornering}} + F_{N_{nominal}} = 0.09kN - 0.50kN + 1.35kN = 0.94kN \\
+	F_{Nro_{max}} = \Delta F_{Nr_{braking}} + \Delta F_{No_{cornering}} + F_{N_{nominal}} = 0.09kN + 0.50kN + 1.35kN = 1.94kN 
+	\end{split}
 \end{equation}
 
+\subsection{Summary}
+Table \ref{tab:forcesSummary} summarizes the loading conditions used for static structural FEA simulations on the suspension and steering components.
+Note that the direction of the force corresponds's with the global coordinate system of the MSXII and the vectors shown in Figure \ref{fig:forcesSummary}.
+The assemblies shown are for the driver's side of the vehicle
+
+\begin{table}[h!]
+\centering
+\caption{Summary of Suspension Loading Conditions}
+\label{tab:summary}
+\begin{tabular}{llll}
+\textbf{Case}                                  & \textbf{Fx (kN)} & \textbf{Fy (kN)} & \textbf{Fz (kN)} \\
+Braking \& Cornering (Inside Front Wheel)      & 1.53             & 1.24             & 2.70             \\
+Braking \& Cornering (Outside Front Wheel)     & -1.53            & 2.37             & 2.70             \\
+Accelerating \& Cornering (Inside Rear Wheel)  & 1.53             & 0.94             & 0.73             \\
+Accelerating \& Cornering (Outside Rear Wheel) & -1.53            & 1.94             & 0.73            
+\end{tabular}
+\end{table}
+
+\begin{figure}[h!]
+	\centering
+    \begin{subfigure}[b]{.48\textwidth}
+		\caption{Font Suspension}
+		\centering
+        \includegraphics[height=9cm]{./LaTex/frontIsoForces.PNG}
+    \end{subfigure}
+    \begin{subfigure}[b]{.48\textwidth}
+		\caption{Rear Suspenion}
+		\centering
+        \includegraphics[height=9cm]{./LaTex/rearIsoForces.PNG}
+    \end{subfigure}
+    \caption{Summary of Suspension Loading Conditions}
+	\label{fig:forcesSummary}
+\end{figure}
+
+Since the lateral force on each of the wheels in particular is dependent on a number of tire properties that are difficult to model (slip angle, tire pressure, etc.).
+It was assumed that 50\% of the net lateral force will be the maximum lateral force experienced by any one of the four wheels at a given time.
+
 \section{Dynamic Loading}
-MSXII will be using Ohlins TTX25 dampers for both the front and rear suspension. From the Ohlins' published data, \cite{tbd}, the maximum damping coefficient achievable is approximately 3.5kNs/m.
+MSXII will be using Ohlins TTX25 dampers for both the front and rear suspension. From the Ohlins' published data, \cite{ohlins}, the maximum damping coefficient achievable is approximately 3.5kNs/m.
 

### 2017-09-28 MotorTorque/main.m

@@ -24,7 +24,7 @@ close all;
 
 % ------------------------------ Constants ---------------------------- %
 
-M = 450;        % Car mass (kg)
+M = 550;        % Car mass (kg)
 r = 0.267;      % Wheel radius (m)
 Cr = 0.0055;    % Rolling resistance coefficient - Winning Solar Car pg 18
 A = 1.8;        % Frontal area (m^2)
@@ -33,7 +33,7 @@ rho = 1.225;    % Density of air (kg/m^3)
 g = 9.81;       % Acceleration of gravity (m/s^2)
 
 % ------------------------------ Variables ---------------------------- %
-
+x
 a = [0:.5:2.5];                      % Acceleration in m/s^2
 v = [0:.1:33.33];                    % Velocity in m/s
 incline = [0, 5,10];            % Incline in %

### 2017-10-14 PVDR/WaterlooMidnightSun_ASC2018PVDR.tex
first draft of mechanical PVDR, pending several critical action items prior to submission, and pending integration of electrical PVDR

% LuaLaTeX
\documentclass[10pt]{article}

% Packages
\usepackage{fontspec}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage[backend=biber]{biblatex}
\usepackage[nonumberlist, toc]{glossaries}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[numbered]{./latex/mcode}
\usepackage{fancyhdr}
\usepackage{amsmath}
\usepackage{siunitx}

% Formatting
\geometry{letterpaper, portrait, margin=.85in}
\defaultfontfeatures{Ligatures=TeX}
\setmainfont[
    BoldFont       = Avenir Medium,
    ItalicFont     = Avenir Book Oblique,
    BoldItalicFont = Avenir Medium Oblique
]{Avenir Book}
\setmonofont{Andale Mono}

\bibliography{./latex/bib}
\newcommand\thetitle{Preliminary Vehicle Design Report}
\newcommand\theauthor{Midnight Sun Solar Car Team}
\newcommand\thedate{\today}

\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\lhead{\thetitle}
\rhead{\theauthor}

\begin{document}

% Title Ppage
\begin{titlepage}
	\vspace*{2cm}
	\centering
	\includegraphics[width=.25\textwidth]{./figures/midnightSunLogoCircle.png}\par
	\vspace{1.5cm}
	{\LARGE \theauthor \par}
	{\large University of Waterloo\par}
	\vspace{2.2cm}
	{\huge\bfseries \thetitle\par} 
	\vspace{0.2cm}
	\large MSXII %
	\vspace{2.2cm}	
	\par Prepared by:\\
	Minghao Ji\\
	minghao.ji@uwmidsun.com\par
	\thedate\par
	\vfill
	www.uwmidsun.com \par
	(519) 888-4567 x32978
\end{titlepage}

% Main Matter
\tableofcontents
\listoffigures
\listoftables

\section{History}
Midnight Sun was founded in 1988 at the University of Waterloo. The team has produced 11 solar-powered vehicles since its inception, numbered MSI through MSXI. MSX and its predecessors have been traditional Challenger class cars. MSXI was the team's first attempt at a Cruiser class vehicle, which ultimately suffered from design issues relating to its monocoque design. With MSXII, the team has regrouped and designed a new Cruiser vehicle from the gound up, focussing strongly on reliability, safety, and manufacturability.

\section{Mechanical}
\subsection{Overview}
MSXII is a dual-occupant Cruiser class vehicle. Its structural design uses a steel alloy tube chassis with non-structural fibreglass body panels. The majority of the vehicle's powertrain systems are located behind the occupant cell, including the battery pack, power junction box, motor controllers, and EV charger. The vehicle is rear-wheel driven by two hub-mounted DC motors. The locations of these components are marked in Figure \ref{fig:msxii-top-view} and Figure \ref{fig:msxii-side-view}. Major vehicle dimensions are summarized in Table \ref{tab:msxii-dimensions}.

\begin{figure}
\centering
%\includegraphics{figures/msxii-top-view}
\caption{Top view showing major dimensions and system positions}
\label{fig:msxii-top-view}
\end{figure}

\begin{figure}
\centering
%\includegraphics{figures/msxii-side-view}
\caption{Side view showing major dimensions and system positions}
\label{fig:msxii-side-view}
\end{figure}

\begin{table}
\centering
\begin{tabular}{cc}
\toprule
Parameter & Value \\
\midrule
Wheelbase & \SI{2.60}{\metre} \\
Track     & \SI{1.60}{\metre} \\
\bottomrule
\end{tabular}
\caption{Table of major vehicle dimensions}
\label{tab:msxii-dimensions}
\end{table}

\subsection{Chassis}
\subsubsection{Design}
The chassis is a tube structure enclosing the occupant cell with small front and rear subframes to support suspension hardpoints. Smaller non-structural tube members will extend from the main frame to provide support to body panels as necessary. The majority of tubes are round, with square tubes being used only for elements intended to mount suspension and powertrain components (battery pack, motor controllers). Chassis elements use SAE 4130N chromoly tubes with diameters or side lengths ranging from 0.750" to 1.250", and wall thicknesses ranging from 0.035" to 0.065". Isometric drawings of the chassis are shown in Figures \ref{fig:chassis-front}, \ref{fig:chassis-top}, and \ref{fig:chassis-side}.

\begin{figure}
\centering
%\includegraphics{figures/chassis-front}
\caption{Chassis, front view}
\label{fig:chassis-front}
\end{figure}
\begin{figure}
\centering
%\includegraphics{figures/chassis-top}
\caption{Chassis, top view}
\label{fig:chassis-top}
\end{figure}
\begin{figure}
\centering
%\includegraphics{figures/chassis-side}
\caption{Chassis, side view}
\label{fig:chassis-side}
\end{figure}

\subsubsection{Manufacturing}
Bending, cutting, and profiling of tubes was done by an external supplier. Welding of the chassis was done by a professional welder using GTAW, with support from team members for jigging and assembly. A complete welding jig was designed in CAD and manufactured using 80/20 aluminum extrusion tubing, shown in Figure \ref{fig:welding-jig}. Custom tube clamps were machined by the team to slot into jig members and fully support all chassis elements through multiple stages of assembly and welding.

\begin{figure}
\centering
%\includegraphics{figures/welding-jig}
\caption{Welding jig under construction}
\label{fig:welding-jig}
\end{figure}

\subsubsection{Roll cage}
The roll cage is integrated into the design of the chassis members enclosing the occupant cell. For the purposes of ASC regulations, the roll cage can be considered to include the A and B pillar hoops and elements connecting these two hoops lengthwise through the car. For clarity, these elements have been highlighted in Figure \ref{fig:roll-cage}.

\begin{figure}
\centering
%\includegraphics{figures/roll-cage}
\caption{Chassis with roll cage elements highlighted}
\label{fig:roll-cage}
\end{figure}

\subsection{Seat Belts}
MSXII is equipped with 5-point seat belts for both occupants. Seat belt hardpoints are mounted to metal tabs welded to chassis members, shown in Figure \ref{fig:seat-belt-tab-positions}. Figure \ref{fig:seat-belt-tab} shows one of these tabs. Figures \ref{fig:seat-belt-side-view} and \ref{fig:seat-belt-top-view} demonstrate compliance with ASC regulations 10.3.E.

\begin{figure}
\centering
%\includegraphics{figures/seat-belt-tab-positions}
\caption{Chassis mounting points for seat belts}
\label{fig:seat-belt-tab-positions}
\end{figure}

\begin{figure}
\centering
%\includegraphics{figures/seat-belt-tab}
\caption{Single seat belt tab}
\label{fig:seat-belt-tab}
\end{figure}

\begin{figure}
\centering
%\includegraphics{figures/seat-belt-side-view}
\caption{Side view of seat belt geometry}
\label{fig:seat-belt-side-view}
\end{figure}

\begin{figure}
\centering
%\includegraphics{figures/seat-belt-top-view}
\caption{Top view of seat belt geometry}
\label{fig:seat-belt-top-view}
\end{figure}

\subsection{Braking System}
\subsubsection{Primary Brakes}
The primary braking system is designed with dual redundant front braking, consisting of two identical systems with separate master cylinders, hydraulic lines, and calipers to each of the left and right wheels. The master cylinders are operated in unision via a single mechanical brake pedal. A schematic of the primary system is shown in Figure \ref{fig:primary-brakes-schematic}.

\begin{figure}
\centering
%\includegraphics{figures/primary-brakes-schematic}
\caption{Schematic of primary braking system}
\label{fig:primary-brakes-schematic}
\end{figure}

Front brake calipers are Willwood parts \href{http://www.wilwood.com/Calipers/CaliperProd.aspx?itemno=120-8374}{120-8374} (left) and \href{http://www.wilwood.com/Calipers/CaliperProd.aspx?itemno=120-8373}{120-8373} (right). Willwood's website indicates the caliper pads have \SI{2}{in\squared} (\SI{12.9}{\centi\metre\squared}) contact area and 0.30" (\SI{7.6}{\milli\metre}) depth including the backing plate.

The master cylinders are Willwood part \href{http://www.wilwood.com/MasterCylinders/MasterCylinderProd.aspx?itemno=260-3374}{206-3374}. They have a 3/8"-24 outlet size which will need to be connected via an adapter to the M10 x 1.25 inlets of the calipers. Brake lines will be DOT-approved hose assemblies sized for 3/8"-24. The brake pedal is Willwood part \href{http://www.wilwood.com/Pedals/PedalProd.aspx?itemno=340-15079}{340-15079} and is mounted in the standard configuration to the left of the accelerator pedal.

Brake rotors will be custom-machined within the supported disk thickness and diameter supported by the Willwood calipers. All mounting geometry on the suspension upright for primary calipers are designed to position the primary calipers in the optimal positions to meet braking specifications.

\subsubsection{Parking Brake}
The parking brake will be implemented by a cable-actuated caliper mounted at one of the front wheels. The caliper is the Tolomatic ME10LA, part number \href{https://www.tolomatic.com/products/product-details/me10-mechanical-disc-brake#/specs-order}{0732-0003}. It will share the same brake disk as the primary braking system but is otherwise mechanically independent. Due to unresolvable dimensioning differences between the Tolomatic and Willwood calipers, the parking brake caliper cannot be mounted perfectly as per the manufacturer's recommendations and will have only 80\% overlap with the brake disk. However, this is expected to provide more than sufficient countertorque to keep the vehicle stationary as per ASC regulations 10.6.

Figure \ref{fig:calipers-mounting-geometry} shows a front suspension upright with all three calipers mounted (two primary and one parking).

\begin{figure}
\centering
%\includegraphics{figures/caliper-mounting-geometry}
\caption{View of dual-primary and parking calipers on front suspension upright }
\label{fig:calipers-mounting-geometry}
\end{figure}

\subsection{Steering System}
The steering system uses a rear-steering, left-hand-drive rack and pinion from Helix (part number \href{http://www.helixsuspension.com/catalog/Steering/Manual-Steering-Racks/HEXSR5/Omni-Manual-Steering-Rack---Rear-Steering}{188754}).

The system uses a collapsible steering column from Sweet Mfg.\@, part number \href{https://sweetmfg.biz/product.php?productid=2836}{405-10310}. The steering system linkage uses multiple U-joints and is shown in Figure \ref{fig:steering-system}.

\begin{figure}
\centering
%\includegraphics{figures/steering-system}
\caption{View of linkage from steering column to rack and pinion}
\label{fig:steering-system}
\end{figure}

\subsection{Suspension}
\subsubsection{Front Suspension}
The front suspension uses dual A-arms with an outboard dampener and coilover spring. The upper and lower control arms are custom-built from 4130 chromoly tubes. The dampener and coilover are a single integrated assembly supplied by \"Ohlins, sold as the TTX25, which is marketed for Formula SAE vehicles. Figure \ref{fig:front-suspension} shows the front suspension assembly.

\begin{figure}
\centering
%\includegraphics{figures/front-suspension}
\caption{View of front suspension assembly}
\label{fig:front-suspension}
\end{figure}

TODO: rod-end sizing justification based on shear loads

\subsubsection{Rear Suspension}
The rear suspension uses independent trailing arms mounted outwards from the rear of the chasssis. The trailing arm itself is custom-machined and mounts the same \"Ohlins TTX25 shock absorber as the front suspension. Figure \ref{fig:rear-suspension} shows the rear suspension assembly.

\begin{figure}
\centering
%\includegraphics{figures/rear-suspension}
\caption{View of rear suspension assembly}
\label{fig:rear-suspension}
\end{figure}

TODO: rod-end sizing justification based on shear loads

\subsection{Structural Analysis}
FEA of suspension and steering components was done using ANSYS software. The specific loading conditions prescribed by ASC regulations are analyzed in this section.

\subsubsection{1G Turn}

\subsubsection{2G Bump}

\subsubsection{1G Braking}

\clearpage
\section{Electrical}

% Bibliography
%\pagebreak
%\printbibliography

% Appendix
%\pagebreak
%\appendix
%\section{Sample Appendix}
%\label{app:simple}
%\lstinputlisting[breaklines=true]{./matlab/script.m}
\end{document}

### 2017-10-14 FEABoundaryConditions/FEABoundaryConditions.tex
Major updates to boundary conditions following PVDR findings

@@
 MSXII will only have brake callipers on the front wheels however regenerative braking from the hub motors will also provide a braking force at the rear contact patches.
-Assuming a generous coefficient of static friction, $\mu _s$, of 1.0 and the extreme case where the both the font and rear tires are about to begin sliding, the combined braking force can be expressed as:
+Assuming the maximum braking condition stipulated in the ASC 2018 regulations where the vehicle decelerates at 1G using only the mechanical breaks, the required braking force can be approximated as follows:
 \begin{equation}
-	F_{braking} = F_{Nf} + F_{Nr} = \mu _s F_{N_{net}} =  \mu _s Mg = (1.0)(550kg)\left(9.81\frac{m}{s^2}\right) = 5.40kN
+	F_{braking} = F_{Nf} + F_{Nr} = Mg = (550kg)\left(9.81\frac{m}{s^2}\right) = 5.40kN
 \end{equation}
@@
 \subsection{Maximum acceleration}
-MSXII will use SCM-150s which have a peak torque of 135Nm, however from the manufacturer provided torque speed curves, a more reasonable upper limit to for peak torque sustainable as the car accelerates is approximately 100Nm \cite{scm}.
-From this torque value and a number of other vehicle parameters the vehicle's maximum acceleration can be calculated as follows:
+MSXII will use SCM-150s which have a peak torque of 135Nm. From this torque value and a number of other vehicle parameters the vehicle's maximum acceleration can be calculated as follows:
 
@@
 	a &= \frac{T_m r}{\frac{1}{2}Mr^2 + J_w} \\
-	a &= \frac{(100Nm)(0.27m)}{\frac{1}{2}(550kg)(0.27m^2)^2 + 0.25kgm^2} = 1.33m/s^2
+	a &= \frac{(135Nm)(0.27m)}{\frac{1}{2}(550kg)(0.27m^2)^2 + 0.25kgm^2} = 1.80m/s^2
 \end{split}
@@
 \begin{equation}
-	F_w = Ma = (550kg)(1.33m/s^2) = 732N
+	F_w = Ma = (550kg)(1.80m/s^2) = 990N
 \end{equation}
@@
 	bF_{Nr} &= hF_{w} + a(Mg - F_{Nr})\\
-	F_{Nr} &= \frac{hF_{w} + aMg}{a+b} = \frac{(0.52m)(732N)+(1.32m)(550kg)\left(9.81\frac{m}{s^2}\right)}{(1.32m)+(1.28m)} = 2.89kN
+	F_{Nr} &= \frac{hF_{w} + aMg}{a+b} = \frac{(0.52m)(990N)+(1.32m)(550kg)\left(9.81\frac{m}{s^2}\right)}{(1.32m)+(1.28m)} = 2.94kN
 \end{split}
 \end{equation}
-Note that 2.88kN is for both rear wheels. Therefore there is approximately \textbf{1.44kN} of normal force on each rear tire during maximum acceleration.
+Note that 2.94kN is for both rear wheels. Therefore there is approximately \textbf{1.47kN} of normal force on each rear tire during maximum acceleration.
 
@@
 The recommended speed limit on these exits 50$km/hr$.
-Assuming a worst case scenario where the vehicle takes an exit with a 100$m$ diameter at 60$km/hr$ (16.67$m/s$), required lateral force required to navigate this corner can be found as follows:
+Assuming a worst case scenario where the vehicle takes an exit with a 100$m$ diameter at 80$km/hr$ (22.22$m/s$), required lateral force required to navigate this corner can be found as follows:
 \begin{equation}
-	F_{steer} = \frac{Mv^2}{r} = \frac{(550kg)(16.67m/s)^2}{50m} = 3.06kN
+	F_{steer} = \frac{Mv^2}{r} = \frac{(550kg)(22.22m/s)^2}{50m} = 5.43kN
 \end{equation}
 
-Since the lateral force required in the second scenario is highest, 3.06kN will be used as the worst case cornering force.
+Since the lateral force required in the second scenario is highest, 5.34kN will be used as the worst case cornering force. Note that this value corresponds with 1G of cornering force.
 
@@
 	\frac{track}{2}F_{No} &= hF_{steer} + \frac{track}{2}(Mg - F_{Ni})\\
-	F_{No} &= \frac{hF_{steer} + \frac{track}{2}Mg}{track} = \frac{(0.52m)(3.06kN)+(0.80m)(550kg)\left(9.81\frac{m}{s^2}\right)}{1.60m} = 3.69kN
+	F_{No} &= \frac{hF_{steer} + \frac{track}{2}Mg}{track} = \frac{(0.52m)(5.43kN)+(0.80m)(550kg)\left(9.81\frac{m}{s^2}\right)}{1.60m} = 4.46kN
 \end{split}
@@
 \begin{equation}
-	F_{Ni} = Mg - F_{No} = (550kg)\left(9.81\frac{m}{s^2}\right) - 3.69kN = 1.71kN
+	F_{Ni} = Mg - F_{No} = (550kg)\left(9.81\frac{m}{s^2}\right) - 4.46kN = 0.93kN
 \end{equation}
 Note that the forces above are for both front and rear outside wheels.
-Assuming the force is evenly split between the front and rear, there is approximately \textbf{1.85kN} of normal force on each outside tire and \textbf{0.85kN} for each inside tire during maximum cornering.
+Assuming the force is evenly split between the front and rear, there is approximately \textbf{2.23kN} of normal force on each outside tire and \textbf{0.47kN} for each inside tire during maximum cornering.
+
+\subsection{Bump}
+As stipulated in the ASC 2018 regulations, the analysis for suspension and steering must include loading considerations for a 2G bump.
+Assuming this requirement is interpreted as the body of the vehicle accelerating upwards at 2G, the additional normal force subjected to a single tire can be expressed as follows:
+
+\begin{equation}
+	F_{N_{bump}} = \frac{2GM}{4} = \frac{2(9.81\frac{m}{s^2})(550kg)}{4} = 2.70kN
+\end{equation}
 
@@
 		\Delta F_{Nf_{braking}} &= F_{Nf_{braking}} - \frac{Mg}{4} =  1.87kN - 1.35kN = 0.52kN\\
-		\Delta F_{Nr_{accelerating}} &= F_{Nf_{braking}} - \frac{Mg}{4} =  1.44kN - 1.35kN = 0.09kN\\
-		\Delta F_{No_{cornering}} &=  F_{No_{cornering}} - \frac{Mg}{4} =  1.85kN - 1.35kN = 0.50kN \\
-		\Delta F_{Ni_{cornering}} &=  F_{No_{cornering}} - \frac{Mg}{4} =  0.85kN - 1.35kN = -0.50kN
+		\Delta F_{Nr_{accelerating}} &= F_{Nf_{braking}} - \frac{Mg}{4} =  1.47kN - 1.35kN = 0.12kN\\
+		\Delta F_{No_{cornering}} &=  F_{No_{cornering}} - \frac{Mg}{4} =  2.23kN - 1.35kN = 0.88kN\\
+		\Delta F_{Ni_{cornering}} &=  F_{No_{cornering}} - \frac{Mg}{4} =  0.47kN - 1.35kN = -0.88kN\\
+		\Delta F_{N_{bump}} &= F_{N_{bump}} - \frac{Mg}{4} =  2.70kN - 1.35kN = 1.35kN
 	\end{split}
@@
 	\begin{split}	
-	F_{Nfi_{max}} = \Delta F_{Nf_{braking}} + \Delta F_{Ni_{cornering}} + F_{N_{nominal}} = 0.52kN - 0.50kN + 1.35kN = 1.24kN \\
-	F_{Nfo_{max}} = \Delta F_{Nf_{braking}} + \Delta F_{No_{cornering}} + F_{N_{nominal}} = 0.52kN + 0.50kN + 1.35kN = 2.37kN \\
-	F_{Nri_{max}} = \Delta F_{Nr_{braking}} + \Delta F_{Ni_{cornering}} + F_{N_{nominal}} = 0.09kN - 0.50kN + 1.35kN = 0.94kN \\
-	F_{Nro_{max}} = \Delta F_{Nr_{braking}} + \Delta F_{No_{cornering}} + F_{N_{nominal}} = 0.09kN + 0.50kN + 1.35kN = 1.94kN 
+	F_{Nfi_{max}} = \Delta F_{Nf_{braking}} + \Delta F_{Ni_{cornering}} + 2F_{N_{nominal}} = 0.52kN - 0.88kN + 2.70kN = 2.34kN \\
+	F_{Nfo_{max}} = \Delta F_{Nf_{braking}} + \Delta F_{No_{cornering}} + 2F_{N_{nominal}} = 0.52kN + 0.88kN + 2.70kN = 4.10kN \\
+	F_{Nri_{max}} = \Delta F_{Nr_{braking}} + \Delta F_{Ni_{cornering}} + 2F_{N_{nominal}} = 0.12kN - 0.88kN + 2.70kN = 1.94kN \\
+	F_{Nro_{max}} = \Delta F_{Nr_{braking}} + \Delta F_{No_{cornering}} + 2F_{N_{nominal}} = 0.12kN + 0.88kN + 2.70kN = 3.70kN 
 	\end{split}
@@
 \centering
-\caption{Summary of Suspension Loading Conditions}
+\caption{Summary of Worst Case Loading Conditions}
 \label{tab:summary}
 \begin{tabular}{llll}
-\textbf{Case}                                  & \textbf{Fx (kN)} & \textbf{Fy (kN)} & \textbf{Fz (kN)} \\
-Braking \& Cornering (Inside Front Wheel)      & 1.53             & 1.24             & 2.70             \\
-Braking \& Cornering (Outside Front Wheel)     & -1.53            & 2.37             & 2.70             \\
-Accelerating \& Cornering (Inside Rear Wheel)  & 1.53             & 0.94             & 0.73             \\
-Accelerating \& Cornering (Outside Rear Wheel) & -1.53            & 1.94             & 0.73            
+\textbf{Worst Case Loading}  			       & \textbf{Fx (kN)} & \textbf{Fy (kN)} & \textbf{Fz (kN)} \\
+Braking \& Cornering (Inside Front Wheel)      & 2.72             & 2.34             & 2.70             \\
+Braking \& Cornering (Outside Front Wheel)     & -2.72            & 4.10             & 2.70             \\
+Accelerating \& Cornering (Inside Rear Wheel)  & 2.72             & 1.94             & 0.99             \\
+Accelerating \& Cornering (Outside Rear Wheel) & -2.72            & 3.70             & 0.99            
+\end{tabular}
+\end{table}
+
+\begin{table}[h!]
+\centering
+\caption{Summary of Bump Loading Conditons}
+\label{tab:summary}
+\begin{tabular}{llll}
+\textbf{Bump Loading Conditions}  			    & \textbf{Fx (kN)} & \textbf{Fy (kN)} & \textbf{Fz (kN)} \\
+All wheels										& 0.00            & 2.70             & 0.00            
+\end{tabular}
+\end{table}
+
+\begin{table}[h!]
+\centering
+\caption{Summary of Braking and Accelerating Loading Conditions}
+\label{tab:summary}
+\begin{tabular}{llll}
+\textbf{Braking and Accelerating}  			       & \textbf{Fx (kN)} & \textbf{Fy (kN)} & \textbf{Fz (kN)} \\
+Braking (Front Wheels)      					   & 0.00             & 1.87             & 2.70             \\
+Accelerating (Rear Wheel)  					   & 0.00             & 1.44             & 0.99             \\
+\end{tabular}
+\end{table}
+
+\begin{table}[h!]
+\centering
+\caption{Summary of Cornering Loading Conditions}
+\label{tab:summary}
+\begin{tabular}{llll}
+\textbf{Cornering Loading}  			       & \textbf{Fx (kN)} & \textbf{Fy (kN)} & \textbf{Fz (kN)} \\
+Inside Wheel     						   & 2.72             & 0.47             & 0.00             \\
+Outside Wheel   						   & -2.72            & 2.23             & 0.00             \\
 \end{tabular}

### 2017-10-14 PVDR/WaterlooMidnightSun_ASC2018PVDR.tex

@@
 \section{Mechanical}
+\subsection{Contacts}
+Questions regarding the mechanical design and implementation of MSXII should be directed to one of the following contacts:
+
+\begin{table}[!h]
+\centering
+\begin{tabular}{llll}
+\toprule
+Name            & Title               & Phone        & Email \\
+\midrule        
+Tak Alguire     & Project Manager     & 519-574-4610 & tak.alguire@uwmidsun.com \\
+Minghao Ji      & Engineering Manager & 519-500-1292 & minghao.ji@uwmidsun.com \\
+Devon Copeland  & Mechanical Lead     & 226-792-7383 & devon.copeland@uwmidsun.com \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \subsection{Overview}
@@
 Track        & \SI{1.60}{\metre} \\
+Total length & TODO              \\
+Total width  & TODO              \\
+Height       & TODO              \\
 \bottomrule
@@
 
+\subsection{Contacts}
+Questions regarding the electrical design and implementation of MSXII should be directed to one of the following contacts:
+
+\begin{table}[!h]
+\centering
+\begin{tabular}{llll}
+\toprule
+Name        & Title               & Phone        & Email \\
+\midrule
+Tak Alguire & Project Manager     & 519-574-4610 & tak.alguire@uwmidsun.com \\
+Minghao Ji  & Engineering Manager & 519-500-1292 & minghao.ji@uwmidsun.com \\
+Titus Chow  & Electrical Lead     & 226-978-7104 & titus.chow@uwmidsun.com \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\subsection{Overview}
+Midnight Sun XII's electrical system consists of high- and low-voltage domains that are electrically isolated for safety.
+The high-voltage system includes the Sunpower E-series solar cells, Nomura maximum power point trackers, Tritium motor controllers and NGM SCM-150 motors.
+These are externally-sourced components purchased by the team and interface with the vehicle's custom embedded systems.
+
+The low-voltage system is comprised of custom circuit-boards serving several functions for monitoring and controlling the vehicle, primarily: driver controls, power distribution, battery management, and external lights.
+Several smaller PCBs are located throughout the vehicle to support various sensor interfacing and data collection functions.
+Most boards communicate over a unified CAN bus, with some nodes supporting smaller subsystems over I2C or SPI.
+The low-voltage power rail is normally provided by the main battery pack via DC-DC converters, but can also be switched to an auxiliary 12V Ni-MH battery during startup or fault modes.
+
+Due to the different CAN specifications of the Tritium WaveSculptor motor controllers, they are allocated a separate Powertrain CAN network which interfaces with the primary CAN bus via dedicated transceiver boards.
+A block diagram of all electrical systems is shown in Figure \ref{fig:msxii-electrical-full-block-diagram}.
+A block diagram of only the vehicle's high-voltage systems is shown in Figure \ref{fig:msxii-electrical-hv-block-diagram}.
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figures/msxii-electrical-full-block-diagram}
+\caption{Block diagram of all electrical systems}
+\label{fig:msxii-electrical-full-block-diagram}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{figures/msxii-electrical-hv-block-diagram}
+\caption{Block diagram of all high-voltage electrical systems}
+\label{fig:msxii-electrical-hv-block-diagram}
+\end{figure}
+
+\subsection{Battery}
+For MSXII the team chose to use MJ1 18650 cells produced by LG Chem Ltd.
+36 MJ1 cells are put in parallel to form a module, and 36 modules are put in series to provide a \SI{130.86}{\volt} nominal, \SI{16.5}{\kilo\watt\hour} capacity battery pack.
+Within each module, cells will be spot welded to nickel tabs, which are themselves soldered to copper bus bars forming a high-ampacity interconnect.
+Bus bars of adjacent modules are mechanically connected together to create series connections through the pack.
+
+The cells were sourced from Liion Wholesale in early 2017.
+
+\subsubsection{Testing Methodology}
+Battery testing is expected to be completed in 4 main phases, outlined below.
+
+\paragraph{Phase One}
+The first phase requires verification of the battery management system (BMS).
+This will test the system's ability to measure and monitor voltage, current, and temperature measurements from each of the modules.
+The system must be able to correctly respond to under-voltage, over-voltage, over-current, and over-temperature conditions to demonstrate that the BMS can actively protect the battery.
+Phase one is currently underway and is expected to be complete by November 2017.
+
+\paragraph{Phase Two}
+The second phase requires testing module performance and reliability.
+A prototype module containing 36 cells in parallel will be built and tested.
+This module will undergo charge cycle tests using a benchtop power supply and an electronic load.
+The BMS will be connected to log data and ensure safety.
+The team will then build 3 battery modules to test the series connections between modules.
+Phase two is expected to be complete by December 2017.
+
+\paragraph{Phase Three}
+The third phase involves the full manufacturing and testing of the battery pack.
+The team will be able to assess the overall performance of the battery and make necessary modifications to the cooling system prior to installing the battery in the car.
+The pack will be discharged at 1C, which is approximately 120A of current.
+This will be accomplished by either powering the vehicle motors mounted to a dynamometer, or by using a salt water load with a resistance designed for the target current draw.
+Phase three is expected to be completed by the end of January 2018.
+
+\paragraph{Phase Four}
+The final phase of battery testing will involve building a mechanical enclosure and integrating the pack into the vehicle.
+With a drivable vehicle, the team will be able to test the pack under real world conditions, and make any changes necessary prior to the race.
+Reliability of the modules and connections under harsh environmental conditions can be evaluated during road tests.
+Since this requires the vehicle to be in a drivable state, it is expected to begin in April or May 2018 and finish before July.
+
 % Bibliography

### 2017-10-16 PVDR/WaterlooMidnightSun_ASC2018PVDR.tex
work in progress

@@
 
-\section{Mechanical}
-\subsection{Contacts}
-Questions regarding the mechanical design and implementation of MSXII should be directed to one of the following contacts:
+\section{Contacts}
+Questions regarding the design and implementation of MSXII should be directed to one of the following contacts:
 
@@
 Devon Copeland  & Mechanical Lead     & 226-792-7383 & devon.copeland@uwmidsun.com \\
+Titus Chow      & Electrical Lead     & 226-978-7104 & titus.chow@uwmidsun.com \\
 \bottomrule
@@
 
+\section{Mechanical}
 \subsection{Overview}
@@
 The vehicle is rear-wheel driven by two hub-mounted DC motors.
-The locations of these components are marked in Figure \ref{fig:msxii-top-view} and Figure \ref{fig:msxii-side-view}.
-Major vehicle dimensions are summarized in Table \ref{tab:msxii-dimensions}.
+The locations of these components are marked in Figure \ref{fig:msxii-top-view-annotated} and Figure \ref{fig:msxii-side-view-annotated}.
+Major vehicle parameters are summarized in Table \ref{tab:msxii-dimensions}.
 
@@
 \centering
-%\includegraphics{figures/msxii-top-view}
+\includegraphics[width=\textwidth]{figures/msxii-top-view-annotated}
 \caption{Top view showing major dimensions and system positions}
-\label{fig:msxii-top-view}
+\label{fig:msxii-top-view-annotated}
 \end{figure}
@@
 \centering
-%\includegraphics{figures/msxii-side-view}
+\includegraphics[width=\textwidth]{figures/msxii-side-view-annotated}
 \caption{Side view showing major dimensions and system positions}
-\label{fig:msxii-side-view}
+\label{fig:msxii-side-view-annotated}
 \end{figure}
@@
 Track        & \SI{1.60}{\metre}    \\
-Total length & TODO              \\
-Total width  & TODO              \\
-Height       & TODO              \\
+Max length   & \SI{4.68}{\metre}    \\
+Max width    & \SI{2.02}{\metre}    \\
+Max height   & \SI{1.23}{\metre}    \\
+Gross weight & \SI{550}{\kilo\gram} \\
 \bottomrule
@@
 \centering
-%\includegraphics{figures/chassis-front}
+\includegraphics[width=0.75\textwidth]{figures/chassis-front}
 \caption{Chassis, front view}
@@
 \centering
-%\includegraphics{figures/chassis-top}
+\includegraphics[width=0.9\textwidth]{figures/chassis-top}
 \caption{Chassis, top view}
@@
 \centering
-%\includegraphics{figures/chassis-side}
+\includegraphics[width=\textwidth]{figures/chassis-side}
 \caption{Chassis, side view}
@@
 \centering
-%\includegraphics{figures/welding-jig}
-\caption{Welding jig under construction}
+\includegraphics[width=0.75\textwidth]{figures/welding-jig}
+\caption{Welding jig on rotating mount}
 \label{fig:welding-jig}
@@
 \centering
-%\includegraphics{figures/roll-cage}
+\includegraphics[width=0.75\textwidth]{figures/roll-cage}
 \caption{Chassis with roll cage elements highlighted}
@@
 \centering
-%\includegraphics{figures/seat-belt-tab-positions}
+\includegraphics[width=0.75\textwidth]{figures/seat-belt-tab-positions}
 \caption{Chassis mounting points for seat belts}
@@
 \centering
-%\includegraphics{figures/seat-belt-tab}
+\includegraphics[width=0.75\textwidth]{figures/seat-belt-tab}
 \caption{Single seat belt tab}
@@
 \centering
-%\includegraphics{figures/seat-belt-side-view}
+\includegraphics[width=0.75\textwidth]{figures/seat-belt-side-view}
 \caption{Side view of seat belt geometry}
@@
 \centering
-%\includegraphics{figures/seat-belt-top-view}
+\includegraphics[width=0.75\textwidth]{figures/seat-belt-top-view}
 \caption{Top view of seat belt geometry}
@@
 \centering
-%\includegraphics{figures/primary-brakes-schematic}
+%\includegraphics[width=\textwidth]{figures/primary-brakes-schematic}
 \caption{Schematic of primary braking system}
@@
 \centering
-%\includegraphics{figures/caliper-mounting-geometry}
+\includegraphics[width=0.5\textwidth]{figures/caliper-mounting-geometry}
 \caption{View of dual-primary and parking calipers on front suspension upright }
@@
 The system uses a collapsible steering column from Sweet Mfg.\@, part number \href{https://sweetmfg.biz/product.php?productid=2836}{405-10310}.
-The steering system linkage uses multiple U-joints and is shown in Figure \ref{fig:steering-system}.
+The steering system linkage uses multiple U-joints and is shown in Figure \ref{fig:steering-assembly}.
 
@@
 \centering
-%\includegraphics{figures/steering-system}
+\includegraphics[width=0.75\textwidth]{figures/steering-assembly}
 \caption{View of linkage from steering column to rack and pinion}
-\label{fig:steering-system}
+\label{fig:steering-assembly}
 \end{figure}
@@
 \centering
-%\includegraphics{figures/front-suspension}
-\caption{View of front suspension assembly}
+\includegraphics[width=0.75\textwidth]{figures/front-suspension}
+\caption{View of front suspension assembly (parking brake obstructed by hub)}
 \label{fig:front-suspension}
@@
 \centering
-%\includegraphics{figures/rear-suspension}
+\includegraphics[width=0.75\textwidth]{figures/rear-suspension}
 \caption{View of rear suspension assembly}
@@
 \section{Electrical}
-
-\subsection{Contacts}
-Questions regarding the electrical design and implementation of MSXII should be directed to one of the following contacts:
-
-\begin{table}[!h]
-\centering
-\begin{tabular}{llll}
-\toprule
-Name        & Title               & Phone        & Email \\
-\midrule
-Tak Alguire & Project Manager     & 519-574-4610 & tak.alguire@uwmidsun.com \\
-Minghao Ji  & Engineering Manager & 519-500-1292 & minghao.ji@uwmidsun.com \\
-Titus Chow  & Electrical Lead     & 226-978-7104 & titus.chow@uwmidsun.com \\
-\bottomrule
-\end{tabular}
-\end{table}
-
 \subsection{Overview}
@@
 
-Due to the different CAN specifications of the Tritium WaveSculptor motor controllers, they are allocated a separate Powertrain CAN network which interfaces with the primary CAN bus via dedicated transceiver boards.
+Due to the different CAN specifications of the Tritium WaveSculptor motor controllers, they are allocated a separate powertrain CAN network which interfaces with the primary CAN bus via dedicated transceiver boards.
 A block diagram of all electrical systems is shown in Figure \ref{fig:msxii-electrical-full-block-diagram}.

### 2017-10-16 PVDR/WaterlooMidnightSun_ASC2018PVDR.tex
Final revision sent to ASC

@@
 % Packages
-\usepackage{fontspec}
+\usepackage[no-math]{fontspec}
 \usepackage{geometry}
@@
 \usepackage{booktabs}
+\usepackage{float}
+\usepackage{subcaption}
 \usepackage[numbered]{./latex/mcode}
@@
 \setmonofont{Andale Mono}
+\sisetup{detect-all}
 
@@
 The locations of these components are marked in Figure \ref{fig:msxii-top-view-annotated} and Figure \ref{fig:msxii-side-view-annotated}.
-Major vehicle parameters are summarized in Table \ref{tab:msxii-dimensions}.
+Major vehicle dimensions are summarized in Table \ref{tab:msxii-dimensions}.
 
@@
 \toprule
-Parameter & Value \\
-\midrule
 Wheelbase    & \SI{2.60}{\metre}    \\
@@
 Max height   & \SI{1.23}{\metre}    \\
-Gross weight & \SI{550}{\kilo\gram} \\
 \bottomrule
@@
 \end{figure}
+
 \begin{figure}
@@
 
-\subsection{Seat Belts}
+\subsection{Seat belts}
 MSXII is equipped with 5-point seat belts for both occupants.
@@
 
-\subsection{Braking System}
-\subsubsection{Primary Brakes}
+\subsection{Braking system}
+\subsubsection{Primary brakes}
 The primary braking system is designed with dual redundant front braking, consisting of two identical systems with separate master cylinders, hydraulic lines, and calipers to each of the left and right wheels.
@@
 \centering
-%\includegraphics[width=\textwidth]{figures/primary-brakes-schematic}
+\includegraphics[width=\textwidth]{figures/primary-brakes-schematic}
 \caption{Schematic of primary braking system}
@@
 The master cylinders are Willwood part \href{http://www.wilwood.com/MasterCylinders/MasterCylinderProd.aspx?itemno=260-3374}{206-3374}.
-They have a 3/8"-24 outlet size which will need to be connected via an adapter to the M10 x 1.25 inlets of the calipers.
-Brake lines will be DOT-approved hose assemblies sized for 3/8"-24.
+They have a 3/8"-24 outlet size which will need to be connected via a suitable brake line to the M10 x 1.25 inlets of the calipers.
+Brake lines will be DOT-approved hose assemblies sized for 3/8"-24 on one end and M10 on the other end via a banjo joint.
 The brake pedal is Willwood part \href{http://www.wilwood.com/Pedals/PedalProd.aspx?itemno=340-15079}{340-15079} and is mounted in the standard configuration to the left of the accelerator pedal.
 
-Brake rotors will be custom-machined within the supported disk thickness and diameter supported by the Willwood calipers.
+Brake rotors will be custom-machined to the disk thickness and diameter supported by the Willwood calipers.
 All mounting geometry on the suspension upright for primary calipers are designed to position the primary calipers in the optimal positions to meet braking specifications.
 
-\subsubsection{Parking Brake}
+\subsubsection{Parking brake}
 The parking brake will be implemented by a cable-actuated caliper mounted at one of the front wheels.
@@
 \includegraphics[width=0.5\textwidth]{figures/caliper-mounting-geometry}
-\caption{View of dual-primary and parking calipers on front suspension upright }
+\caption{View of dual-primary calipers and parking caliper on front suspension upright }
 \label{fig:calipers-mounting-geometry}
@@
 
-\subsection{Steering System}
+\subsection{Steering system}
 The steering system uses a rear-steering, left-hand-drive rack and pinion from Helix (part number \href{http://www.helixsuspension.com/catalog/Steering/Manual-Steering-Racks/HEXSR5/Omni-Manual-Steering-Rack---Rear-Steering}{188754}).
-
 The system uses a collapsible steering column from Sweet Mfg.\@, part number \href{https://sweetmfg.biz/product.php?productid=2836}{405-10310}.
-The steering system linkage uses multiple U-joints and is shown in Figure \ref{fig:steering-assembly}.
+The linkage uses dual U-joints to circumvent interfering brake system components and is shown in Figure \ref{fig:steering-assembly}.
 
@@
 \includegraphics[width=0.75\textwidth]{figures/steering-assembly}
-\caption{View of linkage from steering column to rack and pinion}
+\caption{View of steering assembly}
 \label{fig:steering-assembly}
@@
 \subsection{Suspension}
-\subsubsection{Front Suspension}
+\subsubsection{Front suspension}
 The front suspension uses dual A-arms with an outboard dampener and coilover spring.
@@
 \includegraphics[width=0.75\textwidth]{figures/front-suspension}
-\caption{View of front suspension assembly (parking brake obstructed by hub)}
+\caption{View of front suspension assembly (parking brake caliper obstructed by hub)}
 \label{fig:front-suspension}
@@
 
-TODO: rod-end sizing justification based on shear loads
+The front suspension lower control arm bears the majority of the shear stress from road forces.
+Based on hand calculations for a cantilever beam, the worst-case von Mises stress is \SI{183}{\mega\pascal}.
+This dictates the sizing of the rod end.
 
-\subsubsection{Rear Suspension}
+\subsubsection{Rear suspension}
 The rear suspension uses independent trailing arms mounted outwards from the rear of the chasssis.
@@
 
-TODO: rod-end sizing justification based on shear loads
+\subsection{Finite element analysis}
+Finite element analysis was performed using ANSYS on the chassis and all suspension and steering components (downstream from the rack and pinion).
+FEA of the chassis followed the specifications provided by the ASC regulations.
+A general methodology was to fix a solid beam or wall in space, and apply a 5G acceleration to the chassis against it.
+Images for FEA simulations of each requested scenario are shown in Appendix \ref{sec:fea-chassis-images}.
 
-\subsection{Structural Analysis}
-FEA of suspension and steering components was done using ANSYS software. The specific loading conditions prescribed by ASC regulations are analyzed in this section.
+Worst-case loading forces on each wheel were used to derive appropriate contraints and forces for FEA simulation of the suspension and steering system.
+A complete derivation and summary of loading forces on each wheel is given in Appendix \ref{sec:loading-conditions}.
+Images of all FEA simulations for suspension and steering are available in Appendix \ref{sec:fea-parts-images}.
 
-\subsubsection{1G Turn}
+\subsubsection{Front suspension}
+\paragraph{Clevises}
+Hand calculations were used to resolve the forces from the road as loading onto the clevises.
+The back face of the clevis was constrained with compresion only support.
+Cylindrical supports were used for the bolt holes and fasterner pretention was added.
+Frictional contacts were added between the fasterners and the clevis.
 
-\subsubsection{2G Bump}
+\paragraph{Control arms}
+Contacts were modeled between the control arm tubes and the bearing plate. Forces were applied to the bearing plate based on hand calculations.
 
-\subsubsection{1G Braking}
+\paragraph{Uprights}
+A dummy wheel was modeled to which forces were applied.
+The dummy wheel contacted the spindle with a frictional contact which when applied a load to the upright.
+The brake forces were appled to the caliper mounting holes.
+Assuming only one break caliper were taking all the force.
+
+\paragraph{Spindle}
+A dummy wheel was modeled to which forces were applied. The dummy wheel contacted the spindle with a frictional contact which when applied a load to the upright.
+
+\paragraph{Upper bearing plate}
+Contacts were modeled between the control arm tubes and the bearing plate, forces were applied to the bearing plate based on hand calculations.
+
+\paragraph{Coilover control arm tab}
+The maximum force on the front wheel was resolved as a compressive load in the coilover and applied to this tab.
+
+\paragraph{Coilover mounting plate}
+Hand calculations were used to resolve the forces from the road as loading onto the clevises.
+For this simulation a frictional contact was added between the clevis and the mounting plate.
+Dummy chassis tubes were modeled behind the mounting plate and welded edges were fixted.
+
+\subsubsection{Rear suspension}
+\paragraph{Clevises}
+The worst case of four loading scenarios was taken (2 clevis positions, each with either inside or outside wheel when rounding a courner).
+It was assumed that all the lateral force was applied to a single clevis because this is a statically indeterminate problem on paper (so went with worst case).
+
+\paragraph{Coilover tab}
+The maximum normal force and acceleration force on the rear wheel was resolved as a compressive load in the coilover and applied to this tab.
+
+\subsubsection{Steering}
+\paragraph{Steering arms}
+A compression support was added on the face datuming off of the upright.
+The maximum lateral force on the tire during 1G cornering was resolved as a torque on the upright.
+This toque was applied by a bearing load acting at the end of the steering arm.
+
+\paragraph{Tie rods}
+A similar analysis to that for the steering arm was used.
+Force was applied to one end of the tie rod while the other remained fixed.
+Both sides of the tie rod were tested with an applied force and fixed support.
+
+\subsubsection{Results}
+All FEA results for suspension and steering components are summarized in Table \ref{tab:fea-results-suspension-steering}.
+Note that FEA for the rear suspension trailing arm and brace, as well as front suspension lower control arm, are currently unavailable.
+Simulations of the rear trailing arm are experiencing convergence and stress concentration issues.
+The front lower control arm is undergoing slight modifications to increase its safety factor under maximum loading conditions.
+\begin{table}
+\centering
+\begin{tabular}{lllrr}
+\toprule
+Part Number & Part Name                    & Material             & Max Stress (\si{\mega\pascal}) & Safety Factor \\
+\midrule
+MS00015     & Coilover Clevis              & Aluminum, 7075       &                 147 &           3.4 \\
+MS00029     & Upper Control Arm Clevis     & Aluminum, 7075       &                 191 &           2.6 \\
+MS00102     & Lower Control Arm Clevis     & Aluminum, 7075       &                 198 &           2.5 \\
+MS00017     & Upper Control Arm            & Chromoly Steel, 4130 &                 199 &           2.3 \\
+MS00018     & Lower Control Arm            & Chromoly Steel, 4130 &                   X &             X \\
+MS00019     & Upright                      & Aluminum, 7075       &                 210 &           2.4 \\
+MS00020     & Spindle                      & Chromoly Steel, 4140 &                 177 &           2.3 \\
+MS00030     & Upper Bearing Plate          & Chromoly Steel, 4130 &                 141 &           3.3 \\
+MS00036     & Coilover Control Arm Tab     & Chromoly Steel, 4130 &                 159 &           2.9 \\
+MS00038     & Coilover Mounting Plate      & Chromoly Steel, 4130 &                 147 &           3.1 \\
+MS00086     & Rear Suspension Clevis Mount & Chromoly Steel, 4130 &                 231 &           2.0 \\
+MS00047     & Rear Suspension Coilover Tab & Chromoly Steel, 4130 &                 147 &           3.1 \\
+MS00025     & Trailing Arm                 & Aluminum, 7075       &                   X &             X \\
+MS00026     & Rear Brace                   & Aluminum, 7075       &                   X &             X \\
+MS00074     & Steering Arm                 & Aluminum, 6061-T6    &                  77 &           3.5 \\
+MS00101     & Tie Rod                      & Aluminum, 6061-T6    &                  41 &           6.6 \\
+\bottomrule
+\end{tabular}
+\caption{Summary of FEA results for suspension and steering components}
+\label{tab:fea-results-suspension-steering}
+\end{table}
 
@@
 % Appendix
-%\pagebreak
-%\appendix
-%\section{Sample Appendix}
-%\label{app:simple}
-%\lstinputlisting[breaklines=true]{./matlab/script.m}
+\pagebreak
+\appendix
+\section{Loading conditions}
+\label{sec:loading-conditions}
+FEA of the chassis, suspension, and steering system was done using ANSYS software.
+Loading conditions were derived from both ASC requirements and real-world scenarios.
+Assuming a smooth driving surface, the largest normal force acting on the front tire will occur while braking in a turn.
+Similarly, the largest normal force acting on the rear tire will occur while accelerating in a turn.
+To approximate these loading condition, the a superposition of two static half car models is used; one for cornering and one for hard braking.
+
+Table \ref{tab:loading-conditions-params} lists vehicle parameters that are relevant in determining the loading conditions in the suspension.
+The mass and moment of inertia were computed using CAD software.
+
+\begin{table}
+\centering
+\begin{tabular}{lrl}
+\toprule
+Mass ($M$)                                  &  550 & \si{\kilo\gram}              \\
+Wheelbase                                   & 2.60 & \si{\metre}                  \\
+Track                                       & 1.60 & \si{\metre}                  \\
+Tire Diameter                               & 0.53 & \si{\metre}                  \\
+Wheel Assembly Moment of Inertia ($J_w$)    & 0.25 & \si{\kilogram\metre\squared} \\
+Distance from Front Wheel to CoG ($a$)      & 1.32 & \si{\metre}                  \\
+Distance from Front Wheel to CoG ($b$)      & 1.28 & \si{\metre}                  \\
+Distance from Road to CoG ($h$)             & 0.52 & \si{\metre}                  \\
+Full Wheel Travel in Compression            &   45 & \si{\milli\metre}            \\
+Front Coilover Angle from Vertical          &   40 & \si{\degree}                 \\
+Rear Coilover Angle from Vertical           &   30 & \si{\degree}                 \\
+\bottomrule
+\end{tabular}
+\caption{Important Vehicle Parameters}
+\label{tab:loading-conditions-params}
+\end{table}
+
+\subsection{Braking}
+Figure \ref{fig:braking-fbd} shows a free body diagram of a half car model undergoing braking. The required braking force can be approximated as follows: 
+\begin{equation}
+F_\mathrm{braking} = F_\mathrm{Nf} + F_\mathrm{Nr} = Mg = (\SI{550}{\kilo\gram})(\SI{9.81}{\metre\per\second\squared}) = \SI{5.40}{\kilo\newton}
+\end{equation}
+From Figure \ref{fig:braking-fbd}, summing the moments about the centre of gravity and the vertical forces results in the following equations: 
+\begin{gather}
+aF_\mathrm{Nf} = hF_\mathrm{braking} + bF_\mathrm{Nr} \\
+F_\mathrm{Nf} + F_\mathrm{Nr} = Mg
+\end{gather}
+Solving for the front normal force: 
+\begin{equation}
+\begin{split}
+aF_\mathrm{Nf} &= hF_\mathrm{braking} + b(Mg - F_\mathrm{Nf})\\
+F_\mathrm{Nf} &= \frac{hF_\mathrm{braking} + bMg}{a+b} = \frac{(\SI{0.52}{\metre})(\SI{5.40}{\kilo\newton})+(\SI{1.28}{\metre})(\SI{550}{\kilo\gram})(\SI{9.81}{\metre\per\second\squared})}{(\SI{1.32}{\metre})+(\SI{1.28}{\metre})} = \SI{3.74}{\kilo\newton}
+\end{split}
+\end{equation}
+Note that \SI{3.74}{\kilo\newton} is for both front wheels. Therefore there is approximately \SI{1.87}{\kilo\newton} of normal force on each front tire during hard braking.
+
+\begin{figure}
+\centering
+\includegraphics[width=.5\textwidth]{figures/braking-fbd}
+\caption{Free body diagram of vehicle braking}
+\label{fig:braking-fbd}
+\end{figure}
+
+\subsection{Acceleration}
+MSXII uses NGM SCM-150 motors which have a peak torque of \SI{135}{\newton\metre}.
+From this torque value and a number of other vehicle parameters the vehicle's maximum acceleration can be calculated as follows:
+\begin{gather}
+a = \frac{F_w}{\frac{1}{2}M} \\
+\alpha_w = \frac{T_m - F_wr}{J_w} \\
+\alpha_w = ar
+\end{gather}
+where $a$ is the acceleration of the vehicle, $F_w$ is the friction force acting on a single tire while accelerating, $J_w$ is the moment of inertia of the wheel assembly, $r$ is the radius of the tire, $T_m$ is the torque from the motor and $\alpha_w$ is the angular acceleration of the wheel.
+
+Rearranging the above three equations, the acceleration of the vehicle can be expressed as a function of motor torque as follows: 
+\begin{equation}
+\begin{split}
+a &= \frac{T_m r}{\frac{1}{2}Mr^2 + J_w} \\
+a &= \frac{(\SI{135}{\newton\metre})(\SI{0.27}{\metre})}{\frac{1}{2}(\SI{550}{\kilo\gram})(\SI{0.27}{\metre\squared})^2 + \SI{0.25}{\kilo\gram\metre\squared}} = \SI{1.80}{\metre\per\second\squared}
+\end{split}
+\end{equation}
+Solving for friction force on a single rear tire: 
+\begin{equation}
+F_w = Ma = (\SI{550}{\kilo\gram})(\SI{1.80}{\metre\per\second\squared}) = \SI{990}{\newton}
+\end{equation}
+The resulting normal force on the rear wheels can be calculated as follows: 
+\begin{equation}
+\begin{split}
+bF_\mathrm{Nr} &= hF_\mathrm{w} + a(Mg - F_\mathrm{Nr})\\
+F_\mathrm{Nr} &= \frac{hF_\mathrm{w} + aMg}{a+b} = \frac{(\SI{0.52}{\metre})(\SI{990}{\newton})+(\SI{1.32}{\metre})(\SI{550}{\kilo\gram})\left(\SI{9.81}{\metre\per\second\squared}\right)}{(\SI{1.32}{\metre})+(\SI{1.28}{\metre})} = \SI{2.94}{\kilo\newton}
+\end{split}
+\end{equation}
+Note that \SI{2.94}{\kilo\newton} is for both rear wheels. Therefore there is approximately \SI{1.47}{\kilo\newton} of normal force on each rear tire during maximum acceleration.
+
+\subsection{Cornering}
+ASC regulations require simulations of a 1G turn, at minimum. This corresponds to the following lateral force:
+\begin{equation}
+F_\mathrm{steer} = Mg = (\SI{550}{\kilo\gram})(\SI{9.81}{\metre\per\second\squared}) = \SI{5.40}{\kilo\newton}
+\end{equation}
+
+Additionally, the diameter of several highway exit ramps around the University of Waterloo were measured on satellite maps.
+The recommended speed limit on these exits is \SI{50}{\kilo\metre\per\hour}.
+Assuming a worst case scenario where the vehicle takes an exit with a \SI{100}{\metre} diameter at \SI{80}{\kilo\metre\per\hour} (\SI{22.22}{\metre\per\second}), the required lateral force required to navigate this corner can be found as follows:
+\begin{equation}
+F_\mathrm{steer} = \frac{Mv^2}{r} = \frac{(\SI{550}{\kilo\gram})(\SI{22.22}{\metre\per\second})^2}{\SI{50}{\metre}} = \SI{5.43}{\kilo\newton}
+\end{equation}
+
+This corresponds nearly identically to a 1G turn. Since the lateral force required in the second scenario is slightly higher, \SI{5.34}{\kilo\newton} will be used as the worst case cornering force.
+
+\begin{figure}
+\centering
+\includegraphics[width=.5\textwidth]{./figures/cornering-fbd.jpg}
+\caption{Free Body Diagram of a Car Cornering}
+\label{fig:cornering-fbd}
+\end{figure}
+
+Using the half car model shown if Figure \ref{fig:cornering-fbd}, summing the moments about the centre of gravity and the vertical forces results in the following equations: 
+\begin{gather}
+\frac{1}{2}L_\mathrm{track}F_\mathrm{No} = hF_\mathrm{steer} + \frac{1}{2}L_\mathrm{track}F_\mathrm{Ni} \\
+F_\mathrm{No} + F_\mathrm{Ni} = Mg
+\end{gather}
+Solving for the outside wheel normal force: 
+\begin{equation}
+\begin{split}
+\frac{1}{2}L_\mathrm{track}F_\mathrm{No} &= hF_\mathrm{steer} + \frac{1}{2}L_\mathrm{track}(Mg - F_\mathrm{Ni})\\
+F_\mathrm{No} &= \frac{hF_\mathrm{steer} + \frac{1}{2}L_\mathrm{track}Mg}{L_\mathrm{track}} = \frac{(\SI{0.52}{\metre})(\SI{5.43}{\kilo\newton})+(\SI{0.80}{\metre})(\SI{550}{\kilo\gram})\left(\SI{9.81}{\metre\per\second\squared}\right)}{\SI{1.60}{\metre}} = \SI{4.46}{\kilo\newton}
+\end{split}
+\end{equation}
+Solving for the inside wheel normal force: 
+\begin{equation}
+F_\mathrm{Ni} = Mg - F_\mathrm{No} = (\SI{550}{\kilo\gram})\left(\SI{9.81}{\metre\per\second\squared}\right) - \SI{4.46}{\kilo\newton} = \SI{0.93}{\kilo\newton}
+\end{equation}
+Note that the forces above are for both front and rear outside wheels.
+Assuming the force is evenly split between the front and rear, there is approximately \SI{2.23}{\kilo\newton} of normal force on each outside tire and \SI{0.47}{\kilo\newton} for each inside tire during maximum cornering.
+
+\subsection{Bump}
+As stipulated in the ASC 2018 regulations, the analysis for suspension and steering must include loading considerations for a 2G bump.
+We interpret this requirement as the body of the vehicle accelerating upwards at 2G, in which case the additional normal force subjected to a single tire can be expressed as follows:
+
+\begin{equation}
+F_\mathrm{N,bump} = \frac{2Mg}{4} = \frac{2(\SI{9.81}{\metre\per\second\squared})(\SI{550}{\kilo\gram})}{4} = \SI{2.70}{\kilo\newton}
+\end{equation}
+
+\subsection{Superposition}
+The braking, accelerating, and cornering conditions subject the wheels to normal forces that differ from those experienced when the car is at rest.
+The change in normal force for each loading condition can be expressed as follows:
+\begin{align}
+\Delta F_\mathrm{Nf,braking}      &= F_\mathrm{Nf,braking}   - \frac{Mg}{4} =  \SI{1.87}{\kilo\newton} - \SI{1.35}{\kilo\newton} =  \SI{0.52}{\kilo\newton} \\
+\Delta F_\mathrm{Nr,accelerating} &= F_\mathrm{Nf,braking}   - \frac{Mg}{4} =  \SI{1.47}{\kilo\newton} - \SI{1.35}{\kilo\newton} =  \SI{0.12}{\kilo\newton} \\
+\Delta F_\mathrm{No,cornering}    &= F_\mathrm{No,cornering} - \frac{Mg}{4} =  \SI{2.23}{\kilo\newton} - \SI{1.35}{\kilo\newton} =  \SI{0.88}{\kilo\newton} \\
+\Delta F_\mathrm{Ni,cornering}    &= F_\mathrm{No,cornering} - \frac{Mg}{4} =  \SI{0.47}{\kilo\newton} - \SI{1.35}{\kilo\newton} = -\SI{0.88}{\kilo\newton} \\
+\Delta F_\mathrm{N,bump}          &= F_\mathrm{N,bump}       - \frac{Mg}{4} =  \SI{2.70}{\kilo\newton} - \SI{1.35}{\kilo\newton} =  \SI{1.35}{\kilo\newton}
+\end{align}
+To approximate the worst case loading condition, a superposition of the changes in normal force and the nominal normal force is applied.
+Note that the nominal force used is the same for each wheel, which assumes a centre of gravity approximately centred along the wheelbase.
+\begin{align}
+F_\mathrm{Nfi,max} &= \Delta F_\mathrm{Nf,braking} + \Delta F_\mathrm{Ni,cornering} + 2F_\mathrm{N,nominal} = \SI{0.52}{\kilo\newton} - \SI{0.88}{\kilo\newton} + \SI{2.70}{\kilo\newton} = \SI{2.34}{\kilo\newton} \\
+F_\mathrm{Nfo,max} &= \Delta F_\mathrm{Nf,braking} + \Delta F_\mathrm{No,cornering} + 2F_\mathrm{N,nominal} = \SI{0.52}{\kilo\newton} + \SI{0.88}{\kilo\newton} + \SI{2.70}{\kilo\newton} = \SI{4.10}{\kilo\newton} \\
+F_\mathrm{Nri,max} &= \Delta F_\mathrm{Nr,braking} + \Delta F_\mathrm{Ni,cornering} + 2F_\mathrm{N,nominal} = \SI{0.12}{\kilo\newton} - \SI{0.88}{\kilo\newton} + \SI{2.70}{\kilo\newton} = \SI{1.94}{\kilo\newton} \\
+F_\mathrm{Nro,max} &= \Delta F_\mathrm{Nr,braking} + \Delta F_\mathrm{No,cornering} + 2F_\mathrm{N,nominal} = \SI{0.12}{\kilo\newton} + \SI{0.88}{\kilo\newton} + \SI{2.70}{\kilo\newton} = \SI{3.70}{\kilo\newton}
+\end{align}
+
+\subsection{Summary}
+Table \ref{tab:worst-case-loading} summarizes the loading conditions used for static structural FEA simulations on the suspension and steering components.
+Note that the direction of the force corresponds's with the global coordinate system of the MSXII and the vectors shown in Figure \ref{fig:suspension-coordinate-system}.
+The assemblies shown are for the driver's side of the vehicle
+
+\begin{table}[h!]
+\centering
+\caption{Summary of worst case loading conditions}
+\label{tab:worst-case-loading}
+\begin{tabular}{lrrr}
+\toprule
+Worst Case Loading                             & $F_x$ (\si{\kilo\newton}) & $F_y$ (\si{\kilo\newton}) & $F_z$ (\si{\kilo\newton}) \\
+\midrule
+Braking \& Cornering (Inside Front Wheel)      & 2.72                     & 2.34                     & 2.70                     \\
+Braking \& Cornering (Outside Front Wheel)     & -2.72                    & 4.10                     & 2.70                     \\
+Accelerating \& Cornering (Inside Rear Wheel)  & 2.72                     & 1.94                     & 0.99                     \\
+Accelerating \& Cornering (Outside Rear Wheel) & -2.72                    & 3.70                     & 0.99                     \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{figure}
+\centering
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/front-axis-directions}
+\caption{Font suspension}
+\end{subfigure}
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/rear-axis-directions}
+\caption{Rear suspenion}
+\end{subfigure}
+\caption{Suspension coordinate system}
+\label{fig:suspension-coordinate-system}
+\end{figure}
+
+\clearpage
+\section{FEA images for chassis}
+\label{sec:fea-chassis-images}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/fea/chassis/chassis-collision-bumper-front}
+\caption{Front bumper collision}
+\label{sec:chassis-collision-bumper-front}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/fea/chassis/chassis-collision-bumper-30deg}
+\caption{Front bumper 30 degree collision}
+\label{sec:chassis-collision-bumper-30deg}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/fea/chassis/chassis-collision-bumper-60deg}
+\caption{Front bumper 60 degree collision}
+\label{sec:chassis-collision-bumper-60deg}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/fea/chassis/chassis-collision-bumper-side}
+\caption{Side bumper collision}
+\label{sec:chassis-collision-bumper-side}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/fea/chassis/chassis-collision-bumper-rear}
+\caption{Rear bumper collision}
+\label{sec:chassis-collision-bumper-rear}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/fea/chassis/chassis-collision-hood-60deg}
+\caption{Hood 60 degree collision}
+\label{sec:chassis-collision-hood-60deg}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/fea/chassis/chassis-collision-hood-30deg}
+\caption{Hood 30 degree collision}
+\label{sec:chassis-collision-hood-30deg}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/fea/chassis/chassis-collision-side-60deg}
+\caption{Side 60 degree collision}
+\label{sec:chassis-collision-side-60deg}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/fea/chassis/chassis-collision-side-30deg}
+\caption{Side 30 degree collision}
+\label{sec:chassis-collision-side-30deg}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/fea/chassis/chassis-collision-top}
+\caption{Roof top collision}
+\label{sec:chassis-collision-top}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/fea/chassis/chassis-front-suspension-rest-loading}
+\caption{Vehicle rest front subframe loading}
+\label{sec:chassis-front-suspension-rest-loading}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/fea/chassis/chassis-front-suspension-stress}
+\caption{Vehicle rest front subframe stress}
+\label{sec:chassis-front-suspension-stress}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/fea/chassis/chassis-rear-suspension-stress}
+\caption{Vehicle rest rear subframe stress}
+\label{sec:chassis-rear-suspension-stress}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/fea/chassis/torsion-deformation}
+\caption{Torsion test, deformation}
+\label{sec:torsion-deformation}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.5\textwidth]{figures/fea/chassis/torsion-stress}
+\caption{Torsion test, stress}
+\label{sec:torsion-stress}
+\end{figure}
+
+\clearpage
+\section{FEA images for suspension and steering}
+\label{sec:fea-parts-images}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00015-CoiloverClevis-Setup}
+\caption{Setup}
+\end{subfigure}
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00015-CoiloverClevis-Stress}
+\caption{Stress}
+\end{subfigure}
+\caption{Coilover clevis}
+\label{fig:MS00015-CoiloverClevis}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00017-UpperControlArm-Setup}
+\caption{Setup}
+\end{subfigure}
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00017-UpperControlArm-Stress}
+\caption{Stress}
+\end{subfigure}
+\caption{Upper control arm}
+\label{fig:MS00017-UpperControlArm}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00018-LowerControlArm-Setup}
+\caption{Setup}
+\end{subfigure}
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00018-LowerControlArm-Stress}
+\caption{Stress}
+\end{subfigure}
+\caption{Lower control arm}
+\label{fig:MS00018-LowerControlArm}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00019-Upright-Setup}
+\caption{Setup}
+\end{subfigure}
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00019-Upright-Stress}
+\caption{Stress}
+\end{subfigure}
+\caption{Upright}
+\label{fig:MS00019-Upright}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00020-Spindle-Setup}
+\caption{Setup}
+\end{subfigure}
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00020-Spindle-Stress}
+\caption{Stress}
+\end{subfigure}
+\caption{Spindle}
+\label{fig:MS00020-Spindle}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00029-UpperControlArmClevis-Setup}
+\caption{Setup}
+\end{subfigure}
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00029-UpperControlArmClevis-Stress}
+\caption{Stress}
+\end{subfigure}
+\caption{Upper control arm clevis}
+\label{fig:MS00029-UpperControlArmClevis}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00030-UpperBearingPlate-Setup}
+\caption{Setup}
+\end{subfigure}
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00030-UpperBearingPlate-Stress}
+\caption{Stress}
+\end{subfigure}
+\caption{Upper bearing plate}
+\label{fig:MS00030-UpperBearingPlate}
+\end{figure}
+
+%\begin{figure}[H]
+%\centering
+%\begin{subfigure}[b]{.48\textwidth}
+%\centering
+%\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00036-CoiloverControlArmTab-Setup}
+%\caption{Setup}
+%\end{subfigure}
+%\begin{subfigure}[b]{.48\textwidth}
+%\centering
+%\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00036-CoiloverControlArmTab-Stress}
+%\caption{Stress}
+%\end{subfigure}
+%\caption{Coilover control arm tab}
+%\label{fig:MS00036-CoiloverControlArmTab}
+%\end{figure}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00038-CoiloverMountingPlate-Setup}
+\caption{Setup}
+\end{subfigure}
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00038-CoiloverMountingPlate-Stress}
+\caption{Stress}
+\end{subfigure}
+\caption{Coilover mounting plate}
+\label{fig:MS00038-CoiloverMountingPlate}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00047-RearSuspensionCoiloverTab-Setup}
+\caption{Setup}
+\end{subfigure}
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00047-RearSuspensionCoiloverTab-Stress}
+\caption{Stress}
+\end{subfigure}
+\caption{Rear suspension coilover tab}
+\label{fig:MS00047-RearSuspensionCoiloverTab}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00074-SteeringArm-Setup}
+\caption{Setup}
+\end{subfigure}
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00074-SteeringArm-Stress}
+\caption{Stress}
+\end{subfigure}
+\caption{Steering arm}
+\label{fig:MS00074-SteeringArm}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00086-RearSuspensionClevisMount-Setup}
+\caption{Setup}
+\end{subfigure}
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00086-RearSuspensionClevisMount-Stress}
+\caption{Stress}
+\end{subfigure}
+\caption{Rear suspension clevis}
+\label{fig:MS00086-RearSuspensionClevisMount}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00101-TieRod-Setup}
+\caption{Setup}
+\end{subfigure}
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00101-TieRod-Stress}
+\caption{Stress}
+\end{subfigure}
+\caption{Tie rod}
+\label{fig:MS00101-TieRod}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00102-LowerControlArmClevis-Setup}
+\caption{Setup}
+\end{subfigure}
+\begin{subfigure}[b]{.48\textwidth}
+\centering
+\includegraphics[width=0.9\textwidth]{figures/fea/parts/MS00102-LowerControlArmClevis-Stress}
+\caption{Stress}
+\end{subfigure}
+\caption{Lower control arm clevis}
+\label{fig:MS00102-LowerControlArmClevis}
+\end{figure}
+
 \end{document}

### 2018-10-30 Brakes/forwardDynamicsBrakingCalculations.m
Moved old projects to MSXII folder

@@ -1,126 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Longitudinal Weight Balance - longWeightBalance.m
-% Script to compute the longitudinal weight balance of the car during
-% braking or acceleration.
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2017
-%
-%
-% Notes:
-%   
-%   - Units: 
-%       Distance: meters
-%       Mass: kilograms
-%       Time: seconds
-%       Angles: degrees
-%
-%   - All force vectors are of the form [front, rear]
-%
-%   - Sliding condiditons: 
-%       - 0: No sliding
-%       - 1: Front sliding
-%       - 2: Rear sliding
-%       - 3: both sliding
-%
-% ---------------------------------------------------------------------- %
-
-close all
-clear all
-
-% --------------------------- Constants ------------------------------- %
-
-m = 550;            % car mass in kg
-cogZ = 1.52;         % Center of gravity position in longitudinal direction (wrt global origin)
-cogY = 0.52;         % Center of gravity position in vertical direction (wrt global origin)
-wheelbase = 2.6;   % Wheelbase in m
-g = -9.81;          % Acceleration of gravity (m/s^2)
-muBrakePad = [0.35, 0.4];  % Coefficient of friction between the tire and road [dynamic, static]
-muRoad = [0.6, 0.7];       % Coefficient of friction between the tire and road [dynamic, static]
-dynamic = 1; % Constant to increase code readabliltiy 
-static = 2; % Constant to increase code readabliltiy 
-wheelDiam = 0.533;  % Wheel Diameter in m
-rotorDiam = [.18, .3];    % Rotor diameter at center of brake pad [front, rear]
-padContactArea = 0.00129; % Caliper Contact area in m^2 http://www.wilwood.com/BrakePads/BrakePadsProd.aspx?itemno=150-4091K
-caliperCylinderArea = 0.001; % Area of caliper cylinder 
-masterCylinderArea = 0.001; % Area of master cylinder 
-brakePedalReduction = 3; % 3:1 mechanical advantage on brake pedal
-
-% --------------------------- Variables ------------------------------- %
-
-frontPressureRange = 5e5:5e5:20e6;
-pressureRatioRange = 0:.05:1;
-
-[frontPressure, pressureRatio] = meshgrid(frontPressureRange, pressureRatioRange);
-
[…]

### 2018-10-30 Brakes/inverseDynamicsBrakingCalculations.m

@@ -1,74 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Longitudinal Weight Balance - longWeightBalance.m
-% Script to compute the longitudinal weight balance of the car during
-% braking or acceleration.
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2017
-%
-%
-% Notes:
-%   
-%   - Units: 
-%       Distance: meters
-%       Mass: kilograms
-%       Time: seconds
-%       Angles: degrees
-%
-% ---------------------------------------------------------------------- %
-
-close all
-clear all
-
-% --------------------------- Constants ------------------------------- %
-
-m = 611;            % car mass in kg
-cogZ = 0.8;         % Center of gravity position in longitudinal direction
-cogY = 0.5;         % Center of gravity position in vertical direction
-wheelbase = 1.78;   % Wheelbase in m
-g = -9.81;          % Acceleration of gravity (m/s^2)
-muBrakePad = 0.35;  % Coefficient of friction between the tire and road
-muRoad = 0.6;       % Coefficient of friction between the tire and road
-
-a = 5;              % Target acceleration
-frontWeightBalance = .65; % Proportion of car's weight on front wheels druing hard
-                          % braking
-wheelDiam = 0.533;  % Wheel Diameter in m
-rotorDiam = .18;    % Rotor diameter at center of brake pad
-padContactArea = 0.00129 % Caliper Contact area in m^2 http://www.wilwood.com/BrakePads/BrakePadsProd.aspx?itemno=150-4091K
-caliperCylinderArea = 0.001; % Area of caliper cylinder 
-masterCylinderArea = 0.001; % Area of master cylinder 
-brakePedalReduction = 3; % 3:1 mechanical advantage on brake pedal
-
-% -------------------- Solving for Braking forces --------------------- %
-
-% The dynamic of braking can be modeled by four linearly inpendnet
-% equations: 
-%       1) Sum of forces in the x = mass * car acceleration 
-%       2) Sum of forces in the y = 0
-%       3) Sum of the moments about the center of gravity = 0
-%       4) The front braking force will be at max mu*normal force
-% Expressing the following system of equations in maxtrix form yeilds Af=b
[…]

### 2018-10-30 Brakes/roadReactionForce.m

@@ -1,80 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Function that returns the brake force based for the line pressure, pad
-% area, rotor diameter, and coefficient of friction
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2017
-%
-%
-% Notes:
-%   
-%   - Units: 
-%       Distance: meters
-%       Mass: kilograms
-%       Time: seconds
-%       Angles: degrees
-%       Pressure: pascals
-%
-% The dynamics of braking can be modeled by three linearly inpendnet
-% equations: 
-%       1) Sum of forces in the x = mass * car acceleration 
-%       2) Sum of forces in the y = 0
-%       3) Sum of the moments about the center of gravity = 0
-% Expressing the following system of equations in maxtrix form yeilds Af=b
-% where f is a vector containing: acceleration, a front brake force and 
-% rear brake force (in that order)
-%
-% ---------------------------------------------------------------------- %
-
-function [force, accel] = roadReactionForce(dynamicMuRoad, sliding, brakingForce, wheelbase, cogZ, cogY, m, g)
-    if(sliding == 0)
-        % Setup A matrix
-        A = [m/2 0 0;
-             0 1 1;
-             0 -cogZ (wheelbase-cogZ)]; 
-        % Setup b vector 
-        b = [-(brakingForce(1)+brakingForce(2)); 
-             -m*g/2; 
-             -(brakingForce(1)*cogY+brakingForce(2)*cogY)];
-         
-    elseif(sliding == 1)
-        % Setup A matrix
-        A = [m/2 dynamicMuRoad 0;
-             0 1 1;
-             0 (-cogZ+cogY*dynamicMuRoad) (wheelbase-cogZ)]; 
-        % Setup b vector 
-        b = [-brakingForce(2); 
-             -m*g/2; 
-             -brakingForce(2)*cogY];
-         
-    elseif(sliding == 2)
-        % Setup A matrix
-        A = [m/2 0 dynamicMuRoad;
-             0 1 1;
-             0 -cogZ (wheelbase-cogZ+cogY*dynamicMuRoad)]; 
-        % Setup b vector 
-        b = [-brakingForce(1); 
-             -m*g/2; 
-             -brakingForce(1)*cogY];
-         
-    elseif(sliding == 3)
-        % Setup A matrix
[…]

### 2018-10-30 Brakes/staticBrakingForce.m

@@ -1,25 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Function that returns the brake force based for the line pressure, pad
-% area, rotor diameter, and coefficient of friction
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2017
-%
-%
-% Notes:
-%   
-%   - Units: 
-%       Distance: meters
-%       Mass: kilograms
-%       Time: seconds
-%       Angles: degrees
-%       Pressure: pascals
-%
-% ---------------------------------------------------------------------- %
-
-function force = staticBrakingForce(pressure, mu, area, rotorDiam, tireDiam)
-    force = pressure.*area.*mu.*rotorDiam./tireDiam;
-end
-

### 2018-10-30 FEABoundaryConditions/matlab/Old/createTransferFunctions.m

@@ -1,269 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Suspension Spring Rates Project - createTransferFuncitons.m
-% Function setup transfer functions for a half car model to determine the
-% reponse of MSXII to disturbances in the road conditions
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2017
-%
-%
-% Notes:
-%   
-%   - 
-% ---------------------------------------------------------------------- %
-
-load_system('halfCarModel');
-close all;
-
-% ----------------------------- Constants ------------------------------ %
-
-c1_ = 9670;     % Front damping coefficient (Ns/m)
-c2_ = 9670;     % Rear damping coefficient (Ns/m) 
-k1_ = 650*175.1;    % Front sring rate (N/m) 
-k2_ = 550*175.1;    % Rear sring rate (N/m)
-m_ = 550;       % Mass of the vehicle (kg)
-I_ = 550;       % Moment of interial about global x axis (kg*m^2)
-a_ = 1.3;       % Distance from COG to front tire (m)
-b_ = 1.3;       % Distance from COG to rear tire (m)
-
-% ---------------------- Forcing Function Parameters ------------------- %
-
-amplitude = 0.015;              % Amplitude of bumps on road surface
-bumpSparation = 2;              % Distance between peaks on the road surface (m)
-speed = 22.2;                   % Cruising speed (m/s)
-wb = 2*pi*speed/bumpSparation;  % Frequency of base excitation (rad/sec)
-timeDelay = (a_+b_)/speed;        % delay between input striking front and
-                                % rear wheels (s)
-                                
-% Update simulink parameters
-% set_param('halfCarModel/Road','Frequency',mat2str(wb),'Amplitude',mat2str(amplitude));
-set_param('halfCarModel/DelayA','DelayTime',mat2str(timeDelay));
-set_param('halfCarModel/DelayB','DelayTime',mat2str(timeDelay));
-                              
[…]

### 2018-10-30 FEABoundaryConditions/matlab/alternativeFormulation.m

@@ -1,155 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Suspension Spring Rates Project - alternativeFormulation.m
-% Function to setup transfer functions for a half car model to determine the
-% reponse of MSXII to disturbances in the road conditions
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2017
-%
-%
-% Notes:
-%   
-%   - 
-% ---------------------------------------------------------------------- %
-
-load_system('halfCarModel');
-close all;
-clear all
-
-% ----------------------------- Constants ------------------------------ %
-
-c1_ = 4830;     % Front damping coefficient (Ns/m)
-c2_ = 4830;     % Rear damping coefficient (Ns/m) 
-k1_ = 600*175.1*cosd(35);    % Front sring rate (N/m) 
-k2_ = 500*175.1*cosd(30);    % Rear sring rate (N/m)
-m_ = 550;       % Mass of the vehicle (kg)
-I_ = 550;       % Moment of interial about global x axis (kg*m^2)
-a_ = 1.32;       % Distance from COG to front tire (m)
-b_ = 1.28;       % Distance from COG to rear tire (m)
-
-% ---------------------- Forcing Function Parameters ------------------- %
-
-amplitude = 0.015;              % Amplitude of bumps on road surface
-bumpSparation = 2;              % Distance between peaks on the road surface (m)
-speed = 22.2;                   % Cruising speed (m/s)
-wb = 2*pi*speed/bumpSparation;  % Frequency of base excitation (rad/sec)
-timeDelay = (a_+b_)/speed;      % delay between input striking front and
-                                % rear wheels (s)
-                                
-% Update simulink parameters
-set_param('halfCarModel/DelayA','DelayTime',mat2str(timeDelay));
-set_param('halfCarModel/DelayB','DelayTime',mat2str(timeDelay));
-                              
-% ----------------- Create Symbolic Transfer Functions ----------------- %
-
[…]

### 2018-10-30 FEABoundaryConditions/matlab/simplifiedFormulation.m

@@ -1,91 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Suspension Spring Rates Project - simplifiedFormulation.m
-% Function to setup simplified transfer functions for a half car model to 
-% determine the reponse of MSXII to disturbances in the road conditions
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2017
-%
-%
-% Notes:
-%   
-%   - 
-% ---------------------------------------------------------------------- %
-
-close all;
-clear all;
-clc;
-
-% ----------------------------- Constants ------------------------------ %
-
-c_ = 4830;     % Front damping coefficient (Ns/m)
-k_ = 88600;    % Front sring rate (N/m) 
-M_ = 550;      % Mass of the vehicle (kg)
-I_ = 550;      % Moment of interial about global x axis (kg*m^2)
-a_ = 1.3;      % Distance from COG to front tire and rear tire (m)
-zx_ = 0.49;     % Damping ratio of translational  
-wnx_ = 17.95;   % Natural frequency of pitch
-zth_ = 0.64;    % Damping ratio of pitch
-wnth_ = 23.33;  % Natural frequency of pitch
-
-
-% ---------------------- Forcing Function Parameters ------------------- %
-
-A_ = 0.015;                     % Amplitude of bumps on road surface
-bumpSparation = 2;              % Distance between peaks on the road surface (m)
-speed = 22.2;                   % Cruising speed (m/s)
-wb_ = 2*pi*speed/bumpSparation; % Frequency of base excitation (rad/sec)
-t0_ = (2*a_)/speed;            % delay between input striking front and
-                                % rear wheels (s)
-
-% ----------------- Create Symbolic Transfer Functions ----------------- %                                
-                                
-syms t s M I a t0 A wb c k zx wnx zth wnth
-
-QA = s^2 + 2*c/M*s + wnx^2; % ODE A 
-QB = s^2 + 2*a^2*c/I*s + wnx^2; % ODE B
-
-phaseShift = exp(-s*t0);
-
[…]

### 2018-10-30 Handling Model/handling_revised_4.m

@@ -1,132 +1,0 @@
-close all
-clear all
-
-%=============================== Constants ===============================%
-% Most of the contants are extracted from Ubaidillah et. al. 2015 
-% Constants at Line 73 of main.m and Line 17 of tireModel.m are completely
-% made up, as noted. Any suggestions welcome for proper (or even acceptable approximations)
-% of these values would be appreciated. 
-
-totalMass = 1600;
-sprungMass = 1400; % kgs
-Iyy = 1263.5; % kgm^2
-Ixx = 450.1;
-Izz = 1263.5;
-mufl = 40; murl = 35; murr = 35; mufr = 40;
-K = [18000, 18000, 18000, 18000]; % K values at each corner (see correspondences below)
-C = [1500, 1500, 1500, 1500]; % C values at each corner (see correspondences below)
-
-% Model iterates over each corner (1:4) of the car with the following
-% correspondences:
-% 1 -> front left
-% 2 -> front right
-% 3 -> rear left
-% 4 -> rear right
-
-zRest = (sprungMass*9.81)/(4*18000); % this is the natural deltaZ position
-
-lf = 1.15; lr = 1.15;
-hcg = 0.85; w = 1.3;
-wheelRadius = 0.3; 
-xDistance = [1.15, 1.15, 1.15, 1.15]; 
-yDistance = 1; 
-timeMax = 5; IWheel = 7;
-
-%================================ Inputs =================================%
-TMotor = [5, 5, 0, 0, 0, 0, 0]; % Motor torque at each time 
-TBraking = [0, 0, 0, 0, 0, 0, 0]; % Braking torque at each time
-delta = [0, 0, 0, 0, 0, 0, 0]; % Wheel angle at each time 
-%delta = [0, 0, 0, 0, 0, 0, 0]; % Wheel angle at each time 
-timeStep = 0.5;
-
-for time = 1:7
-     if time == 1
-        
-        for i = 1:4
-            FLat(time,i) = 0; Fdrive(time,i) = 0; 
-            Fx(time,i) = 0; Fy(time,i) = 0; Fz(time,i) = sprungMass*9.81/4;
-            deltaZ(time,i) = zRest; deltaZPrime(time,i) = 0;
-        end
[…]

### 2018-10-30 Handling Model/tireModelY.m

@@ -1,47 +1,0 @@
-function Fy = tireModelY(vx, vy, lf, r, w, deltaf, Fz)
-
-% for i = 1:2
-%     if i == 1
-%         a(i,1) = atan2d((vy+lf*r),vx)-deltaf;
-%         vt(i,1) = sqrt((vy+lf*r)^2+vx^2);
-%         vwx(i,1) = vt(i,1)*cosd(a(i,1));
-%     else
-%         a(i,1) = atan2d((vy+lf*r),vx);
-% 
-%     end
-
-      a(1,1) = deltaf - atan2d((vy+lf*r),(vx-(w*r/2))); 
-      a(2,1) = deltaf - atan2d((vy+lf*r),(vx+(w*r/2)));
-      a(3,1) = -atan2d((vy-lf*r),(vx-(w*r/2)));
-      a(4,1) = -atan2d((vy-lf*r),(vx+(w*r/2))); 
-      
-%       vt(i,1) = sqrt((vy+lf*r)^2+vx^2);
-%       vwx(i,1) = vt(i,1)*cosd(a(i,1));
-%     wheelRadius = 0.3; 
-%     angularVelocity(i,1) = vwx(i,1)/wheelRadius;
-%     longSlipRatio(i,1) = (vwx(i,1) - angularVelocity(i,1)*wheelRadius)/vwx(i,1);
-%     
-    CorneringStiffness = 100; WheelStiffness = 150; mu = 0.45; % completely made-up
-    
-    
-%     a(1,1) = deltaf - atand((vy+lf*r)/(vx-(w*r/2))); 
-%     a(2,1) = deltaf - atand((vy+lf*r)/(vx+(w*r/2)));
-%     a(3,1) = -atand((vy-lf*r)/(vx-(w*r/2)));
-%     a(4,1) = -atand((vy-lf*r)/(vx+(w*r/2))); 
-   
-for i = 1:4
-    Lambda(i,1) = (mu*Fz)/(2*(CorneringStiffness*abs(tand(a(i,1)))));
-    %Lambda(i,1) = (mu*Fz+(1+longSlipRatio(i,1)))/(2*sqrt((WheelStiffness*longSlipRatio(i,1))^2 + ...
-     %   (CorneringStiffness*tand(a(i,1)))^2));
-    
-    if Lambda < 1
-        fLambda(i,1) = (2-Lambda(i,1))*Lambda(i,1);
-    else
[…]

### 2018-10-30 Handling Model/torqueModel.m

@@ -1,8 +1,0 @@
-function Fdrive = torqueModel(TMotor, TBraking, Inertia, a, radius)
-
-angAccel = a/radius; 
-%Fy = (Inertia*angAccel + TMotor - TBraking)/radius; 
-Fdrive = (Inertia*angAccel - TMotor + TBraking)/radius; 
-
-
-    

### 2018-10-30 HighLevelCarModel/archive/computeDrag.m

@@ -1,23 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% High Level Car Model - computeDrag.m
-% Function to compute drag force 
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2017
-%
-%
-% Notes:
-%   
-%   - Units: 
-%       Distance: meters
-%       Mass: kilograms
-%       Time: seconds
-%       Angles: degrees
-%
-% ---------------------------------------------------------------------- %
-
-function drag = computeDrag(rho, v, Cd, A)
-    drag = 0.5.*rho.*(v.^2).*Cd.*A;
-end

### 2018-10-30 HighLevelCarModel/archive/debug.m

@@ -1,4 +1,0 @@
-clc
-clear all
-close all
-rungeKutta4(@test,0,1,1)

### 2018-10-30 HighLevelCarModel/archive/rungeKutta4.m

@@ -1,23 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% High Level Car Model - rungeKutta4.m
-% Performs one iteration of 4th order runge kutta
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2018
-%
-% Notes:
-%   
-%   - The parameter fcn is a differential equation of the 
-%     form y' = fcn(t,y)
-%
-% ---------------------------------------------------------------------- %
-
-function yNext = rungeKutta4(fcn, t, deltat, y)
-    k1 = deltat * fcn(t, y);
-    k2 = deltat * fcn(t+deltat/2, y+k1/2);
-    k3 = deltat * fcn(t+deltat/2, y+k2/2);
-    k4 = deltat * fcn(t+deltat, y+k3);
-    yNext = y + (k1 + 2*k2 + 2*k3 + k4)/6; 
-end

### 2018-10-30 HighLevelCarModel/archive/test.m

@@ -1,3 +1,0 @@
-function testOutput = test(t, y)
-    testOutput = -1/(1+y^2);
-end

### 2018-10-30 HighLevelCarModel/test.py

@@ -1,77 +1,0 @@
-from vehicle import *
-import vehicle_data
-
-"""Trig Functions:"""
-print
-print "Trig Functions:"
-print "sin(30) = ", sind(30)
-print "cos(60) = ", sind(60)
-print "tan(45) = ", tand(45)
-print "asin(0.5) = ", asind(0.5)
-print "acos(0.5) = ", asind(0.5)
-print "atan(1) = ", atand(1)
-
-"""Vehicle Constants Class:"""
-print
-print "Vehicle Constant Class:"
-msxii_data = VehicleConstants()
-msxii_data.load_vehicle(vehicle_data.msxii)
-print "Wheel diameter = ", msxii_data.diam_wheel
-
-"""Coordiante Class:"""
-print
-print "Coordiante Class:"
-test_coord1 = Coordiante(37.455926, -122.065457)
-test_coord2 = Coordiante(37.509186, -122.060822)
-test_coord3 = Coordiante(56.317013, -117.465206)
-test_coord4 = Coordiante(56.335315, -117.506920)
-print "Test Case 1:"
-print test_coord1.get_distance_to(test_coord2)
-print test_coord1.get_direction_to(test_coord2)
-print "Test Case 2:"
-print test_coord3.get_distance_to(test_coord4)
-print test_coord3.get_direction_to(test_coord4)
-print "Test Case 3:"
-print test_coord1.get_distance_to(test_coord4)
-print test_coord1.get_direction_to(test_coord4)
-
-"""Battery Class:"""
-print
-print "Battery Class:"
-test_battery = Battery(1)
-test_battery.soc = 1000
-motor_power = [50,100,-100,-100,100] #Integrate: 25, 75, 0, -100, 0
-rear_array_power = [50, 50, 50, 50, 50] # Integrate: 25, 50, 50, 50, 50
-front_array_power = [50, 50, 50, 50, 50] # Integrate: 25, 50, 50, 50, 50
-lv_losses = [0,0,0,0,0]
-hv_losses = [0,0,0,0,0]
-for i in range(len(motor_power)):
-    test_battery.update(motor_power[i], rear_array_power[i],
-                        front_array_power[i], lv_losses[i], hv_losses[i])
-    print "Net Power = ", test_battery.net_power
-print "Final SOC = ", test_battery.soc #Total 75, 175, 100, 0, 100 -> final SOC = 1450
-
[…]

### 2018-10-30 HighLevelCarModel/vehicle.py

@@ -1,596 +1,0 @@
-#!/usr/bin/env python
-
-""" All vehcile related classes
-
-Module contains all classes related to the vehicle as a whole, vehicle
-parameters.
-
-All units are as follows unless otherwise specified:
-    Distance: meters (m)
-    Mass: kilograms (kg)
-    Time: seconds (s)
-    Angles: degrees (deg)
-    Temperature: degrees celcius (degC)
-
-Midnight Sun Solar Car Team 2018
-University of Waterloo
-"""
-
-__author__ = "Devon Copeland"
-__copyright__ = "Copyright 2018, Midnight Sun Solar Car Team"
-
-from datetime import datetime, timedelta
-import math
-from numpy import sign
-
-"""Trig functions implemented for degrees"""
-def cosd(deg):
-    return math.cos(math.radians(deg))
-
-def sind(deg):
-    return math.sin(math.radians(deg))
-
-def tand(deg):
-    return math.tan(math.radians(deg))
-
-def acosd(ratio):
-    return math.degrees(math.acos(ratio))
-
-def asind(ratio):
-    return math.degrees(math.asin(ratio))
-
-def atand(ratio):
-    return math.degrees(math.atan(ratio))
-
-class VehicleConstants:
-    """Class to store all unchanging vehicle parameters"""
-    
-    def __init__(self):
-        pass
-    
-    def load_vehicle(self, vehicle_dictionary):
-        for key, value in vehicle_dictionary.iteritems():
-            setattr(self, key, value)
-
-class Coordiante:
-    """Class to store waypoint on vehicle route"""
-    
-    TO_RADIANS = 0.0174533
-    TO_DEGREES = 57.2957549575 
-    
-    def __init__(self, input_latitude, input_longitude):
-        self.latitude = input_latitude
-        self.longitude = input_longitude
-        
-    def get_distance_to(self, target_coordiante):
-        """Funtion to compute the distance between two coordiantes in km.
-        Implemented using the haversine formula: https://www.movable-type.co.uk/scripts/latlong.html"""
-        
-        R = 6371e3 # radius of the earth in metres
-        latitude_1 = self.latitude * Coordiante.TO_RADIANS
-        latitude_2 = target_coordiante.latitude * Coordiante.TO_RADIANS
-        delta_latitude = (target_coordiante.latitude-self.latitude) * Coordiante.TO_RADIANS
[…]

### 2018-10-30 HighLevelCarModel/vehicle_data.py

@@ -1,67 +1,0 @@
-#!/usr/bin/env python
-
-""" Constants related to specific vehicles
-
-Module contains a dictionary for each vehicle that may be subjected to the
-model. Key value pairs are loaded into the VehicleConstants class in the model. 
-
-All units are as follows unless otherwise specified:
-    Distance: meters (m)
-    Mass: kilograms (kg)
-    Time: seconds (s)
-    Angles: radians (rad)
-    Temperature: degrees celcius (degC)
-
-Midnight Sun Solar Car Team 2018
-University of Waterloo
-"""
-
-__author__ = "Devon Copeland"
-__copyright__ = "Copyright 2018, Midnight Sun Solar Car Team"
-
-msxii = {
-    """Weight Distribution:"""
-    "mass": 550,                #curb weight, kilograms
-    "cog_x": 0,                 #center of gravity in lateral direction, m
-    "cog_y": 0.52,              #center of gravity in vertical direction, m
-    "cog_z": 1.52,              #center of gravity in longitudinal direction, m
-    "j_wheel": 2,            #polar mass moment of inertia of wheel, kg/m^2
-    "j_motor_rotor": 1,         #polar mass moment of inertia of motor, kg/m^2
-    
-    """Dimensions:"""
-    "wheelbase": 2.6,           #wheelbase, meters
-    "diam_wheel": 0.53,         #wheel diameter, meters
-    "area_frontal": 0.2,        #total frontal area, meters^2
-    
-    """Brakes:"""
-    "area_caliper_cyl": 0.0006387084, #caliper cylinder area, meters^2,
-    "area_master_cyl": 0.0002838704, #caliper cylinder area, meters^2,
-    "mu_k_brakes": 0.2,         #coeficient of drag, unitless
-    "max_line_pressue": 10^7,   #pressure, Pa, assumes 400N pedal force with 7:1 pedal ratio
-    "diam_brake_rotor": 0.53,   #wheel diameter, meters
-    "num_calipers_per_wheel": 2,          #number or brake calipers
-    
-    """Motors:"""
-    "driver_k_p": .1,
-    "driver_k_d": 0,
-    "driver_k_i": 0, 
-    
-    """Motors:"""
-    "stall_torque": 135,        #stall torque of motor, Nm
-    "no_load_speed": 1700,      #no load speed, rpm
-    "regen_gain": 0.01,         #gain applied to angular velocity to get 
-    
-    """Coefficients:"""
[…]

### 2018-10-30 MotorTorque/computeDrag.m

@@ -1,23 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Motor Torque - computeDrag.m
-% Function to compute drag force 
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2017
-%
-%
-% Notes:
-%   
-%   - Units: 
-%       Distance: meters
-%       Mass: kilograms
-%       Time: seconds
-%       Angles: degrees
-%
-% ---------------------------------------------------------------------- %
-
-function drag = computeDrag(rho, v, Cd, A)
-    drag = 0.5.*rho.*(v.^2).*Cd.*A;
-end

### 2018-10-30 MotorTorque/computeRollingResistance.m

@@ -1,25 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Motor Torque - computeRollingResistance.m
-% Function to compute rolling resistance force (Newtons). Equation from 
-% The Winning Solar Car (page 18)
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2017
-%
-%
-% Notes:
-%   
-%   - Units: 
-%       Distance: meters
-%       Mass: kilograms
-%       Time: seconds
-%       Angles: degrees
-%
-% ---------------------------------------------------------------------- %
-
-function rollingResistance = computeRollingResistance(Cr, M, v)
-    g = 9.81; % acceleration of gravity 
-    rollingResistance = Cr.*g.*M.*(1+0.02236.*v);
-end

### 2018-10-30 MotorTorque/main.m

@@ -1,67 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Motor Torque - main.m
-% Script to extimate motor torque requirements at vaious speeds, inclines,
-% and accelerations
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2017
-%
-%
-% Notes:
-%   
-%   - Units: 
-%       Distance: meters
-%       Mass: kilograms
-%       Time: seconds
-%       Angles: degrees
-%
-% ---------------------------------------------------------------------- %
-
-clear all;
-close all;
-
-% ------------------------------ Constants ---------------------------- %
-
-M = 550;        % Car mass (kg)
-r = 0.267;      % Wheel radius (m)
-Cr = 0.0055;    % Rolling resistance coefficient - Winning Solar Car pg 18
-A = 1.8;        % Frontal area (m^2)
-Cd = 0.16;      % Drag coefficient - from Ron's preliminary CFD findings
-rho = 1.225;    % Density of air (kg/m^3)
-g = 9.81;       % Acceleration of gravity (m/s^2)
-
-% ------------------------------ Variables ---------------------------- %
-x
-a = [0:.5:2.5];                      % Acceleration in m/s^2
-v = [0:.1:33.33];                    % Velocity in m/s
-incline = [0, 5,10];            % Incline in %
-theta = atan(incline/100);           % Incline in rad
-
-% ---------------------------- Plot Results ---------------------------- %
-
-figure;
-
-for i = 1:length(theta)
-    subplot(1,length(theta),i);
-    hold on;
-    for j = 1:length(a)
-        torque = 0.5*r.*(M.*a(j) + M.*g.*sin(theta(i))...
-                 + computeRollingResistance(Cr, M, v)...
-                 + computeDrag(rho, v, Cd, A)); 
-        plot(v,torque,'DisplayName', sprintf('%0.5g m/s^2',a(j)));
-    end
-    xlabel('Speed (m/s)');
-    ylabel('Torque (Nm)')
-    title(sprintf('%d%% incline',incline(i)))
-    ylim([0,400]);
-    grid on;
-end
-legend('show');
-
-
-% --------------------- Plot Cornering Results ------------------------ %
-
-% Save
-% saveas(1,'outputs/Suspension Simulation Results - Cornering','png');

### 2018-10-30 SpringRateProject/matlab/Old/timeDomainAnalysis.m

@@ -1,25 +1,0 @@
-syms t c1 c2 k1 k2 m I a b t0 A wb xddot xdot x thetaddot thetadot theta c k
-
-Ax2 = m;% X terms of ODE (A) 
-Ax1 = (c1+c2) ; 
-Ax0 = (k1+k2);
-Atheta2 = 0;
-Atheta1 = (b*c2-a*c1);
-Atheta0 = (b*k2-a*k1);
-Bx2 = 0;
-Bx1 = (b*c2-a*c1); 
-Bx0 = (b*k2-a*k1); 
-Btheta2 = I;
-Btheta1 = (b^2*c2+a^2*c1);
-Btheta0 = (b^2*k2+a^2*k1);
-
-PA = c1*A*wb*cos(wb*t) + c2*A*wb*cos(wb*(t-t0)) + k1*A*sin(wb*t) + k2*A*sin(wb*(t-t0)); 
-PB = -c1*a*A*wb*cos(wb*t) + c2*b*A*wb*cos(wb*(t-t0)) + k1*a*A*sin(wb*t) + k2*b*A*sin(wb*(t-t0)); 
-
-% Create linear systems in to solve for transfer functions
-
-A = [Ax2, Ax1, Ax0, Atheta2, Atheta1, Atheta0, PA;
-     Bx2, Bx1, Bx0, Btheta2, Btheta1, Btheta0, PB];
- 
-A = subs(A, {b, c1, c2, k1, k2}, {a, c, c, k, k});
-A

### 2018-10-30 SteeringLoadingConditions/KingPinAxisTorque.m

@@ -1,35 +1,0 @@
-contactPatch = [0.8,0,0];
-UCAKingPinAxis = [0.706,0.420,0.0133];
-LCAKingPinAxis = [0.740,0.190,0.0068];
-
-fx = [1530,0,0];
-fy = [0,1240,0];
-fz = [0,0,2700];
-
-upright = UCAKingPinAxis - LCAKingPinAxis;
-uprightUnitVector = upright./norm(upright)
-momentArm = contactPatch - LCAKingPinAxis;
-pointOnKingPinAxis = LCAKingPinAxis + dot(momentArm,uprightUnitVector).*uprightUnitVector;
-perpMomentArm = contactPatch - pointOnKingPinAxis;
-
-torqueFx = cross(perpMomentArm,fx);
-torqueFy = cross(perpMomentArm,fy);
-torqueFz = cross(perpMomentArm,fz);
-
-netKingPinAxisTorque = dot(torqueFx,uprightUnitVector) +...
-                       dot(torqueFy,uprightUnitVector) +...
-                       dot(torqueFz,uprightUnitVector)
-
-% Visualize geometry:                    
-% figure; 
-% hold on;
-% plot3(0,0,0,'ko');
-% plot3([-.8;0.8],[0;0],[0;0],'-b');
-% plot3([-.8;0.8],[0;0],[-2.6;-2.6],'-g');
-% plot3([0;0],[0;0],[0;-2.6],'-m');
-% plot3(contactPatch(1),contactPatch(2),contactPatch(3),'*r');
-% plot3(UCAKingPinAxis(1),UCAKingPinAxis(2),UCAKingPinAxis(3),'*b');
-% plot3(LCAKingPinAxis(1),LCAKingPinAxis(2),LCAKingPinAxis(3),'*b');
-% plot3([LCAKingPinAxis(1);UCAKingPinAxis(1)],...
-%       [LCAKingPinAxis(2);UCAKingPinAxis(2)],...
-%       [LCAKingPinAxis(3);UCAKingPinAxis(3)],'-b');

### 2018-10-30 SuspensionDynamics/computeCamberVSTravel.m

@@ -1,48 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Suspension - computeCamberVSTravel.m
-% Scripte to greate a fit a polynomial function to camber vs travel  
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2017
-%
-%
-% Notes:
-%   
-%   - 
-%
-% ---------------------------------------------------------------------- %
-
-clc
-clear all
-close all
-
-angleVSTravel = [-35.95, 1.043
-                 -30.21, 1.091
-                 -24.90, 1.128
-                 -23.02, 1.139
-                 -19.79, 1.155
-                 -16.47, 1.170
-                 -14.65, 1.176
-                 -11.15, 1.187
-                 -8.52, 1.193
-                 -6.64, 1.196
-                 -2.92, 1.2
-                 -0.69, 1.201
-                 2.85, 1.199
-                 7.98, 1.193
-                 11.28, 1.185
-                 15.49, 1.172
-                 18.72, 1.16
-                 22.28, 1.143
-                 24.45, 1.132
-                 27.21, 1.115
-                 28.46, 1.108];
-            
-plot(angleVSTravel(:,1),angleVSTravel(:,2));
-coeffs = polyfit(angleVSTravel(:,1),angleVSTravel(:,2),3)
-
-x = floor(angleVSTravel(1,1)):1:floor(angleVSTravel(end,1));
-hold on;
-plot(x,polyval(coeffs,x),'r');

### 2018-10-30 SuspensionDynamics/lowerControlArmForces.m

@@ -1,16 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Suspension - lowerControlArmForces.m
-% Function to compute the forces acting on the lower control arm 
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2017
-%
-%
-% Notes:
-%   
-%   - Vector quantities are of the form [x; y; z]
-% ---------------------------------------------------------------------- %
-
-function [fPushrod] = pushRodForce(fUpright, pHardpoint, pUpright, p)

### 2018-10-30 SuspensionDynamics/staticScalarTrailingArmForces.m

@@ -1,37 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Suspension - staticTrailingArmForces.m
-% Function to compute the forces acting on the trailing arm 
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2017
-%
-%
-% Notes:
-%   
-%   - MSXII global coordinat system used
-%   - Scalar verison of staticTrailingArmForces()
-%   - Considers loading only in the static case
-%   - Simplifies problem to 2D
-%   - coiloverAngle is the angle that the coilover strut makes with the
-%     central axis of the trailing arm
-%   - trailingArmAngle is the angle the trailing arm makes with horizontal
-% ---------------------------------------------------------------------- %
-
-function [fMountY, fMountZ, fCoilover] = staticScalarTrailingArmForces(fRoadY, fRoadyZ, coiloverAngle,...
-                                                       trailingArmAngle,... 
-                                                       wheelMomentArm,...
-                                                       coiloverLateralMomentArm,...
-                                                       coiloverLongitudianalMomentArm)
-                                                   
-    fCoilover = (fRoadY.*cosd(trailingArmAngle) - fRoadyZ.*cosd(trailingArmAngle)).*wheelMomentArm /...
-                (cosd(trailingArmAngle).*coiloverLongitudianalMomentArm +...
-                 sind(coiloverAngle).*coiloverLateralMomentArm);
-                
-    fCoiloverZ = fCoilover*sind(coiloverAngle-trailingArmAngle);
-    fCoiloverY = fCoilover*cosd(coiloverAngle-trailingArmAngle);
-    
-    fMountZ = fRoadyZ - fCoiloverZ;
-    fMountY = fCoiloverY - fRoadY;
-end

### 2018-10-30 SuspensionDynamics/staticTrailingArmForces.m

@@ -1,37 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Suspension - staticTrailingArmForces.m
-% Function to compute the forces acting on the trailing arm 
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2017
-%
-%
-% Notes:
-%   
-%   - Vectors are of the form [x; y; z] in the MSXII coordinate system
-%   - Considers loading only in the static case
-%   - Simplifies problem to 2D
-%   - coiloverAngle is the angle that the coilover strut makes with the
-%     central axis of the trailing arm
-%   - trailingArmAngle is the angle the trailing arm makes with horizontal
-% ---------------------------------------------------------------------- %
-
-function [fMount, fCoilover] = staticTrailingArmForces(fRoad, coiloverAngle,...
-                                                       trailingArmAngle,... 
-                                                       wheelMomentArm,...
-                                                       coiloverLateralMomentArm,...
-                                                       coiloverLongitudianalMomentArm)
-                                                   
-    fCoiloverMag = (fRoad(2)*cosd(trailingArmAngle) - fRoad(3)*cosd(trailingArmAngle))*wheelMomentArm /...
-                   (cosd(trailingArmAngle)*coiloverLongitudianalMomentArm +...
-                    sind(coiloverAngle)*coiloverLateralMomentArm);
-                
-    fCoilover(3) = fCoiloverMag*sind(coiloverAngle-trailingArmAngle);
-    fCoilover(2) = fCoiloverMag*cosd(coiloverAngle-trailingArmAngle);
-    
-    fMount = [0;0;0];
-    fMount(3) = fRoad(3) - fCoilover(3);
-    fMount(2) = fCoilover(2) - fRoad(2);
-end

### 2018-10-30 SuspensionDynamics/trailingArmForces.m

@@ -1,52 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% SuspensionDynamics - trailingArmForces.m
-% Script to extimate the forces on the trailing arm based on loads from the
-% wheel 
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2017
-%
-%
-% Notes:
-%   
-%   - Units: 
-%       Distance: meters
-%       Mass: kilograms
-%       Time: seconds
-%       Angles: degrees
-%
-% ---------------------------------------------------------------------- %
-
-clear all;
-close all;
-
-% ------------------------------ Constants ---------------------------- %
-
-coiloverAngle = 45; % angle that the coilover strut makes with the central axis of the trailing arm in degrees
-trailingArmAngle = 15; % angle the trailing arm makes with horizontal in degrees
-wheelMomentArm = 0.300; % (meters) distance between wheel center and pivot point
-coiloverLateralMomentArm = 0.05;% (meters) distance between trailing arm central axis and coilover mounting point
-coiloverLongitudianalMomentArm = 0.100; % (meters) distance along trailing arm centerline between pivot point and coilover mounting point
-
-% ------------------------------ Variables ---------------------------- %
-
-fRoadNormalRange = [0:50:(500*0.5*9.81)]; % Newtons
-fRoadFricitonRange = [0:5:(60/(.5*.533))]; % Newtons
-
-[fRoadNormal, fRoadFriciton] = meshgrid(fRoadNormalRange, fRoadFricitonRange);
-
-% ------------------------------ Results ------------------------------ %
-
-[fMountY, fMountZ, fCoilover] = staticScalarTrailingArmForces(fRoadNormal,... 
-                                fRoadFriciton, coiloverAngle, trailingArmAngle,... 
-                                wheelMomentArm, coiloverLateralMomentArm,...
-                                coiloverLongitudianalMomentArm);
-                           
-figure;
-hold on;
-surf(fRoadNormal, fRoadFriciton, fCoilover);
-surf(fRoadNormal, fRoadFriciton, fMountY);
-surf(fRoadNormal, fRoadFriciton, fMountZ);
-

### 2018-10-30 SuspensionDynamics/uprightForces.m

@@ -1,43 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Suspension - uprightForces.m
-% Function to compute the forces acting on the upright 
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2017
-%
-%
-% Notes:
-%   
-%   - Vector quantities are of the form [x; y; z]
-%   - The function solves 6 equations 
-%       - sum of the forces in all 3 axes = mass*accel - applied forces
-%       - sum of the moments in all 3 axes = moment of inertia*accel -
-%       applied moments
-% ---------------------------------------------------------------------- %
-
-function [fUCA, fLCA] = uprightForces(mass, moi, cog, linearAccel,...
-                                      angularAccel, fRoad, posUCA, posLCA,... 
-                                      posContact)
-    % Compute moment arms
-    rUCA = posUCA - cog;
-    rLCA = posLCA - cog;
-    rRoad = posContact - cog;
-    % Compute weight vector
-    weight = mass.*[0;-9.81;0];
-    % Setup linear equations
-    A = [ 1        0        0        1        0        0
-          0        1        0        0        1        0
-          0        0        1        0        0        1
-          0        -rUCA(3) rUCA(2)  0        -rLCA(3) rLCA(2)
-          rUCA(3)  0        -rUCA(1) rLCA(3)  0        -rLCA(1)
-          -rUCA(2) rUCA(1)  0        -rLCA(2) rLCA(1)  0 ];
-    b = [(mass.*linearAccel - fRoad - weight); 
-         (moi.*angularAccel - fRoad.*rRoad)];
-    x = linsovle(A,b);
-    % Parse results
-    fUCA = x(1:3);
-    fLCA = x(4:6);
-
-end

### 2018-10-30 SuspensionSimulator/ZZcomputeNodalPositions.m

@@ -1,61 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Double Wishbone Suspension Simulator - computeNodalPositions.m
-% Function to compute x,y position of nodes from properties
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2016
-%
-%
-% Notes:
-%
-%   - The order of params is as follows:
-%        1)  UCA mount (x)
-%        2)  UCA mount (y)
-%        3)  LCA mount (x)
-%        4)  LCA mount (y)
-%        5)  UCA length
-%        6)  LCA length
-%        7)  upright length
-%        8)  spindle height
-%        9)  wheel offset
-%        10) wheel diameter
-%        11) inclination angle
-%
-% ---------------------------------------------------------------------- %
-
-function nodes = computeNodalPositions(params, lcaAngle)
-    
-    % Control arm mounting postions
-    nodes(1,1) = params(1);
-    nodes(1,2) = params(2);
-    nodes(2,1) = params(3);
-    nodes(2,2) = params(4);
-    
-    % 
-    
-    
-    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    
-    % End of LCA
-    nodes(5,1) = nodes(2,1) + params(6) * cosd(lcaAngle);
-    nodes(5,2) = nodes(2,2) + params(6) * sind(lcaAngle);
-    
-    % End of UCA 
-    [xout,yout] = circcirc(x1,y1,r1,x2,y2,r2)
-    nodes(3,1) = nodes(1,1) + ((nodes(1,1) + UCA - y)*(UCA - Ay + y))^(1/2)
-    
-    % Distance from the UCA mount to the end of the LCA
-    ucaMount2LCAend = sqrt((nodes(1,1)-nodes(5,1))^2 ... 
-                         + (nodes(1,2)-nodes(5,2))^2); % by distance formula
-    
-    % Temp angle is the angle formed btwn the UCA and the line from the UCA
-    % mount to the end of the LCA
-    tempAngle = acosd( (params(5)^2 + ucaMount2LCAend^2 - params(7)^2)...
-                      /(2 * params(5) * ucaMount2LCAend)); % by cosine law
-    
-    % End of UCA 
-    
-    
-end 

### 2018-10-30 SuspensionSimulator/assembleSolution.m

@@ -1,53 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Double Wishbone Suspension Simulator - assembleSolution.m
-% A function that returns solved suspension geometry from the output of 
-% fsolve and the input parameters
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2016
-%
-%
-% Notes:
-%
-%    - The order of params is as follows:
-%        1)  UCA mount (x)
-%        2)  UCA mount (y)
-%        3)  LCA mount (x)
-%        4)  LCA mount (y)
-%        5)  UCA length
-%        6)  LCA length
-%        7)  upright length
-%        8)  spindle height
-%        9)  wheel offset
-%        10) wheel diameter
-%        11) inclination angle
-%
-%    - The order of unknonws is as follows 
-%        1)  Node 3 (x)
-%        2)  Node 3 (y)
-%        3)  Node 4 (x)
-%        4)  Node 4 (y)
-%        5)  Node 5 (x)
-%        6)  Node 5 (y)
-%        7)  Node 6 (x)
-%        8)  Node 6 (y)
-%        9)  Node 7 (x)
-%        10) Node 7 (y)
-%        11) Node 8 (x)
-%
-%    - In nodes, each row represents a node: (x,y)
-%
-% ---------------------------------------------------------------------- %
-
-function nodes = assembleSolution(solvedNodes, params)
-    nodes = [ params(1),       params(2);       % UCA mount
-              params(3),       params(4);       % LCA mount
-              solvedNodes(1),  solvedNodes(2);  % upright top
-              solvedNodes(3),  solvedNodes(4);  % upright spindle mount
-              solvedNodes(5),  solvedNodes(6);  % upright btm
-              solvedNodes(7),  solvedNodes(8);  % wheel top
-              solvedNodes(9),  solvedNodes(10); % wheel center
-              solvedNodes(11), 0    ];          % wheel btm 
-end 

### 2018-10-30 SuspensionSimulator/computeDistance.m

@@ -1,18 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Double Wishbone Suspension Simulator - computeDistance.m
-% A function that returns the distance between two cartesian points
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2016
-%
-%
-% Notes:
-%
-%   - 
-% ---------------------------------------------------------------------- %
-
-function d = computeDistance(x1,y1,x2,y2)
-    d = sqrt((x1-x2)^2 + (y1-y2)^2);
-end 

### 2018-10-30 SuspensionSimulator/computeSuspensionError.m

@@ -1,94 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Double Wishbone Suspension Simulator - computeSuspensionError.m
-% An error function to be minimized by fsolve and solve for nodal positions
-% in the process
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2016
-%
-%
-% Notes:
-%
-%   - parmams is a vector containing the parameters that define the
-%     suspension geometry. guess is an vector of unknown nodal coordinates. 
-%
-%   - The order of params is as follows:
-%        1)  UCA mount (x)
-%        2)  UCA mount (y)
-%        3)  LCA mount (x)
-%        4)  LCA mount (y)
-%        5)  UCA length
-%        6)  LCA length
-%        7)  upright length
-%        8)  spindle height
-%        9)  wheel offset
-%        10) wheel diameter
-%        11) spindle angle
-%
-%    - The order of the guess is as follows 
-%        1)  Node 3 (x)
-%        2)  Node 3 (y)
-%        3)  Node 4 (x)
-%        4)  Node 4 (y)
-%        5)  Node 5 (x)
-%        6)  Node 5 (y)
-%        7)  Node 6 (x)
-%        8)  Node 6 (y)
-%        9)  Node 7 (x)
-%        10) Node 7 (y)
-%        11) Node 8 (x)
-%
-% ---------------------------------------------------------------------- %
-
-function error = computeSuspensionError(guess, params)
-    
-    % The first six equations are driven by the lengths of the elements 
-    % in the suspension (parameters 5 to 10) and are driven by distance
-    % formula. 
-    error(1) = computeDistance(params(1),params(2),guess(1),guess(2)) ...
-               - params(5);  % UCA length
-           
-    error(2) = computeDistance(params(3),params(4),guess(5),guess(6)) ...
-               - params(6);  % LCA length
-           
-    error(3) = computeDistance(guess(1),guess(2),guess(5),guess(6)) ...
-               - params(7);  % upright length
-           
-    error(4) = computeDistance(guess(3),guess(4),guess(5),guess(6)) ...
-               - params(8);  % spindle height
-           
-    error(5) = computeDistance(guess(3),guess(4),guess(9),guess(10)) ...
-               - params(9);  % wheel offset
-           
-    error(6) = computeDistance(guess(7),guess(8),guess(11),0) ...
-               - params(10); % wheel diameter
-    
[…]

### 2018-10-30 SuspensionSimulator/cosLawGetOppositeLength.m

@@ -1,24 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Double Wishbone Suspension Simulator - computeSuspensionError.m
-% A function that returns the length of a triangle's legs given the angle
-% opposite to the desigred leg and the lengths of the other two legs. 
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2016
-%
-%
-% Notes:
-%
-%   - Uses the cosine law: http://hyperphysics.phy-astr.gsu.edu/hbase/lcos.html
-%
-%   - angleC is the known angle opposite of leg c in degrees
-%
-%   - a and b are the other two legs of known length
-%
-% ---------------------------------------------------------------------- %
-
-function c = cosLawGetOppositeLength(a,b,angleC)
-    c = sqrt(a^2+b^2-(2*a*b*cosd(angleC)));
-end

### 2018-10-30 SuspensionSimulator/extractUnknownNodes.m

@@ -1,41 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Double Wishbone Suspension Simulator - extractUnknownNodes.m
-% A function that returns the unknown nodes in the suspension 
-% This function is used to parse the guess. 
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2016
-%
-%
-% Notes:
-%
-%    - The order of unknonws is as follows 
-%        1)  Node 3 (x)
-%        2)  Node 3 (y)
-%        3)  Node 4 (x)
-%        4)  Node 4 (y)
-%        5)  Node 5 (x)
-%        6)  Node 5 (y)
-%        7)  Node 6 (x)
-%        8)  Node 6 (y)
-%        9)  Node 7 (x)
-%        10) Node 7 (y)
-%        11) Node 8 (x)
-%
-% ---------------------------------------------------------------------- %
-
-function unknowns = extractUnknownNodes(nodes)
-    unknowns = [ nodes(3,1);
-                 nodes(3,2);
-                 nodes(4,1);
-                 nodes(4,2);
-                 nodes(5,1);
-                 nodes(5,2);
-                 nodes(6,1);
-                 nodes(6,2);
-                 nodes(7,1);
-                 nodes(7,2);
-                 nodes(8,1)  ];
-end 

### 2018-10-30 SuspensionSimulator/getCamber.m

@@ -1,21 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Double Wishbone Suspension Simulator - getCamber.m
-% A function that returns the camber angle in degrees 
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2016
-%
-%
-% Notes:
-%
-%   - Negative camber occurs when the top of the wheel is close to the
-%     center of the car than the bottom
-% 
-% ---------------------------------------------------------------------- %
-
-function camber = getCamber(nodes)
-    camber = atan2d(nodes(8,1)-(nodes(6,1)),(nodes(6,2)-nodes(8,2)));
-end 
-

### 2018-10-30 SuspensionSimulator/getKingPinAngle.m

@@ -1,19 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Double Wishbone Suspension Simulator - getKingPinAngle.m
-% A function that returns the king pin angle in degrees 
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2016
-%
-%
-% Notes:
-%
-%   - 
-% 
-% ---------------------------------------------------------------------- %
-
-function kingPinAngle = getKingPinAngle(nodes)
-    kingPinAngle = atan2d(nodes(5,1)-(nodes(3,1)),(nodes(3,2)-nodes(5,2)));
-end 

### 2018-10-30 SuspensionSimulator/getLateralScrub.m

@@ -1,19 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Double Wishbone Suspension Simulator - getLateralScrub.m
-% A function that returns the lateral scrub distance in mm 
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2016
-%
-%
-% Notes:
-%
-%   - 
-% 
-% ---------------------------------------------------------------------- %
-
-function lateralScrub = getLateralScrub(nodes,origNodes)
-    lateralScrub = nodes(8,1) - origNodes(8,1);
-end 

### 2018-10-30 SuspensionSimulator/getRollCenter.m

@@ -1,69 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Double Wishbone Suspension Simulator - getRollCenter.m
-% A function that computes the roll center of the vehicle 
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2016
-%
-%
-% Notes:
-%
-%   - 
-% 
-% ---------------------------------------------------------------------- %
-
-function [x,y] = getRollCenter(nodesRight, nodesLeft, PLOT_ROLL_CENTER, limits)
-    
-    % Right instantaneous center
-    rUCAEqn = polyfit([nodesRight(1,1),nodesRight(3,1)], ...
-                      [nodesRight(1,2),nodesRight(3,2)],1);
-    rLCAEqn = polyfit([nodesRight(2,1),nodesRight(5,1)], ...
-                      [nodesRight(2,2),nodesRight(5,2)],1);
-    A = [-1 * rUCAEqn(1), 1; -1 * rLCAEqn(1), 1];
-    b = [rUCAEqn(2); rLCAEqn(2)];
-    rightInstCenter = linsolve(A,b)';
-    
-    % Left instantaneous center 
-    lUCAEqn = polyfit([nodesLeft(1,1),nodesLeft(3,1)], ...
-                      [nodesLeft(1,2),nodesLeft(3,2)],1);
-    lLCAEqn = polyfit([nodesLeft(2,1),nodesLeft(5,1)], ...
-                      [nodesLeft(2,2),nodesLeft(5,2)],1);
-    A = [-1 * lUCAEqn(1), 1; -1 * lLCAEqn(1), 1];
-    b = [lUCAEqn(2); lLCAEqn(2)];
-    leftInstCenter = linsolve(A,b)';
-    
-    % Roll center
-    rightInst2Wheel = polyfit([rightInstCenter(1),nodesRight(8,1)], ...
-                              [rightInstCenter(2),0],1);
-    leftInst2Wheel = polyfit([leftInstCenter(1),nodesLeft(8,1)], ...
-                             [leftInstCenter(2),0],1);
-    A = [-1 * rightInst2Wheel(1), 1; -1 * leftInst2Wheel(1), 1];
-    b = [rightInst2Wheel(2); leftInst2Wheel(2)];
-    rollCenter = linsolve(A,b)';
-    x = rollCenter(1);
-    y = rollCenter(2);
-    
-    % Plot roll center geometry
-    if PLOT_ROLL_CENTER
[…]

### 2018-10-30 SuspensionSimulator/getScrubRadius.m

@@ -1,21 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Double Wishbone Suspension Simulator - getScrubRadius.m
-% A function that returns the scrub radius in mm 
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2016
-%
-%
-% Notes:
-%
-%   - 
-% 
-% ---------------------------------------------------------------------- %
-
-function scrubRadius = getScrubRadius(nodes)
-    dXdY = (nodes(3,1) - nodes(5,1)) / (nodes(3,2) - nodes(5,2));
-    roadIntersectionPoint = nodes(5,1) - dXdY * nodes(5,2); 
-    scrubRadius = nodes(8,1) - roadIntersectionPoint;
-end 

### 2018-10-30 SuspensionSimulator/main.m

@@ -1,274 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Double Wishbone Suspension Simulator - main.m
-% Script to simulate the suspension geometry of MSXII
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2016
-%
-%
-% Notes:
-%
-%   - Coordinte system follows standard for MSXII: 
-%     https://uwmidsun.atlassian.net/wiki/display/MECH/MSXII
-%   
-%   - Units: 
-%       Distance: millimeters
-%       Mass: kilograms
-%       Time: seconds
-%       Angles: degrees
-%
-% ---------------------------------------------------------------------- %
-
-clear all;
-close all;
-
-% ------------------------------ Config -------------------------------- %
-
-% Plots
-VISUALIZER_ON = true;
-PLOT_ROLL_CENTER_GEOMETRY = false;
-
-% Longitudial shift
-START_DISP = 0;
-END_DISP = 100;
-DISP_STEP = 5;
-NEUTRAL_DISP = 0;
-
-% Cornering (roll)
-ROLL_STEP = .01;
-END_ROLL = 4; 
-
-% Override 
-if PLOT_ROLL_CENTER_GEOMETRY
-    VISUALIZER_ON = true;
-end
-
-% --------------------------- Input Parameters ------------------------- %
-
-% Vector contianing the parameters which define the suspension geometry
-parametersRight = [ 1199.0;   % UCA mount (x)
-                    600.0;    % UCA mount (y)
-                    999.0;    % LCA mount (x)
-                    199.0;    % LCA mount (y)
-                    250.0;    % UCA length
-                    500.0;    % LCA length
-                    450.0;    % upright length
-                    157.0;    % spindle height
-                    130.0;    % wheel offset
-                    700.0;    % wheel diameter
-                    83   ];   % spindle angle
-
-% Generate parameters for left wheel
-parametersLeft = parametersRight; 
-parametersLeft(1) = parametersLeft(1) * -1;
-parametersLeft(3) = parametersLeft(3) * -1;
-
-% ------------------------- Initialize Geometry ------------------------ %
-
-% An approximation of double wishbode suspension to give fsolve a good
-% place to start:
-% Each row represents a node: (x,y)
[…]

### 2018-10-30 SuspensionSimulator/printLabels.m

@@ -1,32 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Double Wishbone Suspension Simulator - printRoad.m
-% Fuction to add the road to the visualizer plot and format it
-% 
-%
-% Notes:
-%
-%   - 
-%
-% ---------------------------------------------------------------------- %
-
-function printLabels(limits) 
-
-    % Define plot limits
-    xlimits = limits(1,:); 
-    ylimits = limits(2,:); 
-    
-    % Show ground
-    plot([xlimits(1)*10 xlimits(2)*10],[0 0],'k--');
-    
-    % Legend
-    legend('Neutral Position','Dynamic','Location','southoutside');
-    
-    hold off;
-    xlabel('X-Position (mm)');
-    ylabel('Y-Position (mm)');
-    xlim(xlimits);
-    ylim(ylimits);
-    title('Suspension Visualizer');
-end
-    

### 2018-10-30 SuspensionSimulator/printSimulationParams.m

@@ -1,31 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Double Wishbone Suspension Simulator - printSimulationParams.m
-% Fuction to add the simulation parameters to the plot
-% 
-%
-% Notes:
-%
-%   - 
-%
-% ---------------------------------------------------------------------- %
-
-function printSimulationParams(params) 
-
-    % Add parameters
-    dim = [.15 .6 0.3 0.3];
-    params = { 'Simulation Parameters:',...
-               sprintf('R-UCA x: %d', params(1)),... 
-               sprintf('R-UCA y: %d', params(2)),... 
-               sprintf('R-LCA x: %d', params(3)),... 
-               sprintf('R-LCA y: %d', params(4)),... 
-               sprintf('UCA length: %d', params(5)),... 
-               sprintf('LCA length: %d', params(6)),... 
-               sprintf('Upright length: %d', params(7)),... 
-               sprintf('Spindle height: %d', params(8)),... 
-               sprintf('Wheel offset: %d', params(9)),... 
-               sprintf('Wheel diameter: %d', params(10)),... 
-               sprintf('Spindle angle: %d', params(11)) };
-    annotation('textbox',dim,'String',params,'FitBoxToText','on');
-    
-end

### 2018-10-30 SuspensionSimulator/printSuspension.m

@@ -1,68 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Double Wishbone Suspension Simulator - printSuspension.m
-% Fuction to print the current state of the suspension to a figure
-% 
-% Original version by                                                
-%   Dr. Haim Waisman, Columbia University, New York, Copyright 2009    
-% Modified 2013                                                      
-%   Dr. Robert Gracie, University of Waterloo
-% Modified 2014
-%   Christopher Kohar, University of Waterloo
-% Modified 2016
-%   Devon Copeland, Midnight Sun, University of Waterloo
-%
-% Notes:
-%
-%   - nodes is a matrix with n rows of the format x,y representing n nodes
-%   - initialState is a matrix of nodes at the suspensions neutral position
-%   - elmntMap maps has m rows of the format node1,node2 representing
-%     the m element in the suspension
-%
-% ---------------------------------------------------------------------- %
-
-function printSuspension(initialState,nodes,elmntMap,limits) 
-
-    numbElmnts = size(elmntMap,1); % number of elements
-    xInit = initialState(:,1); % isolate x coordinates
-    yInit = initialState(:,2); % isolate y coordinates
-    x = nodes(:,1); % isolate x coordinates
-    y = nodes(:,2); % isolate y coordinates
-
-    % Define plot limits
-    xlimits = limits(1,:); 
-    ylimits = limits(2,:); 
-    
-    % Plot the structure
-    for i = 1:numbElmnts
-
-        % Plot initial state of the suspension
-        XXI = [xInit(elmntMap(i,1)), xInit(elmntMap(i,2))];
-        YYI = [yInit(elmntMap(i,1)), yInit(elmntMap(i,2))];
-        line(XXI,YYI,'LineWidth',1);
-        hold on;
-
-        % PLOT FINAL DEFORMED STRUCTURE
-        XXF = [x(elmntMap(i,1)), x(elmntMap(i,2))];
-        YYF = [y(elmntMap(i,1)), y(elmntMap(i,2))];
-        line(XXF,YYF,'LineWidth',2,'Color','RED','LineStyle','--');hold on;
-        % Initial Node Location Labels
-        text(XXI(1),YYI(1),sprintf('%0.5g',elmntMap(i,1)));
[…]

### 2018-10-30 SuspensionSimulator/rotateCoords.m

@@ -1,26 +1,0 @@
-% ---------------------------------------------------------------------- %
-%
-% Double Wishbone Suspension Simulator - rotateCoords.m
-% A function that rotates an input coordinate about a specified center of
-% rotation by a specified angle in degrees
-% 
-% Author: Devon Copeland
-% 
-% © Midnight Sun 2016
-%
-%
-% Notes:
-%
-%   - Positive angles are CCW
-% 
-% ---------------------------------------------------------------------- %
-
-function [x,y] = rotateCoords(inputCoords, centerOfRotation, angle)
-    inputCoords = inputCoords - centerOfRotation;
-    [theta,rho] = cart2pol(inputCoords(1),inputCoords(2));
-    theta = theta + deg2rad(angle); 
-    [outputCoords(1),outputCoords(2)] = pol2cart(theta,rho);
-    outputCoords = outputCoords + centerOfRotation;
-    x = outputCoords(1);
-    y = outputCoords(2);
-end 

### 2018-10-30 MSXII/AerobodyFlowVisualization/MSXIIAerobodyFlowVisualization.tex
Moved old projects to MSXII folder

\documentclass[12pt]{article}

% Packages
\usepackage{lipsum}
\usepackage{setspace}
\usepackage{indentfirst}
\usepackage{geometry}
\usepackage[nonumberlist, toc]{glossaries}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage [autostyle, english = american]{csquotes}
\usepackage{fancyhdr}
\usepackage{amsmath}
\usepackage{cite}

% Formatting
\geometry{letterpaper, portrait, margin=.85in}
\MakeOuterQuote{"}
\pagestyle{fancy}
\lhead{MSXII Aerobody Flow Visualization} 
\rhead{Midnight Sun Solar Car Team}

\begin{document}

% Title Ppage
\begin{titlepage}
	\vspace*{2cm}
	\centering
	\includegraphics[width=.25\textwidth]{./images/midnightSunLogoCircle.png}\par
	\vspace{1.5cm}
	{\scshape\LARGE Midnight Sun Solar Car Team \par}
	{\scshape\large University of Waterloo\par}
	\vspace{2.2cm}
	{\huge\bfseries MSXII Aerobody Flow Visualization\par}
	\vspace{0.2cm}
	\large Prepared for ME351 Project 1 
	\vspace{2.2cm}	
	\par Prepared by:\par
	Danyon Chu - 20563165\par
	Devon Copeland - 20553468\par

	\vspace{0.6cm} 
	\today\par
	\vfill
	www.uwmidsun.com \par
	(519) 888-4567 x 32978
\end{titlepage}

% Main Matter
\section{Background Information and Description of Flow}
Midnight Sun Twelve (MSXII) is a cruiser class, solar electric vehicle being designed with the goal of competing in the 2018 American Solar Challenge (ASC 2018) and the 2019 World Solar Challenge (WSC 2019). By definition, cruiser class solar vehicles must be multi-occupant and are designed with the intent of being more practical than a typical, challenger class solar car. The figure of merit for scoring for this class is predominantly efficiency based however teams are also awarded points for practicality and incorporating features that bring the car closer to what one might find in a consumer vehicle. Because of the unique challenges of this competition the areobody of MSXII must be designed to minimize drag while still keeping vehicle aesthetics and driver comfort in mind. This report summarizes an experiment that provides qualitative learnings regarding the flow of air over MSXII's aerobody by using a 3D printed scale model of the aerobody and a wind tunnel.

\section{Experimental Setup}
In order to visualize the flow of air over MSXII while driving, a wind tunnel was built and a 1:30 scale model of MSXII's aerobody was 3D printed. The wind tunnel was made using three boards of masonite and one piece of cast acrylic which provided a window such that the air flow could be seen from the outside. Opposite the acrylic, a dark coloured piece of paper was glued to the masonite on the inside of the tunnel to provide contrast against smoke. A fan with adjustable speed was then placed at one end of the tunnel to generate the airflow. Since the air flow downstream from the fan would likely be turbulent, the air was directed through a large bundle of straws before making contact with the car. The straws helped to create laminar airflow and removed disturbances from the fan. In order to visualize the airflow, a draft detector (smoke stick) was purchased to generate smoke. The smoke was then sucked from a reservoir outside the tunnel through a straw and then released just in front of the 3D printed model.

\begin{figure}[h!]
	\centering
	\includegraphics[width=.7\textwidth]{./images/setup1.jpg}
	\caption{Wind tunnel construction}
	\label{fig:setup1}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width=.6\textwidth]{./images/draftDetector.png}
	\caption{The draft detector used to generate smoke}
	\label{fig:draftDetector}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width=.7\textwidth]{./images/setup2.jpg}
    \captionsetup{justification=centering,margin=2cm}
	\caption{Smoke is transported through a straw from a reservoir outside the tunnel and released directly in front of the 3D printed part}
	\label{fig:setup2}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width=.7\textwidth]{./images/setup3.jpg}
    \captionsetup{justification=centering,margin=2cm}
	\caption{Straws placed downstream of the fan help to remove turbulence from the moving air column}
	\label{fig:setup3}
\end{figure}

\section{Observations}
Several observations can be made based on the flow of smoke over the 3D printed aerobody. One such observation relates to the experimental setup. As noted in the previous section, several steps were taken to ensure that the air flow in the wind tunnel was laminar; including the wall of straws and adjustable speed on the fan. While conducting the experiment, the speed of the fan was reduced to it's lowest setting and the smoke in the wind tunnel created a fairly clean streakline before striking the body of the vehicle. This indicates that these design considerations have helped accomplished the goal of reducing turbulence in the airflow. These observations also coincide with trends in the literature; where higher velocity air flows tend to be more turbulent than lower velocity air flows and small diameter flows (like that of the straws) tend to be more laminar. Explained differently, based on the the governing equation for Reynolds Number, speed plays a direct roll in determining whether a flow is turbulent or laminar:

\begin{center}
$Re = \frac{VL}{\mu}$ 
\end{center}

where: V = velocity, L = length, $\mu$ = kinematic velocity and that a high Re is an indicator of turbulent air flow.

Another notable observation relates to the position from which the smoke was released. Through experimentation it was determined that releasing the smoke from a small aperture led to a cleaner streakline. For this reason, the smoke from the draft detector was fed into the wind tunnel through a single straw connected to a reservoir rather than simply through the fan.

A third key observation relates to the flat area on the rear of the car that was originally designed to accommodate a road legal license plate. This flat area was predicted to lead to a less aerodynamic design. Although an eddy can still be seen at the rear of the vehicle in Figure \ref{fig:sideFlow}, the flow moving underneath the car through the catamaran seen in Figure \ref{fig:rearFlow} seemed to help mitigate or reduce the amount of negative pressure and eddies seen behind the vehicle.

Overall, the most important observation that the custom wind tunnel has enabled is a high level view of the flow separation across the vehicle. Where possible, MSXII was designed to emulate the optimal teardrop shape while still resembling an actual vehicle. In doing so a number of simulations were done to try to optimize certain features for efficiency. By providing a physical representation of the streaklines over the car, this study has helped to corroborate data from CFD and draw connections between these simulations and the real world.\\

To view a video of the experiment, please visit: www.tinyurl.com/msxiiflow

\begin{figure}[h!]
	\centering
	\includegraphics[width=.7\textwidth]{./images/sideFlow1.jpg}
    \captionsetup{justification=centering,margin=2cm}
	\caption{Laminar air stream over the MSXII model as seen from the side view}
	\label{fig:sideFlow}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width=.7\textwidth]{./images/topFlow1.jpg}
    \captionsetup{justification=centering,margin=2cm}
	\caption{Laminar air stream over the MSXII model as seen from the top view}
	\label{fig:topFlow}
\end{figure}

\begin{figure}[h!]
	\centering
	\includegraphics[width=.7\textwidth]{./images/rearFlow1.jpg}
    \captionsetup{justification=centering,margin=2cm}
	\caption{Laminar air stream over the MSXII model as seen from the rear view}
	\label{fig:rearFlow}
\end{figure}

% Bibliography
\pagebreak
\bibliography{./latex/bib}{}
\bibliographystyle{plain}

\end{document}

### 2018-10-30 MSXII/Brakes/forwardDynamicsBrakingCalculations.m

@@ -1,0 +1,126 @@
+% ---------------------------------------------------------------------- %
+%
+% Longitudinal Weight Balance - longWeightBalance.m
+% Script to compute the longitudinal weight balance of the car during
+% braking or acceleration.
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2017
+%
+%
+% Notes:
+%   
+%   - Units: 
+%       Distance: meters
+%       Mass: kilograms
+%       Time: seconds
+%       Angles: degrees
+%
+%   - All force vectors are of the form [front, rear]
+%
+%   - Sliding condiditons: 
+%       - 0: No sliding
+%       - 1: Front sliding
+%       - 2: Rear sliding
+%       - 3: both sliding
+%
+% ---------------------------------------------------------------------- %
+
+close all
+clear all
+
+% --------------------------- Constants ------------------------------- %
+
+m = 550;            % car mass in kg
+cogZ = 1.52;         % Center of gravity position in longitudinal direction (wrt global origin)
+cogY = 0.52;         % Center of gravity position in vertical direction (wrt global origin)
+wheelbase = 2.6;   % Wheelbase in m
+g = -9.81;          % Acceleration of gravity (m/s^2)
+muBrakePad = [0.35, 0.4];  % Coefficient of friction between the tire and road [dynamic, static]
+muRoad = [0.6, 0.7];       % Coefficient of friction between the tire and road [dynamic, static]
+dynamic = 1; % Constant to increase code readabliltiy 
+static = 2; % Constant to increase code readabliltiy 
+wheelDiam = 0.533;  % Wheel Diameter in m
+rotorDiam = [.18, .3];    % Rotor diameter at center of brake pad [front, rear]
+padContactArea = 0.00129; % Caliper Contact area in m^2 http://www.wilwood.com/BrakePads/BrakePadsProd.aspx?itemno=150-4091K
+caliperCylinderArea = 0.001; % Area of caliper cylinder 
+masterCylinderArea = 0.001; % Area of master cylinder 
+brakePedalReduction = 3; % 3:1 mechanical advantage on brake pedal
+
+% --------------------------- Variables ------------------------------- %
+
+frontPressureRange = 5e5:5e5:20e6;
+pressureRatioRange = 0:.05:1;
+
+[frontPressure, pressureRatio] = meshgrid(frontPressureRange, pressureRatioRange);
+
[…]

### 2018-10-30 MSXII/Brakes/inverseDynamicsBrakingCalculations.m

@@ -1,0 +1,74 @@
+% ---------------------------------------------------------------------- %
+%
+% Longitudinal Weight Balance - longWeightBalance.m
+% Script to compute the longitudinal weight balance of the car during
+% braking or acceleration.
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2017
+%
+%
+% Notes:
+%   
+%   - Units: 
+%       Distance: meters
+%       Mass: kilograms
+%       Time: seconds
+%       Angles: degrees
+%
+% ---------------------------------------------------------------------- %
+
+close all
+clear all
+
+% --------------------------- Constants ------------------------------- %
+
+m = 611;            % car mass in kg
+cogZ = 0.8;         % Center of gravity position in longitudinal direction
+cogY = 0.5;         % Center of gravity position in vertical direction
+wheelbase = 1.78;   % Wheelbase in m
+g = -9.81;          % Acceleration of gravity (m/s^2)
+muBrakePad = 0.35;  % Coefficient of friction between the tire and road
+muRoad = 0.6;       % Coefficient of friction between the tire and road
+
+a = 5;              % Target acceleration
+frontWeightBalance = .65; % Proportion of car's weight on front wheels druing hard
+                          % braking
+wheelDiam = 0.533;  % Wheel Diameter in m
+rotorDiam = .18;    % Rotor diameter at center of brake pad
+padContactArea = 0.00129 % Caliper Contact area in m^2 http://www.wilwood.com/BrakePads/BrakePadsProd.aspx?itemno=150-4091K
+caliperCylinderArea = 0.001; % Area of caliper cylinder 
+masterCylinderArea = 0.001; % Area of master cylinder 
+brakePedalReduction = 3; % 3:1 mechanical advantage on brake pedal
+
+% -------------------- Solving for Braking forces --------------------- %
+
+% The dynamic of braking can be modeled by four linearly inpendnet
+% equations: 
+%       1) Sum of forces in the x = mass * car acceleration 
+%       2) Sum of forces in the y = 0
+%       3) Sum of the moments about the center of gravity = 0
+%       4) The front braking force will be at max mu*normal force
+% Expressing the following system of equations in maxtrix form yeilds Af=b
[…]

### 2018-10-30 MSXII/Handling Model/torqueModel.m

@@ -1,0 +1,8 @@
+function Fdrive = torqueModel(TMotor, TBraking, Inertia, a, radius)
+
+angAccel = a/radius; 
+%Fy = (Inertia*angAccel + TMotor - TBraking)/radius; 
+Fdrive = (Inertia*angAccel - TMotor + TBraking)/radius; 
+
+
+    

### 2018-10-30 MSXII/SpringRateProject/matlab/Old/createTransferFunctions.m

@@ -1,0 +1,269 @@
+% ---------------------------------------------------------------------- %
+%
+% Suspension Spring Rates Project - createTransferFuncitons.m
+% Function setup transfer functions for a half car model to determine the
+% reponse of MSXII to disturbances in the road conditions
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2017
+%
+%
+% Notes:
+%   
+%   - 
+% ---------------------------------------------------------------------- %
+
+load_system('halfCarModel');
+close all;
+
+% ----------------------------- Constants ------------------------------ %
+
+c1_ = 9670;     % Front damping coefficient (Ns/m)
+c2_ = 9670;     % Rear damping coefficient (Ns/m) 
+k1_ = 650*175.1;    % Front sring rate (N/m) 
+k2_ = 550*175.1;    % Rear sring rate (N/m)
+m_ = 550;       % Mass of the vehicle (kg)
+I_ = 550;       % Moment of interial about global x axis (kg*m^2)
+a_ = 1.3;       % Distance from COG to front tire (m)
+b_ = 1.3;       % Distance from COG to rear tire (m)
+
+% ---------------------- Forcing Function Parameters ------------------- %
+
+amplitude = 0.015;              % Amplitude of bumps on road surface
+bumpSparation = 2;              % Distance between peaks on the road surface (m)
+speed = 22.2;                   % Cruising speed (m/s)
+wb = 2*pi*speed/bumpSparation;  % Frequency of base excitation (rad/sec)
+timeDelay = (a_+b_)/speed;        % delay between input striking front and
+                                % rear wheels (s)
+                                
+% Update simulink parameters
+% set_param('halfCarModel/Road','Frequency',mat2str(wb),'Amplitude',mat2str(amplitude));
+set_param('halfCarModel/DelayA','DelayTime',mat2str(timeDelay));
+set_param('halfCarModel/DelayB','DelayTime',mat2str(timeDelay));
+                              
[…]

### 2018-10-30 MSXII/SteeringLoadingConditions/KingPinAxisTorque.m

@@ -1,0 +1,35 @@
+contactPatch = [0.8,0,0];
+UCAKingPinAxis = [0.706,0.420,0.0133];
+LCAKingPinAxis = [0.740,0.190,0.0068];
+
+fx = [1530,0,0];
+fy = [0,1240,0];
+fz = [0,0,2700];
+
+upright = UCAKingPinAxis - LCAKingPinAxis;
+uprightUnitVector = upright./norm(upright)
+momentArm = contactPatch - LCAKingPinAxis;
+pointOnKingPinAxis = LCAKingPinAxis + dot(momentArm,uprightUnitVector).*uprightUnitVector;
+perpMomentArm = contactPatch - pointOnKingPinAxis;
+
+torqueFx = cross(perpMomentArm,fx);
+torqueFy = cross(perpMomentArm,fy);
+torqueFz = cross(perpMomentArm,fz);
+
+netKingPinAxisTorque = dot(torqueFx,uprightUnitVector) +...
+                       dot(torqueFy,uprightUnitVector) +...
+                       dot(torqueFz,uprightUnitVector)
+
+% Visualize geometry:                    
+% figure; 
+% hold on;
+% plot3(0,0,0,'ko');
+% plot3([-.8;0.8],[0;0],[0;0],'-b');
+% plot3([-.8;0.8],[0;0],[-2.6;-2.6],'-g');
+% plot3([0;0],[0;0],[0;-2.6],'-m');
+% plot3(contactPatch(1),contactPatch(2),contactPatch(3),'*r');
+% plot3(UCAKingPinAxis(1),UCAKingPinAxis(2),UCAKingPinAxis(3),'*b');
+% plot3(LCAKingPinAxis(1),LCAKingPinAxis(2),LCAKingPinAxis(3),'*b');
+% plot3([LCAKingPinAxis(1);UCAKingPinAxis(1)],...
+%       [LCAKingPinAxis(2);UCAKingPinAxis(2)],...
+%       [LCAKingPinAxis(3);UCAKingPinAxis(3)],'-b');

### 2018-10-30 MSXII/SuspensionDynamics/computeCamberVSTravel.m

@@ -1,0 +1,48 @@
+% ---------------------------------------------------------------------- %
+%
+% Suspension - computeCamberVSTravel.m
+% Scripte to greate a fit a polynomial function to camber vs travel  
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2017
+%
+%
+% Notes:
+%   
+%   - 
+%
+% ---------------------------------------------------------------------- %
+
+clc
+clear all
+close all
+
+angleVSTravel = [-35.95, 1.043
+                 -30.21, 1.091
+                 -24.90, 1.128
+                 -23.02, 1.139
+                 -19.79, 1.155
+                 -16.47, 1.170
+                 -14.65, 1.176
+                 -11.15, 1.187
+                 -8.52, 1.193
+                 -6.64, 1.196
+                 -2.92, 1.2
+                 -0.69, 1.201
+                 2.85, 1.199
+                 7.98, 1.193
+                 11.28, 1.185
+                 15.49, 1.172
+                 18.72, 1.16
+                 22.28, 1.143
+                 24.45, 1.132
+                 27.21, 1.115
+                 28.46, 1.108];
+            
+plot(angleVSTravel(:,1),angleVSTravel(:,2));
+coeffs = polyfit(angleVSTravel(:,1),angleVSTravel(:,2),3)
+
+x = floor(angleVSTravel(1,1)):1:floor(angleVSTravel(end,1));
+hold on;
+plot(x,polyval(coeffs,x),'r');

### 2018-10-30 MSXII/SuspensionDynamics/lowerControlArmForces.m

@@ -1,0 +1,16 @@
+% ---------------------------------------------------------------------- %
+%
+% Suspension - lowerControlArmForces.m
+% Function to compute the forces acting on the lower control arm 
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2017
+%
+%
+% Notes:
+%   
+%   - Vector quantities are of the form [x; y; z]
+% ---------------------------------------------------------------------- %
+
+function [fPushrod] = pushRodForce(fUpright, pHardpoint, pUpright, p)

### 2018-10-30 MSXII/SuspensionDynamics/staticScalarTrailingArmForces.m

@@ -1,0 +1,37 @@
+% ---------------------------------------------------------------------- %
+%
+% Suspension - staticTrailingArmForces.m
+% Function to compute the forces acting on the trailing arm 
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2017
+%
+%
+% Notes:
+%   
+%   - MSXII global coordinat system used
+%   - Scalar verison of staticTrailingArmForces()
+%   - Considers loading only in the static case
+%   - Simplifies problem to 2D
+%   - coiloverAngle is the angle that the coilover strut makes with the
+%     central axis of the trailing arm
+%   - trailingArmAngle is the angle the trailing arm makes with horizontal
+% ---------------------------------------------------------------------- %
+
+function [fMountY, fMountZ, fCoilover] = staticScalarTrailingArmForces(fRoadY, fRoadyZ, coiloverAngle,...
+                                                       trailingArmAngle,... 
+                                                       wheelMomentArm,...
+                                                       coiloverLateralMomentArm,...
+                                                       coiloverLongitudianalMomentArm)
+                                                   
+    fCoilover = (fRoadY.*cosd(trailingArmAngle) - fRoadyZ.*cosd(trailingArmAngle)).*wheelMomentArm /...
+                (cosd(trailingArmAngle).*coiloverLongitudianalMomentArm +...
+                 sind(coiloverAngle).*coiloverLateralMomentArm);
+                
+    fCoiloverZ = fCoilover*sind(coiloverAngle-trailingArmAngle);
+    fCoiloverY = fCoilover*cosd(coiloverAngle-trailingArmAngle);
+    
+    fMountZ = fRoadyZ - fCoiloverZ;
+    fMountY = fCoiloverY - fRoadY;
+end

### 2018-10-30 MSXII/SuspensionDynamics/staticTrailingArmForces.m

@@ -1,0 +1,37 @@
+% ---------------------------------------------------------------------- %
+%
+% Suspension - staticTrailingArmForces.m
+% Function to compute the forces acting on the trailing arm 
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2017
+%
+%
+% Notes:
+%   
+%   - Vectors are of the form [x; y; z] in the MSXII coordinate system
+%   - Considers loading only in the static case
+%   - Simplifies problem to 2D
+%   - coiloverAngle is the angle that the coilover strut makes with the
+%     central axis of the trailing arm
+%   - trailingArmAngle is the angle the trailing arm makes with horizontal
+% ---------------------------------------------------------------------- %
+
+function [fMount, fCoilover] = staticTrailingArmForces(fRoad, coiloverAngle,...
+                                                       trailingArmAngle,... 
+                                                       wheelMomentArm,...
+                                                       coiloverLateralMomentArm,...
+                                                       coiloverLongitudianalMomentArm)
+                                                   
+    fCoiloverMag = (fRoad(2)*cosd(trailingArmAngle) - fRoad(3)*cosd(trailingArmAngle))*wheelMomentArm /...
+                   (cosd(trailingArmAngle)*coiloverLongitudianalMomentArm +...
+                    sind(coiloverAngle)*coiloverLateralMomentArm);
+                
+    fCoilover(3) = fCoiloverMag*sind(coiloverAngle-trailingArmAngle);
+    fCoilover(2) = fCoiloverMag*cosd(coiloverAngle-trailingArmAngle);
+    
+    fMount = [0;0;0];
+    fMount(3) = fRoad(3) - fCoilover(3);
+    fMount(2) = fCoilover(2) - fRoad(2);
+end

### 2018-10-30 MSXII/SuspensionDynamics/uprightForces.m

@@ -1,0 +1,43 @@
+% ---------------------------------------------------------------------- %
+%
+% Suspension - uprightForces.m
+% Function to compute the forces acting on the upright 
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2017
+%
+%
+% Notes:
+%   
+%   - Vector quantities are of the form [x; y; z]
+%   - The function solves 6 equations 
+%       - sum of the forces in all 3 axes = mass*accel - applied forces
+%       - sum of the moments in all 3 axes = moment of inertia*accel -
+%       applied moments
+% ---------------------------------------------------------------------- %
+
+function [fUCA, fLCA] = uprightForces(mass, moi, cog, linearAccel,...
+                                      angularAccel, fRoad, posUCA, posLCA,... 
+                                      posContact)
+    % Compute moment arms
+    rUCA = posUCA - cog;
+    rLCA = posLCA - cog;
+    rRoad = posContact - cog;
+    % Compute weight vector
+    weight = mass.*[0;-9.81;0];
+    % Setup linear equations
+    A = [ 1        0        0        1        0        0
+          0        1        0        0        1        0
+          0        0        1        0        0        1
+          0        -rUCA(3) rUCA(2)  0        -rLCA(3) rLCA(2)
+          rUCA(3)  0        -rUCA(1) rLCA(3)  0        -rLCA(1)
+          -rUCA(2) rUCA(1)  0        -rLCA(2) rLCA(1)  0 ];
+    b = [(mass.*linearAccel - fRoad - weight); 
+         (moi.*angularAccel - fRoad.*rRoad)];
+    x = linsovle(A,b);
+    % Parse results
+    fUCA = x(1:3);
+    fLCA = x(4:6);
+
+end

### 2018-10-30 MSXII/SuspensionSimulator/ZZcomputeNodalPositions.m

@@ -1,0 +1,61 @@
+% ---------------------------------------------------------------------- %
+%
+% Double Wishbone Suspension Simulator - computeNodalPositions.m
+% Function to compute x,y position of nodes from properties
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2016
+%
+%
+% Notes:
+%
+%   - The order of params is as follows:
+%        1)  UCA mount (x)
+%        2)  UCA mount (y)
+%        3)  LCA mount (x)
+%        4)  LCA mount (y)
+%        5)  UCA length
+%        6)  LCA length
+%        7)  upright length
+%        8)  spindle height
+%        9)  wheel offset
+%        10) wheel diameter
+%        11) inclination angle
+%
+% ---------------------------------------------------------------------- %
+
+function nodes = computeNodalPositions(params, lcaAngle)
+    
+    % Control arm mounting postions
+    nodes(1,1) = params(1);
+    nodes(1,2) = params(2);
+    nodes(2,1) = params(3);
+    nodes(2,2) = params(4);
+    
+    % 
+    
+    
+    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+    
+    % End of LCA
+    nodes(5,1) = nodes(2,1) + params(6) * cosd(lcaAngle);
+    nodes(5,2) = nodes(2,2) + params(6) * sind(lcaAngle);
+    
+    % End of UCA 
+    [xout,yout] = circcirc(x1,y1,r1,x2,y2,r2)
+    nodes(3,1) = nodes(1,1) + ((nodes(1,1) + UCA - y)*(UCA - Ay + y))^(1/2)
+    
+    % Distance from the UCA mount to the end of the LCA
+    ucaMount2LCAend = sqrt((nodes(1,1)-nodes(5,1))^2 ... 
+                         + (nodes(1,2)-nodes(5,2))^2); % by distance formula
+    
+    % Temp angle is the angle formed btwn the UCA and the line from the UCA
+    % mount to the end of the LCA
+    tempAngle = acosd( (params(5)^2 + ucaMount2LCAend^2 - params(7)^2)...
+                      /(2 * params(5) * ucaMount2LCAend)); % by cosine law
+    
+    % End of UCA 
+    
+    
+end 

### 2018-10-30 MSXII/SuspensionSimulator/computeDistance.m

@@ -1,0 +1,18 @@
+% ---------------------------------------------------------------------- %
+%
+% Double Wishbone Suspension Simulator - computeDistance.m
+% A function that returns the distance between two cartesian points
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2016
+%
+%
+% Notes:
+%
+%   - 
+% ---------------------------------------------------------------------- %
+
+function d = computeDistance(x1,y1,x2,y2)
+    d = sqrt((x1-x2)^2 + (y1-y2)^2);
+end 

### 2018-10-30 MSXII/SuspensionSimulator/computeSuspensionError.m

@@ -1,0 +1,94 @@
+% ---------------------------------------------------------------------- %
+%
+% Double Wishbone Suspension Simulator - computeSuspensionError.m
+% An error function to be minimized by fsolve and solve for nodal positions
+% in the process
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2016
+%
+%
+% Notes:
+%
+%   - parmams is a vector containing the parameters that define the
+%     suspension geometry. guess is an vector of unknown nodal coordinates. 
+%
+%   - The order of params is as follows:
+%        1)  UCA mount (x)
+%        2)  UCA mount (y)
+%        3)  LCA mount (x)
+%        4)  LCA mount (y)
+%        5)  UCA length
+%        6)  LCA length
+%        7)  upright length
+%        8)  spindle height
+%        9)  wheel offset
+%        10) wheel diameter
+%        11) spindle angle
+%
+%    - The order of the guess is as follows 
+%        1)  Node 3 (x)
+%        2)  Node 3 (y)
+%        3)  Node 4 (x)
+%        4)  Node 4 (y)
+%        5)  Node 5 (x)
+%        6)  Node 5 (y)
+%        7)  Node 6 (x)
+%        8)  Node 6 (y)
+%        9)  Node 7 (x)
+%        10) Node 7 (y)
+%        11) Node 8 (x)
+%
+% ---------------------------------------------------------------------- %
+
+function error = computeSuspensionError(guess, params)
+    
+    % The first six equations are driven by the lengths of the elements 
+    % in the suspension (parameters 5 to 10) and are driven by distance
+    % formula. 
+    error(1) = computeDistance(params(1),params(2),guess(1),guess(2)) ...
+               - params(5);  % UCA length
+           
+    error(2) = computeDistance(params(3),params(4),guess(5),guess(6)) ...
+               - params(6);  % LCA length
+           
+    error(3) = computeDistance(guess(1),guess(2),guess(5),guess(6)) ...
+               - params(7);  % upright length
+           
+    error(4) = computeDistance(guess(3),guess(4),guess(5),guess(6)) ...
+               - params(8);  % spindle height
+           
+    error(5) = computeDistance(guess(3),guess(4),guess(9),guess(10)) ...
+               - params(9);  % wheel offset
+           
+    error(6) = computeDistance(guess(7),guess(8),guess(11),0) ...
+               - params(10); % wheel diameter
+    
[…]

### 2018-10-30 MSXII/SuspensionSimulator/cosLawGetOppositeLength.m

@@ -1,0 +1,24 @@
+% ---------------------------------------------------------------------- %
+%
+% Double Wishbone Suspension Simulator - computeSuspensionError.m
+% A function that returns the length of a triangle's legs given the angle
+% opposite to the desigred leg and the lengths of the other two legs. 
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2016
+%
+%
+% Notes:
+%
+%   - Uses the cosine law: http://hyperphysics.phy-astr.gsu.edu/hbase/lcos.html
+%
+%   - angleC is the known angle opposite of leg c in degrees
+%
+%   - a and b are the other two legs of known length
+%
+% ---------------------------------------------------------------------- %
+
+function c = cosLawGetOppositeLength(a,b,angleC)
+    c = sqrt(a^2+b^2-(2*a*b*cosd(angleC)));
+end

### 2018-10-30 MSXII/SuspensionSimulator/getCamber.m

@@ -1,0 +1,21 @@
+% ---------------------------------------------------------------------- %
+%
+% Double Wishbone Suspension Simulator - getCamber.m
+% A function that returns the camber angle in degrees 
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2016
+%
+%
+% Notes:
+%
+%   - Negative camber occurs when the top of the wheel is close to the
+%     center of the car than the bottom
+% 
+% ---------------------------------------------------------------------- %
+
+function camber = getCamber(nodes)
+    camber = atan2d(nodes(8,1)-(nodes(6,1)),(nodes(6,2)-nodes(8,2)));
+end 
+

### 2018-10-30 MSXII/SuspensionSimulator/getKingPinAngle.m

@@ -1,0 +1,19 @@
+% ---------------------------------------------------------------------- %
+%
+% Double Wishbone Suspension Simulator - getKingPinAngle.m
+% A function that returns the king pin angle in degrees 
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2016
+%
+%
+% Notes:
+%
+%   - 
+% 
+% ---------------------------------------------------------------------- %
+
+function kingPinAngle = getKingPinAngle(nodes)
+    kingPinAngle = atan2d(nodes(5,1)-(nodes(3,1)),(nodes(3,2)-nodes(5,2)));
+end 

### 2018-10-30 MSXII/SuspensionSimulator/getLateralScrub.m

@@ -1,0 +1,19 @@
+% ---------------------------------------------------------------------- %
+%
+% Double Wishbone Suspension Simulator - getLateralScrub.m
+% A function that returns the lateral scrub distance in mm 
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2016
+%
+%
+% Notes:
+%
+%   - 
+% 
+% ---------------------------------------------------------------------- %
+
+function lateralScrub = getLateralScrub(nodes,origNodes)
+    lateralScrub = nodes(8,1) - origNodes(8,1);
+end 

### 2018-10-30 MSXII/SuspensionSimulator/getScrubRadius.m

@@ -1,0 +1,21 @@
+% ---------------------------------------------------------------------- %
+%
+% Double Wishbone Suspension Simulator - getScrubRadius.m
+% A function that returns the scrub radius in mm 
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2016
+%
+%
+% Notes:
+%
+%   - 
+% 
+% ---------------------------------------------------------------------- %
+
+function scrubRadius = getScrubRadius(nodes)
+    dXdY = (nodes(3,1) - nodes(5,1)) / (nodes(3,2) - nodes(5,2));
+    roadIntersectionPoint = nodes(5,1) - dXdY * nodes(5,2); 
+    scrubRadius = nodes(8,1) - roadIntersectionPoint;
+end 

### 2018-10-30 MSXII/SuspensionSimulator/printLabels.m

@@ -1,0 +1,32 @@
+% ---------------------------------------------------------------------- %
+%
+% Double Wishbone Suspension Simulator - printRoad.m
+% Fuction to add the road to the visualizer plot and format it
+% 
+%
+% Notes:
+%
+%   - 
+%
+% ---------------------------------------------------------------------- %
+
+function printLabels(limits) 
+
+    % Define plot limits
+    xlimits = limits(1,:); 
+    ylimits = limits(2,:); 
+    
+    % Show ground
+    plot([xlimits(1)*10 xlimits(2)*10],[0 0],'k--');
+    
+    % Legend
+    legend('Neutral Position','Dynamic','Location','southoutside');
+    
+    hold off;
+    xlabel('X-Position (mm)');
+    ylabel('Y-Position (mm)');
+    xlim(xlimits);
+    ylim(ylimits);
+    title('Suspension Visualizer');
+end
+    

### 2018-10-30 MSXII/SuspensionSimulator/printSuspension.m

@@ -1,0 +1,68 @@
+% ---------------------------------------------------------------------- %
+%
+% Double Wishbone Suspension Simulator - printSuspension.m
+% Fuction to print the current state of the suspension to a figure
+% 
+% Original version by                                                
+%   Dr. Haim Waisman, Columbia University, New York, Copyright 2009    
+% Modified 2013                                                      
+%   Dr. Robert Gracie, University of Waterloo
+% Modified 2014
+%   Christopher Kohar, University of Waterloo
+% Modified 2016
+%   Devon Copeland, Midnight Sun, University of Waterloo
+%
+% Notes:
+%
+%   - nodes is a matrix with n rows of the format x,y representing n nodes
+%   - initialState is a matrix of nodes at the suspensions neutral position
+%   - elmntMap maps has m rows of the format node1,node2 representing
+%     the m element in the suspension
+%
+% ---------------------------------------------------------------------- %
+
+function printSuspension(initialState,nodes,elmntMap,limits) 
+
+    numbElmnts = size(elmntMap,1); % number of elements
+    xInit = initialState(:,1); % isolate x coordinates
+    yInit = initialState(:,2); % isolate y coordinates
+    x = nodes(:,1); % isolate x coordinates
+    y = nodes(:,2); % isolate y coordinates
+
+    % Define plot limits
+    xlimits = limits(1,:); 
+    ylimits = limits(2,:); 
+    
+    % Plot the structure
+    for i = 1:numbElmnts
+
+        % Plot initial state of the suspension
+        XXI = [xInit(elmntMap(i,1)), xInit(elmntMap(i,2))];
+        YYI = [yInit(elmntMap(i,1)), yInit(elmntMap(i,2))];
+        line(XXI,YYI,'LineWidth',1);
+        hold on;
+
+        % PLOT FINAL DEFORMED STRUCTURE
+        XXF = [x(elmntMap(i,1)), x(elmntMap(i,2))];
+        YYF = [y(elmntMap(i,1)), y(elmntMap(i,2))];
+        line(XXF,YYF,'LineWidth',2,'Color','RED','LineStyle','--');hold on;
+        % Initial Node Location Labels
+        text(XXI(1),YYI(1),sprintf('%0.5g',elmntMap(i,1)));
[…]

### 2018-10-30 MSXII/SuspensionSimulator/rotateCoords.m

@@ -1,0 +1,26 @@
+% ---------------------------------------------------------------------- %
+%
+% Double Wishbone Suspension Simulator - rotateCoords.m
+% A function that rotates an input coordinate about a specified center of
+% rotation by a specified angle in degrees
+% 
+% Author: Devon Copeland
+% 
+% © Midnight Sun 2016
+%
+%
+% Notes:
+%
+%   - Positive angles are CCW
+% 
+% ---------------------------------------------------------------------- %
+
+function [x,y] = rotateCoords(inputCoords, centerOfRotation, angle)
+    inputCoords = inputCoords - centerOfRotation;
+    [theta,rho] = cart2pol(inputCoords(1),inputCoords(2));
+    theta = theta + deg2rad(angle); 
+    [outputCoords(1),outputCoords(2)] = pol2cart(theta,rho);
+    outputCoords = outputCoords + centerOfRotation;
+    x = outputCoords(1);
+    y = outputCoords(2);
+end 

### 2018-10-30 MSXIV/MassBrakesModel/main.m
Initial commit of brakes solution space model

@@ -1,0 +1,53 @@
+% ---------------------------------------------------------------------- %
+%
+% Brakes & CoM Solution Space - main.m
+% Script to develop solution space for front/back CoM position of MSXIV
+% 
+% Author: Neeraj Nair
+% 
+% © Midnight Sun 2018
+%
+%
+% Notes:
+%
+%   - Units: 
+%       Distance: millimeters
+%       Mass: kilograms
+%       Time: seconds
+%       Angles: degrees
+%
+% ---------------------------------------------------------------------- %
+
+Length = 0; % Length of wheelbase 
+a = 0; % Distance of Center of Gravity from front wheel
+Mass = 0; % Mass of car
+Nf = 0; % Normal reaction force of front wheels
+Nr = 0; % Normal reaction force of rear wheels 
+Dc = 0; % Deceleration of car 
+t= 0; % Braking time of car 
+h = 0; % Distance between ground and Center of Gravity 
+ 
+ 
+% For this case, we assume Length = 2.25 m, Dc = -4.47, t = 3 
+
+Length = 2.25;
+Dc = -4.47;
+time  = 3;
+h = 1.5;
+
+
+syms Mass Nf Nr
+Nr = ((Mass * 9.81) - Nf) /1000
+
+ 
+Nf = (((Mass * 9.81 * (Length - a)) + (Mass * Dc * h))/Length)/1000
+ 
+n = 10;
+m = .25;
+o =10;
+
+
+[xx,yy,zz] = meshgrid(0:n:800,0:m:2.25,-100:o:100);
+
+xx = Mass;
+yy = a; 

### 2018-10-30 MSXIV/BrakesSolutionSpace/main.m
Update MSXIV/BrakesSolutionSpace/main.m
Removed plot command (test commit)

@@ -44,10 +44,10 @@ Nf = (((Mass * 9.81 * (Length - a)) + (Mass * Dc * h))/Length)/1000
  
 n = 10;
 m = .25;
-o =10;
+% o =10;
 
 
-[xx,yy,zz] = meshgrid(0:n:800,0:m:2.25,-100:o:100);
-
-xx = Mass;
-yy = a; 
+% % [xx,yy,zz] = meshgrid(0:n:800,0:m:2.25);
+% 
+% xx = Mass;
+% yy = a; 

### 2018-11-02 MSXIV/BrakesSolutionSpace/main.m
Update MSXIV/BrakesSolutionSpace/main.m
- Improved code readability and formatting
- Added labels to plot and set grid properties

@@ -16,38 +16,47 @@
 %       Time: seconds
 %       Angles: degrees
 %
-% ---------------------------------------------------------------------- %
 
-Length = 0; % Length of wheelbase 
-a = 0; % Distance of Center of Gravity from front wheel
-Mass = 0; % Mass of car
-Nf = 0; % Normal reaction force of front wheels
-Nr = 0; % Normal reaction force of rear wheels 
-Dc = 0; % Deceleration of car 
-t= 0; % Braking time of car 
-h = 0; % Distance between ground and Center of Gravity 
- 
- 
+% --------------------------- Constants -------------------------------- %
+
 % For this case, we assume Length = 2.25 m, Dc = -4.47, t = 3 
 
-Length = 2.25;
-Dc = -4.47;
-time  = 3;
-h = 1.5;
+Length = 2.25;          % Length of wheelbase
+Dc = -4.47;             % Deceleration of car
+T  = 3;                 % Braking time of car 
+H = 1.5;                % Distance between ground and Center of Gravity 
+k = 1;                  % Counter for  loop
 
+% --------------------------- Variables -------------------------------- %
 
-syms Mass Nf Nr
-Nr = ((Mass * 9.81) - Nf) /1000
+Mass2= zeros(1,40500);
+A2= zeros(1,40500);
+Nf= zeros(1,40500);     % Normal reaction force of front wheels
+Nr = 0;                 % Normal reaction force of rear wheels 
+               
 
  
-Nf = (((Mass * 9.81 * (Length - a)) + (Mass * Dc * h))/Length)/1000
- 
-n = 10;
-m = .25;
-% o =10;
+% Populating Nf values for all possilbe combinations of weight and alpha
 
+for Mass = 700: 1: 900
 
-% % [xx,yy,zz] = meshgrid(0:n:800,0:m:2.25);
-% 
-% xx = Mass;
-% yy = a; 
+   for A= 0.1: 0.00825 : 1.75
+Nf(k) = (((Mass* 9.81 * (Length - A ))+(Mass * Dc * H)) / Length)/1000;
+Mass2(k)= Mass;
+A2(k)= A;
+k = k +1;
+  end
+
+end
+
+% Plot for feasible space
+
+figure
+plot3(A2, Mass2, Nf)
+xlabel({'Distance of COG from front wheels, A','(Meters)'},'Color','r')
[…]
